### 2016-04-23 report/paper.tex
Add Latex template

\documentclass[10pt,conference,a4paper]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Correct date format in references but with american hyphenation, quotes, ...
% Trick from http://tex.stackexchange.com/a/129209
\usepackage[australian,american]{babel}

% *** GRAPHICS RELATED PACKAGES ***
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
\else
  \usepackage[dvips]{graphicx}
\fi

% *** MATH PACKAGES ***
\usepackage{amssymb}
\usepackage[cmex10]{amsmath}

% Figure captions in small font
\makeatletter
\let\MYcaption\@makecaption
\makeatother
\usepackage[font=footnotesize]{subcaption}
\makeatletter
\let\@makecaption\MYcaption
\makeatother

\usepackage{pgfplots}
\pgfplotsset{compat=1.11}
%TODO minted ?
\usepackage[binary-units]{siunitx}
\usepackage{booktabs}
\usepackage{pifont}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage[backend=biber,style=ieee,minbibnames=1,maxbibnames=3]{biblatex}
\usepackage{url}
%\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage[hidelinks]{hyperref}

\bibliography{paper}
% Small font for references
\renewcommand*{\bibfont}{\small}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% paper title
% can use linebreaks \\ within to get better formatting as desired
\def\mytitle{EpTO: An Epidemic Total Order Algorithm for Large-Scale Distributed Systems}
\title{\mytitle}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\def\jocelyn{Jocelyn Thode}
\def\ehsan{Ehsan Farhadi}
\author{\IEEEauthorblockN{\jocelyn\IEEEauthorrefmark{1},
\ehsan\IEEEauthorrefmark{2}}
\IEEEauthorblockA{Université de Neuchâtel\\
Neuchâtel, Switzerland\\
Email: \IEEEauthorrefmark{1}\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch},
\IEEEauthorrefmark{2}\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}}}

% Set PDF file properties
\hypersetup{
	pdftitle=\mytitle,
}

% Tick marks
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

% Columns balancing
\renewbibmacro{finentry}{%
  \iffieldequalstr{entrykey}{jujjuri2010virtfs}%<- key after which you want the break
   {\finentry\newpage}
   {\finentry}}

\begin{document}
\graphicspath{{figures/}}

% make the title area
\maketitle

\begin{abstract}
%TODO
asdfasdf
\end{abstract}

\section{Introduction}
%TODO

\section{TODO}

\begin{figure}
\includegraphics[width=\linewidth]{figures/epto-architecture.png}
\caption{test\protect\footnotemark{}.}
\label{fig:test}
\end{figure}

\section{Conclusion}
%TODO

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

% references section, with correct date format
\begin{otherlanguage}{australian}
\printbibliography
\end{otherlanguage}

% that's all folks
\end{document}

### 2016-04-23 src/src/epto/utilities/App.java
add App class

@@ -3,9 +3,16 @@ package epto.utilities;
 
 import net.sf.neem.impl.Application;
 
-import java.nio.ByteBuffer;
+    import java.nio.ByteBuffer;
 
+/**
+ * Implementation of an Application
+ */
 public class App implements Application{
+
+    /**
+     * {@inheritDoc}
+     */
     @Override
     public void deliver(ByteBuffer[] byteBuffers) {
 

### 2016-04-24 report/paper.tex
report#1

@@
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.11}
+\pgfplotsset{compat=1.9}
 %TODO minted ?
@@
 
-\bibliography{paper}
+\addbibresource{•}{paper.bib}
 % Small font for references
@@
 \begin{abstract}
-%TODO
-asdfasdf
+One of the fundamental problems of distributed computing, is the ordering of events through all peers.
+From all the available orderings, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
+unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.	 EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
+EpTO provides deterministic safety and probabilistic liveness: integrity, total order and validity are always preserved, while agreement is achieved with arbitrarily high probability.
+We are going to implement EpTO using NeEM library and show EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with currently being used ordering algorithms.
 \end{abstract}
@@
 %TODO
+The ordering of events is one of the most fundamental problems in distributed systems, and it has been studied over past few decades.
+A lot of researchers have been working on an algorithm with different guarantees and tradeoffs such as synchronization, agreement or state machine replication.
+But because these properties are strong guarantees, the algorithms that implements them do not scale very well.
+Existing probabilistic protocols are highly scalable and resilient, but they can not provide total ordering in large scale distributed systems, and existing deterministic protocols which provide total ordering, are not resilient and scalable.
+\par
+The problem with current deterministic total ordering protocols, deterministic total ordering algorithm, is that they need some sort of agreement between all peers in the system.
+An agreement on the order of messages, which cause a massive amount of network traffic and overhead on the system.
+Plus, an agreement feature for an asynchronous system requires to
+explicitly maintain a group and have access to a failure detector.
+Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits scalability of the algorithm.
+\subsection{Contribution}
+EpTO guarantees that processes eventually agree on the set of received events with high probability and deliver these events in total order to the application.
+The intuition behind EpTO is that events are available quickly at all nodes with high probability,
+Once events are thought to be available in every peer, each peer deterministically order them by timestamp of each event and breaking ties with the id of the broadcaster peer, and deliver them to the application accordingly.
+\par
+The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach.
+\textit{Balls-and-bins} is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
+\par
+A balls-and-bins approach abstract peers as bins and messages (events) as balls, and studies how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
+using this approach The number of messages transmitted per process per round is logarithmic in the number of processes, and the total number of messages
+transmitted in the network before an event is delivered is low and uniform over all peers.
 
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/BnB.jpeg}
+\caption{balls-and-bins\protect\footnotemark{}.}
+\label{fig:balls-and-bins}
+\end{figure}
+
+\section{EPTO ALGORITHM DESCRIPTION}
+\subsubsection{EpTO Dissemination Component}
+\subsubsection{EpTO Ordering Component}
+\subsubsection{EpTO Stability Oracle}
 \section{TODO}
-
+\section{Work Plan}
 \begin{figure}

### 2016-04-24 report/paper.tex
report#1

@@
 
-\addbibresource{•}{paper.bib}
+\addbibresource{paper.bib}
 % Small font for references
@@
 \par
-The problem with current deterministic total ordering protocols, deterministic total ordering algorithm, is that they need some sort of agreement between all peers in the system.
+The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system.
 An agreement on the order of messages, which cause a massive amount of network traffic and overhead on the system.

### 2016-04-24 report/paper.tex
Add skeleton

--- 2016-04-23 report/paper.tex
@@
 \usepackage{url}
-%\usepackage[backgroundcolor=lightgray]{todonotes}
+\usepackage[backgroundcolor=lightgray]{todonotes}
 \usepackage[hidelinks]{hyperref}
@@
 \begin{abstract}
-%TODO
-asdfasdf
+One of the fundamental problems of distributed computing, is the ordering of events through all peers.
+From all the available orderings, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
+unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.
+EpTO \autocite{matos2015epto} is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
+EpTO provides deterministic safety and probabilistic liveness: integrity, total order and validity are always preserved, while agreement is achieved with arbitrarily high probability.
+We are going to implement EpTO using the NeEM \autocite{neem} library and show EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with currently being used ordering algorithms.
 \end{abstract}
@@
 \section{Introduction}
-%TODO
-
-\section{TODO}
+The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over past few decades.
+A lot of researchers have been working on an algorithm with different guarantees and trade-offs such as synchronization, agreement or state machine replication \todo{Cite papers}
+.
+But because these properties are strong guarantees, the algorithms that implements them do not scale very well.
+Existing probabilistic protocols are highly scalable and resilient, but they can not provide total ordering in large scale distributed systems, and existing deterministic protocols which provide total ordering, are not resilient and are scalable.
+\par
+The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system on the order of messages.
+This causes a massive amount of network traffic and overhead on the system.
+Moreover, an agreement feature for an asynchronous system requires to
+explicitly maintain a group and have access to a failure detector.
+Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits scalability of the algorithm.
+\subsection{Contribution}
+EpTO guarantees that processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
+The intuition behind EpTO is that events are available quickly at all nodes with high probability,
+Once events are thought to be available in every peer, each peer deterministically order them using the timestamp of each event and breaking ties with the id of the broadcaster peer, deliver them to the application accordingly.
+\par
+The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach.
+The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
+\par
+A balls-and-bins approach abstracts peers as bins and messages (events) as balls, and studies how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
+Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, and the total number of messages
+transmitted in the network before an event is delivered is low and uniform over all peers.
 
 \begin{figure}
+\includegraphics[width=\linewidth]{figures/BnB.jpeg}
+\caption{Balls-and-Bins\protect\footnotemark{}.}
+\label{fig:balls-and-bins}
+\end{figure}
+\par
+In this paper, we will implement EpTO and compare it against a known deterministic total order algorithm, namely JGroups \autocite{jgroups}.
+
+These tasks will be split in two tasks.
+The first task will assume that all peers are known by all peers.
+We will implement EpTO using this assumption and test it against JGroups.
+The second task will assume unknown membership on all peers.
+We will update EpTO and use a PSS\footnote{Peer Sampling Service}.
+We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
+
+\section{Ordering Algorithms}
+\todo[inline]{We will explain why Ordering algorithms are necessary, the difference between partial order and Total order and why partial order might not be sufficient in some cases}
+\subsection{Partial Order Algorithms}
+\subsection{Total Order Algorithms}
+
+\section{EpTO}
+\todo[inline]{Here we will present EpTO and the different components}
+\begin{figure}
 	\includegraphics[width=\linewidth]{figures/epto-architecture.png}
-\caption{test\protect\footnotemark{}.}
-\label{fig:test}
+	\caption{\todo[inline]{We need a better quality figure (SVG)}\protect\footnotemark{}.}
+	\label{fig:epto-architecture}
 \end{figure}
+\subsection{Definitions}
+\todo[inline]{In this section we will define what events and process means in the context of EpTO}
+\subsection{EpTO Dissemination Component}
+\todo[inline]{We will explain what is the role of the dissemination component}
+\subsection{EpTO Ordering Component}
+\todo[inline]{We will explain what is the role of the ordering component}
+\subsection{EpTO Stability Oracle}
+\todo[inline]{We will explain what is the role of the stability clock and how it doesn't affect performances}
+\section{Performance Comparisons}
+\todo[inline]{Here we will present the different test we ran, the methodology used and the result we obtained}
+\subsection{Methodology}
+\todo[inline]{Here we explain how we ran our tests and on what type of machines}
+\subsection{Peer membership known}
+\todo[inline]{We will present the results obtained and try to explain the results}
+\subsection{Peer membership unknown}
+\todo[inline]{We will present the results obtained and try to explain the results}
 
 \section{Conclusion}
-%TODO
+\todo[inline]{He we will summarize the results we found and present some future tasks that could be accomplished on EpTO}
 

### 2016-04-24 report/paper.tex
Add some text

@@
 \subsection{Definitions}
-\todo[inline]{In this section we will define what events and process means in the context of EpTO}
-\subsection{EpTO Dissemination Component}
-\todo[inline]{We will explain what is the role of the dissemination component}
-\subsection{EpTO Ordering Component}
-\todo[inline]{We will explain what is the role of the ordering component}
-\subsection{EpTO Stability Oracle}
-\todo[inline]{We will explain what is the role of the stability clock and how it doesn't affect performances}
+A process or peer is defined as an actor in our system running the application that needs total order.
+Each process will communicate with other processes in the distributed systems and exchange events and order them together.
+\par
+An event is defined as something that happens at a given time.
+For example, we could imagine a system where each process publish some sort of data.
+The moment where this data is published combined with the data is called an event.
+\par
+A ball is a set of events bundled together and sent as one package. We use balls in EpTO to reduce network traffic and make it scalable in terms of processes and events.
+
+EpTO is scalable in the sense that \todo[inline]{Explain the logarithmic scalability}
+
+\todo[inline]{Maybe speak about holes created by the probabilistic nature of the algorithm}
+\subsection{Dissemination Component}
+The Dissemination component is the component that bridges EpTO with the rest of the network.
+As we can see in Figure \ref*{fig:epto-architecture}.
+It is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the fan-out of the gossip.
+\par
+When an application wants to publish an event, it will broadcast this event to the Dissemination component.
+
+\subsection{Ordering Component}
+The Ordering component is the component that orders the events before delivering them to the application.
+\todo[inline]{Explain how it orders the different events and how it can order them with a high probability}
+\subsection{Stability Oracle}
+The Stability oracle is the component that outputs timestamps.
+It will increment its local clock every round as well as synchronize itself using timestamps of newly received events, to make sure the local clock does not drift too much.
+\par
+This Stability oracle offers an API to the Ordering component letting it know when an event is mature enough to be delivered to the application.
+\par
+As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it doesn't impact the scalability of EpTO.
 \section{Performance Comparisons}

### 2016-04-25 report/paper.tex
add ordering component to paper

@@
 \par
-A balls-and-bins approach abstracts peers as bins and messages (events) as balls, and studies how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
+A balls-and-bins approach abstracts peers as bins and messages (events) as balls, and studies how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability \autocite{Koldehofe02simplegossiping}.
 Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, and the total number of messages
@@
 \includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins\protect\footnotemark{}.}
+\caption{Balls-and-Bins}
 \label{fig:balls-and-bins}
@@
 \todo[inline]{We will explain why Ordering algorithms are necessary, the difference between partial order and Total order and why partial order might not be sufficient in some cases}
+
+Distributed systems, like centralized systems needs to preserve the temporal order of event produced by concurrent process in the system.
+When there are spatially separated processes that can only communicate through messages, you can’t tell what in the system happened before something else.
+Therefore we need ordering algorithms to overcome this problem.
+\par
+We have two types of order \autocite{lamport1978time}:
 \subsection{Partial Order Algorithms}
+Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
+\begin{itemize}
+	\item Reflexivity: $a \leq a$ for all $a \in S$.
+	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
+	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
+\end{itemize}
+
 \subsection{Total Order Algorithms}
+A totally ordered set of events is a partially ordered set which satisfies one additional property:
+\begin{itemize}
+	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
+\end{itemize}
 
@@
 \todo[inline]{Maybe speak about holes created by the probabilistic nature of the algorithm}
+\par
+Since EpTO is using a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
+In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered event will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
 \subsection{Dissemination Component}
@@
 \subsection{Ordering Component}
-The Ordering component is the component that orders the events before delivering them to the application.
-\todo[inline]{Explain how it orders the different events and how it can order them with a high probability}
+The Ordering component is responsible for orders events before delivering them to the application.
+To do so, this component has a \textit{received} hashmap of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hashmap containing all the events which are delivered.
+\par
+In brief, ordering component first increment the timestamp of all the events which has been received in previous rounds to indicate the start of a new round.
+Then it process new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event) to prevent delivering duplicate event, remaining event in the received ball will add to the \textit{received} hashmap, and wait to be delivered based on the stability oracle and other undelivered events on each round.
 \subsection{Stability Oracle}

### 2016-04-25 report/paper.tex
Fix references in Abstract

@@
 unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.
-EpTO \autocite{matos2015epto} is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
+EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
 EpTO provides deterministic safety and probabilistic liveness: integrity, total order and validity are always preserved, while agreement is achieved with arbitrarily high probability.
-We are going to implement EpTO using the NeEM \autocite{neem} library and show EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with currently being used ordering algorithms.
+We are going to implement EpTO using the NeEM library and show EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with currently being used ordering algorithms.
 \end{abstract}
@@
 \subsection{Contribution}
-EpTO guarantees that processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
+EpTO \autocite{matos2015epto} guarantees that processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
 The intuition behind EpTO is that events are available quickly at all nodes with high probability,
@@
 We will update EpTO and use a PSS\footnote{Peer Sampling Service}.
-We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
+We will also update the underlying layer NeEM \autocite{neem} to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
 

### 2016-04-25 report/paper.tex
Correct errors

@@
 
-Distributed systems, like centralized systems needs to preserve the temporal order of event produced by concurrent process in the system.
-When there are spatially separated processes that can only communicate through messages, you can’t tell what in the system happened before something else.
+Distributed systems, like centralized systems need to preserve the temporal order of event produced by concurrent process in the system.
+When there are separated processes that can only communicate through messages, you can’t easily order these messages.
 Therefore we need ordering algorithms to overcome this problem.
 \par
-We have two types of order \autocite{lamport1978time}:
+We have two types of ordering algorithms \autocite{lamport1978time}, the partial order algorithms and the total order algorithms.
 \subsection{Partial Order Algorithms}
@@
 \section{EpTO}
-\todo[inline]{Here we will present EpTO and the different components}
 \begin{figure}
@@
 
-\todo[inline]{Maybe speak about holes created by the probabilistic nature of the algorithm}
 \par
@@
 The Dissemination component is the component that bridges EpTO with the rest of the network.
-As we can see in Figure \ref*{fig:epto-architecture}.
-It is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the fan-out of the gossip.
+As we can see in Figure \ref*{fig:epto-architecture}, it is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the gossip fan-out.
 \par
@@
 \subsection{Ordering Component}
-The Ordering component is responsible for orders events before delivering them to the application.
-To do so, this component has a \textit{received} hashmap of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hashmap containing all the events which are delivered.
+The Ordering component is responsible for ordering events before delivering them to the application.
+To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which are delivered.
 \par
-In brief, ordering component first increment the timestamp of all the events which has been received in previous rounds to indicate the start of a new round.
-Then it process new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event) to prevent delivering duplicate event, remaining event in the received ball will add to the \textit{received} hashmap, and wait to be delivered based on the stability oracle and other undelivered events on each round.
+In brief, the Ordering component first increments the timestamp of all the events which have been received in previous rounds to indicate the start of a new round.
+Then, it processes new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event).
+This is done to prevent delivering duplicate events.
+The remaining events in the received ball will be added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
 \subsection{Stability Oracle}
@@
 \par
-As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it doesn't impact the scalability of EpTO.
+As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
 \section{Performance Comparisons}
@@
 \section*{Acknowledgment}
+\todo[inline]{Write acknowledgment}
 

### 2016-04-25 report/paper.tex
Add column
Add a todo

@@
 \par
-We have two types of ordering algorithms \autocite{lamport1978time}, the partial order algorithms and the total order algorithms.
+We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
 \subsection{Partial Order Algorithms}
@@
 \end{itemize}
+\todo[inline]{Explain advantages and usefulness of both systems}
 

### 2016-04-26 report/paper.tex
change senteces from orihinal paper

@@
 One of the fundamental problems of distributed computing, is the ordering of events through all peers.
-From all the available orderings, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
+Between different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
 unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.
 EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
-EpTO provides deterministic safety and probabilistic liveness: integrity, total order and validity are always preserved, while agreement is achieved with arbitrarily high probability.
-We are going to implement EpTO using the NeEM library and show EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with currently being used ordering algorithms.
+EpTO uses a probabilistic agreement to disseminate events through the system with high probability, and also provides integrity, total order and validity.
+We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with a currently being used ordering algorithm.
 \end{abstract}
@@
 The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over past few decades.
-A lot of researchers have been working on an algorithm with different guarantees and trade-offs such as synchronization, agreement or state machine replication \todo{Cite papers}
-.
-But because these properties are strong guarantees, the algorithms that implements them do not scale very well.
-Existing probabilistic protocols are highly scalable and resilient, but they can not provide total ordering in large scale distributed systems, and existing deterministic protocols which provide total ordering, are not resilient and are scalable.
+Many researches have been done to create algorithms with different properties and trade-offs such as synchronization, agreement or state machine replication \autocites[]{defago2004total}[]{lamport1978time}.
+But due the fact that these properties are strong guarantees, the algorithms that implements them do not scale as expected because of the churn and failures in an non-ideal environment.
+Because of scalability problem of deterministic approaches, probabilistic protocols emerged.
+Nowadays, existing probabilistic protocols are highly scalable and resilient, but they do not provide total ordering.
 \par
@@
 \subsection{Contribution}
-EpTO \autocite{matos2015epto} guarantees that processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
-The intuition behind EpTO is that events are available quickly at all nodes with high probability,
-Once events are thought to be available in every peer, each peer deterministically order them using the timestamp of each event and breaking ties with the id of the broadcaster peer, deliver them to the application accordingly.
+EpTO \autocite{matos2015epto} guarantees that all processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
+The intuition behind EpTO is make events available in all processes quickly with high probability,
+Once events are available in every peer, each peer deterministically order them using the timestamp of each event and breaking ties with the id of the broadcaster peer, and afterwards deliver them to the application accordingly.
 \par
-The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach.
-The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
+The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+\textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
 \par
-A balls-and-bins approach abstracts peers as bins and messages (events) as balls, and studies how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability \autocite{Koldehofe02simplegossiping}.
-Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, and the total number of messages
-transmitted in the network before an event is delivered is low and uniform over all peers.
+Using balls-and-bins approach we model processes an bins and events (or set of events) as balls.
+Base on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
+Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore number of messages sent in the system is low and uniform over all processes.
+Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provide total order.
 
@@
 \todo[inline]{Explain advantages and usefulness of both systems}
+\par
+In other words, total order is an ordering that defines the exact order of every event in the system, But partial ordering only defines the order between certain key event that depend on each other.
+Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example in order to have an exact image of the state of the system at all time, we have to use total order.
 

### 2016-04-26 report/paper.tex
Fix errors

@@
 One of the fundamental problems of distributed computing, is the ordering of events through all peers.
-Between different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
+Among different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
 unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.
@@
 EpTO uses a probabilistic agreement to disseminate events through the system with high probability, and also provides integrity, total order and validity.
-We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with a currently being used ordering algorithm.
+We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems.
+Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups.
 \end{abstract}
@@
 \section{Introduction}
-The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over past few decades.
+The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over the past few decades.
 Many researches have been done to create algorithms with different properties and trade-offs such as synchronization, agreement or state machine replication \autocites[]{defago2004total}[]{lamport1978time}.
-But due the fact that these properties are strong guarantees, the algorithms that implements them do not scale as expected because of the churn and failures in an non-ideal environment.
-Because of scalability problem of deterministic approaches, probabilistic protocols emerged.
+Unfortunately, due the fact that these properties are strong guarantees, the algorithms that implement them do not scale as expected because of churn and failures in an non-ideal environment.
+Because of the scalability problem inherent to deterministic approaches, probabilistic protocols have emerged.
 Nowadays, existing probabilistic protocols are highly scalable and resilient, but they do not provide total ordering.
+\todo[inline]{Maybe add a paper to justify this claim}
 \par
@@
 explicitly maintain a group and have access to a failure detector.
-Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits scalability of the algorithm.
+Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits the scalability of the algorithm.
 \subsection{Contribution}
 EpTO \autocite{matos2015epto} guarantees that all processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
-The intuition behind EpTO is make events available in all processes quickly with high probability,
-Once events are available in every peer, each peer deterministically order them using the timestamp of each event and breaking ties with the id of the broadcaster peer, and afterwards deliver them to the application accordingly.
+The intuition behind EpTO is to make events available in all processes quickly with high probability.
+Once events are thought to be available in every peer, each peer deterministically orders them using the timestamp of each event and breaking ties with the id of the broadcaster peer, delivers them to the application accordingly.
 \par
 The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
-\textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
+The \textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
 \par
-Using balls-and-bins approach we model processes an bins and events (or set of events) as balls.
-Base on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
-Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore number of messages sent in the system is low and uniform over all processes.
-Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provide total order.
+Using the balls-and-bins approach we model processes as bins and events (or set of events) as balls.
+Based on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
+Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
+Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.
 
@@
 \par
-In other words, total order is an ordering that defines the exact order of every event in the system, But partial ordering only defines the order between certain key event that depend on each other.
+In other words, total order is an ordering that defines the exact order of every event in the system.
+On the other hand, partial ordering only defines the order between certain key events that depend on each other.
 Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example in order to have an exact image of the state of the system at all time, we have to use total order.

### 2016-04-26 src/src/epto/Peer.java
Fix ByteBuffer read

@@ -52,7 +52,7 @@ public class Peer implements Runnable{
                 ByteBuffer bb = ByteBuffer.wrap(buf);
 
                 neem.read(bb);
-                ByteArrayInputStream byteIn = new ByteArrayInputStream(byteOut.toByteArray()); //TODO fix this
+                ByteArrayInputStream byteIn = new ByteArrayInputStream(buf); //TODO Check correctness
                 ObjectInputStream in = new ObjectInputStream(byteIn);
                 disseminationComponent.receive((HashMap<UUID, Event>) in.readObject());
             }

### 2016-04-26 src/src/epto/utilities/Event.java
Add a toString method

@@ -117,4 +117,17 @@ public class Event implements Comparable<Event> {
         return EQUAL;
 
     }
+
+    /**
+     * @return A string representing the Event object
+     */
+    @Override
+    public String toString() {
+        return "Event{" +
+                "id=" + id +
+                ", timeStamp=" + timeStamp +
+                ", ttl=" + ttl +
+                ", sourceId=" + sourceId +
+                '}';
+    }
 }

### 2016-04-26 src/src/epto/DisseminationComponent.java
Fix some issues in the components

@@ -2,7 +2,6 @@ package epto;
 
 import epto.utilities.Event;
 import net.sf.neem.MulticastChannel;
-import net.sf.neem.impl.Overlay;
 import net.sf.neem.impl.Periodic;
 import net.sf.neem.impl.Transport;
 
@@ -11,7 +10,9 @@ import java.io.IOException;
 import java.io.ObjectOutputStream;
 import java.nio.ByteBuffer;
 import java.nio.channels.ClosedChannelException;
-import java.util.*;
+import java.util.HashMap;
+import java.util.Random;
+import java.util.UUID;
 
 /**
  * Implementation of the Dissemination Component  of EpTO. This class is in charge of

### 2016-04-26 src/src/epto/OrderingComponent.java

@@ -1,13 +1,12 @@
 package epto;
 
-import epto.utilities.Event;
 import epto.utilities.App;
+import epto.utilities.Event;
 
 import java.io.ByteArrayOutputStream;
 import java.io.IOException;
 import java.io.ObjectOutputStream;
 import java.nio.ByteBuffer;
-import java.nio.channels.ClosedChannelException;
 import java.util.HashMap;
 import java.util.UUID;
 
@@ -109,9 +108,9 @@ public class OrderingComponent {
                 e.printStackTrace();
             }
             //delivering the event
-            ByteBuffer[] byteoutArray = new ByteBuffer[1];
-            byteoutArray[0] = ByteBuffer.wrap(byteOut.toByteArray());
-            app.deliver(byteoutArray);
+            ByteBuffer[] byteOutArray = new ByteBuffer[1];
+            byteOutArray[0] = ByteBuffer.wrap(byteOut.toByteArray());
+            app.deliver(byteOutArray);
 
         }
 

### 2016-04-26 src/src/epto/Peer.java
Start workin on the App logic

@@ -1,8 +1,8 @@
 package epto;
 
+import epto.utilities.App;
 import epto.utilities.Event;
 import net.sf.neem.MulticastChannel;
-import net.sf.neem.apps.Addresses;
 
 import java.io.ByteArrayInputStream;
 import java.io.ObjectInputStream;
@@ -29,10 +29,10 @@ public class Peer implements Runnable{
      *
      * @param neem MultiCast object
      */
-    Peer(MulticastChannel neem){
+    public Peer(MulticastChannel neem, App app){
         this.neem = neem;
         this.oracle = new StabilityOracle();
-        this.orderingComponent = new OrderingComponent(oracle);
+        this.orderingComponent = new OrderingComponent(oracle, app);
         this.uuid = neem.getProtocolMBean().getLocalId();
         this.disseminationComponent = new DisseminationComponent(new Random(), neem.getNet(), oracle, this, neem, orderingComponent);
         neem.getProtocolMBean().setGossipFanout(DisseminationComponent.K);
@@ -52,41 +52,12 @@ public class Peer implements Runnable{
                 ByteBuffer bb = ByteBuffer.wrap(buf);
 
                 neem.read(bb);
-                ByteArrayInputStream byteIn = new ByteArrayInputStream(buf); //TODO Check correctness
+                ByteArrayInputStream byteIn = new ByteArrayInputStream(bb.array());
                 ObjectInputStream in = new ObjectInputStream(byteIn);
                 disseminationComponent.receive((HashMap<UUID, Event>) in.readObject());
             }
         } catch (AsynchronousCloseException ace) {
             // Exiting.
-        } catch (Exception e) {
-            e.printStackTrace();
-        }
-    }
-
-    public static void main(String[] args) {
-        if (args.length < 1) {
-            System.err.println("Usage: apps.Chat local peer1 ... peerN");
-            System.exit(1);
-        }
-
-        try {
-
-            MulticastChannel neem = new MulticastChannel(Addresses.parse(args[0], true));
-
-            System.out.println("Started: " + neem.getLocalSocketAddress());
-
-            if (neem.getLocalSocketAddress().getAddress().isLoopbackAddress())
-                System.out.println("WARNING: Hostname resolves to loopback address! Please fix network configuration\nor expect only local peers to connect.");
-
-            //Todo should use start ?
-            Peer peer = new Peer(neem);
-            new Thread(peer).start();
-
-            for (int i = 1; i < args.length; i++)
[…]

### 2016-04-26 src/src/epto/utilities/App.java

@@ -1,12 +1,19 @@
 package epto.utilities;
 
 
+import epto.Peer;
+import net.sf.neem.MulticastChannel;
+import net.sf.neem.apps.Addresses;
 import net.sf.neem.impl.Application;
 
-    import java.nio.ByteBuffer;
+import java.io.ByteArrayInputStream;
+import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.nio.ByteBuffer;
 
 /**
  * Implementation of an Application
+ * TODO broadcast + create a start application
  */
 public class App implements Application{
 
@@ -15,6 +22,49 @@ public class App implements Application{
      */
     @Override
     public void deliver(ByteBuffer[] byteBuffers) {
+        for (ByteBuffer byteBuffer : byteBuffers) {
+            byte[] content = byteBuffer.array();
+            ByteArrayInputStream byteIn = new ByteArrayInputStream(content);
+            try {
+                ObjectInputStream in = new ObjectInputStream(byteIn);
+                Event event = (Event) in.readObject();
+                System.out.println(event.toString());
+            } catch (IOException e) {
+                e.printStackTrace();
+            } catch (ClassNotFoundException e) {
+                e.printStackTrace();
+            }
+        }
+    }
 
+
+    public static void main(String[] args) {
+        if (args.length < 1) {
+            System.err.println("Usage: apps.Chat local peer1 ... peerN");
+            System.exit(1);
+        }
+
+        try {
+
+            MulticastChannel neem = new MulticastChannel(Addresses.parse(args[0], true));
+
+            System.out.println("Started: " + neem.getLocalSocketAddress());
+
+            if (neem.getLocalSocketAddress().getAddress().isLoopbackAddress())
+                System.out.println("WARNING: Hostname resolves to loopback address! Please fix network configuration\nor expect only local peers to connect.");
+
+            App app = new App();
+            //Todo should use start ?
+            Peer peer = new Peer(neem, app);
+            new Thread(peer).start();
+
+            for (int i = 1; i < args.length; i++)
+                neem.connect(Addresses.parse(args[i], false));
+
+            neem.close();
+
+        } catch (Exception e) {
+            e.printStackTrace();
+        }
     }
 }

### 2016-04-26 src/src/epto/utilities/App.java
Add starting mecanism

@@ -13,9 +13,17 @@ import java.nio.ByteBuffer;
 
 /**
  * Implementation of an Application
- * TODO broadcast + create a start application
+ * TODO broadcast ?
  */
 public class App implements Application{
+
+    private Peer peer;
+    private MulticastChannel neem;
+
+    public App(MulticastChannel neem) {
+        this.neem = neem;
+        this.peer = new Peer(neem, this);
+    }
 
     /**
      * {@inheritDoc}
@@ -37,6 +45,10 @@ public class App implements Application{
         }
     }
 
+    private void start() {
+        new Thread(peer).start();
+    }
+
 
     public static void main(String[] args) {
         if (args.length < 1) {
@@ -53,13 +65,14 @@ public class App implements Application{
             if (neem.getLocalSocketAddress().getAddress().isLoopbackAddress())
                 System.out.println("WARNING: Hostname resolves to loopback address! Please fix network configuration\nor expect only local peers to connect.");
 
-            App app = new App();
+
             //Todo should use start ?
-            Peer peer = new Peer(neem, app);
-            new Thread(peer).start();
+            App app = new App(neem);
 
             for (int i = 1; i < args.length; i++)
                 neem.connect(Addresses.parse(args[i], false));
+
+            app.start();
 
             neem.close();
 

### 2016-04-26 report/paper.tex
rewrite the intro

@@
 \begin{abstract}
-One of the fundamental problems of distributed computing, is the ordering of events through all peers.
-Among different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications.
-unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package.
-EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events.
-EpTO uses a probabilistic agreement to disseminate events through the system with high probability, and also provides integrity, total order and validity.
-We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems.
-Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups.
+
 \end{abstract}
-
 \section{Introduction}
-The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over the past few decades.
-Many researches have been done to create algorithms with different properties and trade-offs such as synchronization, agreement or state machine replication \autocites[]{defago2004total}[]{lamport1978time}.
-Unfortunately, due the fact that these properties are strong guarantees, the algorithms that implement them do not scale as expected because of churn and failures in an non-ideal environment.
-Because of the scalability problem inherent to deterministic approaches, probabilistic protocols have emerged.
-Nowadays, existing probabilistic protocols are highly scalable and resilient, but they do not provide total ordering.
-\todo[inline]{Maybe add a paper to justify this claim}
+Ordering of events is one the fundamental problems of distributed systems.
+Creating an algorithm which provide scalability, integrity and validity, along with an ordering for the event through all the peers and processes is one the hot topics in distributed systems research areas for many years.
+One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
+EpTO is an algorithm which is claimed to provide integrity, validity and total order throughout a distributed system.
+In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.
 \par 
-The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system on the order of messages.
+There are many other algorithm for dissemination and ordering events in a distributed systems.
+There are some deterministic algorithms, which guarantee total order, agreement or other strong properties deterministically.
+But this type of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
+\par
+The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system.
 This causes a massive amount of network traffic and overhead on the system.
 Moreover, an agreement feature for an asynchronous system requires to
-explicitly maintain a group and have access to a failure detector.
+explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
 Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits the scalability of the algorithm.
+\par
+As an alternative for deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
+These algorithms guarantee to disseminate events through the system with high probability.
+In this way, there is not need for failure detectors and redundant traffic thus, these algorithms can be quit scalable.
+Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as ordering.
+\par
+This is where EpTO is coming into the light. EpTO with mixing these two types of algorithms, provide total order along with scalability, validity and integrity.
+
 \subsection{Contribution}
-EpTO \autocite{matos2015epto} guarantees that all processes eventually agree on the set of received events with high probability and deliver these in total order to the application.
-The intuition behind EpTO is to make events available in all processes quickly with high probability.
-Once events are thought to be available in every peer, each peer deterministically orders them using the timestamp of each event and breaking ties with the id of the broadcaster peer, delivers them to the application accordingly.
+EpTO consists of two parts.
+First part is probabilistic dissemination.
+EpTO guarantees that all peers will receive an event with high probability.
+Second part of EpTO is deterministic ordering.
+Once peers receive all events, they will deterministically order them using the events timestamp, and in case of a tie using event's broadcaster's id.
 \par
-The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
-The \textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
-\par
-Using the balls-and-bins approach we model processes as bins and events (or set of events) as balls.
-Based on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability.
-Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.
-
+For the first part, EpTO is using \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls.
+In this scenario, one the natural questions that comes to mind is that: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?
 \begin{figure}
@@
 \par
-In this paper, we will implement EpTO and compare it against a known deterministic total order algorithm, namely JGroups \autocite{jgroups}.
+Using the balls-and-bins approach we model processes as bins and events as balls.
+Based on this model we will calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability.
+Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
+Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.
 
-These tasks will be split in two tasks.
-The first task will assume that all peers are known by all peers.
+\par
+Until now, the creators of EpTO test this algorithm in a simulated environment, and not in real-world conditions.
+We are going implement EpTO in pure JAVA using NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
+\par
+These tasks will be split in two parts.
+The first part will assume that all peers are known by all peers.
 We will implement EpTO using this assumption and test it against JGroups.
-The second task will assume unknown membership on all peers.
+I the second part of the tasks, will assume unknown membership on all peers.
 We will update EpTO and use a PSS\footnote{Peer Sampling Service}.
-We will also update the underlying layer NeEM \autocite{neem} to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
+We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
 

### 2016-04-26 report/paper.tex
rewrite abstract

@@
 \begin{abstract}
-
+One of the fundamental problems of distributed systems is that a message between two processes, may be delivered zero, or many times.
+This is way ordering events and messages is an essence in distrusted systems.
+There has been many algorithm to overcome this problem.
+EpTO is one of the recently introduced total order algorithms for large-scale distributed systems.
+EpTO is the first algorithm that claim to provide total order and scalability at the same time.
+We are going to verify this claim by implementing EpTO and evaluate it in a real-world condition.
 \end{abstract}
@@
 These algorithms guarantee to disseminate events through the system with high probability.
-In this way, there is not need for failure detectors and redundant traffic thus, these algorithms can be quit scalable.
+In this way, there is not need for failure detectors and redundant traffic.
+Thus, these algorithms can be quit scalable.
 Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as ordering.
 \par
-This is where EpTO is coming into the light. EpTO with mixing these two types of algorithms, provide total order along with scalability, validity and integrity.
-
-\subsection{Contribution}
+This is where EpTO is coming into the light.
+EpTO with mixing these two types of algorithms, provide total order along with scalability, validity and integrity.
 EpTO consists of two parts.
@@
 
+\subsection{Contribution}
+Until now, the creators of EpTO test this algorithm in a simulated environment, and not in a real-world scenario.
+We are going implement EpTO in pure JAVA using NeEM library \autocite{neem}.
+Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
+This comparison help us to verify if EpTO is actually perform as expected in real-world conditions or not.
 \par
-Until now, the creators of EpTO test this algorithm in a simulated environment, and not in real-world conditions.
-We are going implement EpTO in pure JAVA using NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
-Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
-\par
-These tasks will be split in two parts.
+These tasks will be split into two parts.
 The first part will assume that all peers are known by all peers.
@@
 On the other hand, partial ordering only defines the order between certain key events that depend on each other.
-Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example in order to have an exact image of the state of the system at all time, we have to use total order.
+Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example when we need an exact image of the state of the system at all time, we have to use total order.
 

### 2016-04-26 report/paper.tex
Fix some grammar errors

@@
 \section{Introduction}
-Ordering of events is one the fundamental problems of distributed systems.
-Creating an algorithm which provide scalability, integrity and validity, along with an ordering for the event through all the peers and processes is one the hot topics in distributed systems research areas for many years.
+The ordering of events is one the most prominent problems in distributed systems \todo[inline]{cite paper}.
+Creating an algorithm which provides scalability, integrity and validity, along with an ordering for the events through all peers and processes has been one the hot topics in distributed systems research for many years.
 One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
-EpTO is an algorithm which is claimed to provide integrity, validity and total order throughout a distributed system.
+EpTO is an algorithm which claims to provide integrity, validity and total order in a large scale distributed system.
 In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.
 \par 
-There are many other algorithm for dissemination and ordering events in a distributed systems.
-There are some deterministic algorithms, which guarantee total order, agreement or other strong properties deterministically.
-But this type of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
+There are many other algorithms for disseminating and ordering events in a distributed system.
+There are some deterministic algorithms, which guarantee total order, agreement or other strong properties.
+Unfortunately, these types of algorithm are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
 \par
@@
 As an alternative for deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
-These algorithms guarantee to disseminate events through the system with high probability.
-In this way, there is not need for failure detectors and redundant traffic.
-Thus, these algorithms can be quit scalable.
-Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as ordering.
+These algorithms guarantee the dissemination of events in the system with high probability.
+This way, there is no need for failure detectors and redundant traffic thus, these algorithms are highly scalable.
+Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as total ordering.
 \par
-This is where EpTO is coming into the light.
-EpTO with mixing these two types of algorithms, provide total order along with scalability, validity and integrity.
+This is where EpTO comes into light. EpTO, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
 EpTO consists of two parts.
-First part is probabilistic dissemination.
-EpTO guarantees that all peers will receive an event with high probability.
-Second part of EpTO is deterministic ordering.
-Once peers receive all events, they will deterministically order them using the events timestamp, and in case of a tie using event's broadcaster's id.
+The first part is probabilistic dissemination.
+EpTO guarantees that all peers will receive an event with an arbitrarily high probability.
+The second part of EpTO is deterministic ordering.
+Once peers have received all events, they will deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.
 \par
-For the first part, EpTO is using \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+For the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
 The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls.
@@
 Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.
+Using these approaches, EpTO becomes highly scalable and resilient, while still providing total order.
 
 \subsection{Contribution}
-Until now, the creators of EpTO test this algorithm in a simulated environment, and not in a real-world scenario.
-We are going implement EpTO in pure JAVA using NeEM library \autocite{neem}.
+Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
+We are going to implement EpTO in pure JAVA using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
 Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
-This comparison help us to verify if EpTO is actually perform as expected in real-world conditions or not.
+This comparison helps us verify if EpTO is actually performing as expected in a real-world scenario.
 \par
-These tasks will be split into two parts.
+These tasks will be split in two parts.
 The first part will assume that all peers are known by all peers.
 We will implement EpTO using this assumption and test it against JGroups.
-I the second part of the tasks, will assume unknown membership on all peers.
+In the second part of the tasks, we will assume unknown membership on all peers.
 We will update EpTO and use a PSS\footnote{Peer Sampling Service}.

### 2016-04-26 report/paper.tex
Rewrite some parts

@@
 \def\ehsan{Ehsan Farhadi}
-\author{\IEEEauthorblockN{\jocelyn\IEEEauthorrefmark{1},
-\ehsan\IEEEauthorrefmark{2}}
+\author{
+	\IEEEauthorblockN{\jocelyn}
+	\IEEEauthorblockA{Université de Fribourg\\
+	Fribourg, Switzerland\\
+	Email:\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch}
+	}
+	\and
+	\IEEEauthorblockN{\ehsan}
 	\IEEEauthorblockA{Université de Neuchâtel\\
 		Neuchâtel, Switzerland\\
-Email: \IEEEauthorrefmark{1}\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch},
-\IEEEauthorrefmark{2}\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}}}
+		Email:\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}
+	}
+
+}
 
@@
 \begin{abstract}
-One of the fundamental problems of distributed systems is that a message between two processes, may be delivered zero, or many times.
-This is way ordering events and messages is an essence in distrusted systems.
-There has been many algorithm to overcome this problem.
-EpTO is one of the recently introduced total order algorithms for large-scale distributed systems.
-EpTO is the first algorithm that claim to provide total order and scalability at the same time.
-We are going to verify this claim by implementing EpTO and evaluate it in a real-world condition.
+Total ordering has always been of interest in large-scale distributed systems.
+EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
+In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions.
+We then compare it to a deterministic total order algorithm named JGroups.
 \end{abstract}
 \section{Introduction}
-The ordering of events is one the most prominent problems in distributed systems \todo[inline]{cite paper}.
-Creating an algorithm which provides scalability, integrity and validity, along with an ordering for the events through all peers and processes has been one the hot topics in distributed systems research for many years.
+Creating an algorithm which provides scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
 One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
-EpTO is an algorithm which claims to provide integrity, validity and total order in a large scale distributed system.
+EpTO is an algorithm which claims to provide integrity, validity and total order in a large-scale distributed system.
 In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.
@@
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits the scalability of the algorithm.
+Due to faults and churn in large-scale distributed systems, failure detector turns into a bottleneck for the system and thus, limits the scalability of the algorithm.
 \par
@@
 \section{Ordering Algorithms}
-\todo[inline]{We will explain why Ordering algorithms are necessary, the difference between partial order and Total order and why partial order might not be sufficient in some cases}
-
 Distributed systems, like centralized systems need to preserve the temporal order of event produced by concurrent process in the system.
@@
 \end{itemize}
-\todo[inline]{Explain advantages and usefulness of both systems}
 \par
@@
 On the other hand, partial ordering only defines the order between certain key events that depend on each other.
-Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example when we need an exact image of the state of the system at all time, we have to use total order.
+Partial order can be useful since it is less costly to implement.
+However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order, we have to use total order, otherwise we could end up in an inconsistent state.
 
@@
 
-EpTO is scalable in the sense that \todo[inline]{Explain the logarithmic scalability}
+We define scalable as it was defined in  \autocite{matos2015epto}:  ``The number
+of messages transmitted per process per round is logarithmic
+in the number of processes, ...''
 

### 2016-04-27 report/paper.tex
Fix errors

@@
 EpTO is an algorithm which claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, EpTO is designed to work without a global clock or synchronization process, that is why EpTO is well-suited for dynamic large-scale distributed systems.
+In addition, EpTO is designed to work without a global clock, that is why EpTO is well-suited for dynamic large-scale distributed systems.
 \par 
@@
 \par
-As an alternative for deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
+As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
 These algorithms guarantee the dissemination of events in the system with high probability.
-This way, there is no need for failure detectors and redundant traffic thus, these algorithms are highly scalable.
-Although because these type of algorithms focus on reliability of dissemination, they often ignore other needed properties such as total ordering.
+This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable.
+Although as these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
 \par
 This is where EpTO comes into light. EpTO, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
-EpTO consists of two parts.
+EpTO consists of two distinct parts.
 The first part is probabilistic dissemination.
-EpTO guarantees that all peers will receive an event with an arbitrarily high probability.
+EpTO guarantees that all peers will receive an event with arbitrarily high probability.
 The second part of EpTO is deterministic ordering.
@@
 \par
-For the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+To model the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
 The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls.
-In this scenario, one the natural questions that comes to mind is that: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?
+In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?
 \begin{figure}
@@
 Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-We are going to implement EpTO in pure JAVA using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+We are going to implement EpTO in pure Java using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
 Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
-This comparison helps us verify if EpTO is actually performing as expected in a real-world scenario.
+These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
 \par
@@
 In the second part of the tasks, we will assume unknown membership on all peers.
-We will update EpTO and use a PSS\footnote{Peer Sampling Service}.
-We will also update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
+We will update EpTO and use a PSS\footnote{Peer Sampling Service} to find peers to exchange with.
+We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
 
 \section{Ordering Algorithms}
-Distributed systems, like centralized systems need to preserve the temporal order of event produced by concurrent process in the system.
+Distributed systems, like centralized systems need to preserve the temporal order of events produced by concurrent processes in the system.
 When there are separated processes that can only communicate through messages, you can’t easily order these messages.

### 2016-04-27 report/paper.tex
Hugues comments

@@
 \section{Introduction}
-Creating an algorithm which provides scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
+Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
 One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
 EpTO is an algorithm which claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, EpTO is designed to work without a global clock, that is why EpTO is well-suited for dynamic large-scale distributed systems.
+In addition, EpTO is designed to work without a global clock, thus is well-suited for dynamic large-scale distributed systems.
 \par 
@@
 \par
-The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system.
+The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the system.
 This causes a massive amount of network traffic and overhead on the system.
@@
 Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-We are going to implement EpTO in pure Java using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
-Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
+In this work, We are going to implement EpTO in pure Java using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
 These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
+\todo[inline]{Here we present our conclusion briefly}
 \par
-These tasks will be split in two parts.
+These tasks are split in two parts.
 The first part will assume that all peers are known by all peers.
@@
 We will update EpTO and use a PSS\footnote{Peer Sampling Service} to find peers to exchange with.
-We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2\footnote{Non-blocking I/O available from Java 7}.
+We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2 \todo{NIO2?} \autocite{nio}.
 
 \section{Ordering Algorithms}
-Distributed systems, like centralized systems need to preserve the temporal order of events produced by concurrent processes in the system.
+Distributed systems, like centralized systems, need to preserve the temporal order of events produced by concurrent processes in the system.
 When there are separated processes that can only communicate through messages, you can’t easily order these messages.
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{figures/epto-architecture.png}
-	\caption{\todo[inline]{We need a better quality figure (SVG)}\protect\footnotemark{}.}
+	\includegraphics[width=\linewidth]{figures/architecture.pdf}
+	\caption{EpTO architecture \autocite{matos2015epto}}
 	\label{fig:epto-architecture}
@@
 The Dissemination component is the component that bridges EpTO with the rest of the network.
-As we can see in Figure \ref*{fig:epto-architecture}, it is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the gossip fan-out.
+As we can see in  \todo{Hugues comment} Figure \ref*{fig:epto-architecture}, it is in charge of receiving balls it opens them, passes them to the Ordering component and then forwards them to \textit{K} other processes, where K denotes the gossip fan-out.
 \par
@@
 \par
-As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
+As this component is local and only corrects itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
 \section{Performance Comparisons}

### 2016-04-27 src/src/epto/OrderingComponent.java

@@ -53,7 +53,7 @@ public class OrderingComponent {
             Event event = received.get(key);
             if(!delivered.containsKey(key) && event.getTimeStamp() > lastDeliveredTs){
                 if (received.containsKey(key)){
-                    if(received.get(key).getTtl() < event.getTtl()){ //TODO check later
+                    if(received.get(key).getTtl() < event.getTtl()){
                         received.get(key).setTtl(event.getTtl());
                     }
                 }
@@ -82,7 +82,7 @@ public class OrderingComponent {
             Event event = received.get(key);
             if (event.getTimeStamp() > minQueuedTs) {
                 // ignore deliverable events with timestamp greater than all non-deliverable events
-                deliverableEvents.remove(key); //TODO check later
+                deliverableEvents.remove(key);
             }
             else {
                 // event can be delivered, remove from received events

### 2016-04-27 report/paper.tex
Finish corrections

@@
 These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
-\todo[inline]{Here we present our conclusion briefly}
+\todo[inline]{Here we will present our conclusion briefly}
 \par
@@
 We will update EpTO and use a PSS\footnote{Peer Sampling Service} to find peers to exchange with.
-We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2 \todo{NIO2?} \autocite{nio}.
+We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2 \autocite{nio}.
 
@@
 The Dissemination component is the component that bridges EpTO with the rest of the network.
-As we can see in  \todo{Hugues comment} Figure \ref*{fig:epto-architecture}, it is in charge of receiving balls it opens them, passes them to the Ordering component and then forwards them to \textit{K} other processes, where K denotes the gossip fan-out.
+As we can see in Figure \ref*{fig:epto-architecture}, it receives balls, it opens them, it passes them to the Ordering component and then it forwards them to \textit{K} other processes, where K denotes the gossip fan-out.
 \par

### 2016-04-27 src/src/epto/utilities/App.java
Fix usage in App

@@ -15,7 +15,7 @@ import java.nio.ByteBuffer;
  * Implementation of an Application
  * TODO broadcast ?
  */
-public class App implements Application{
+public class App implements Application {
 
     private Peer peer;
     private MulticastChannel neem;
@@ -52,7 +52,7 @@ public class App implements Application{
 
     public static void main(String[] args) {
         if (args.length < 1) {
-            System.err.println("Usage: apps.Chat local peer1 ... peerN");
+            System.err.println("Usage: apps.App local peer1 ... peerN");
             System.exit(1);
         }
 
@@ -66,7 +66,6 @@ public class App implements Application{
                 System.out.println("WARNING: Hostname resolves to loopback address! Please fix network configuration\nor expect only local peers to connect.");
 
 
-            //Todo should use start ?
             App app = new App(neem);
 
             for (int i = 1; i < args.length; i++)

### 2016-04-27 src/src/epto/utilities/App.java
add a broadcast method to app for broadcasting events

@@ -6,10 +6,10 @@ import net.sf.neem.MulticastChannel;
 import net.sf.neem.apps.Addresses;
 import net.sf.neem.impl.Application;
 
-import java.io.ByteArrayInputStream;
-import java.io.IOException;
-import java.io.ObjectInputStream;
+import java.io.*;
 import java.nio.ByteBuffer;
+import java.util.HashMap;
+import java.util.UUID;
 
 /**
  * Implementation of an Application
@@ -50,6 +50,30 @@ public class App implements Application {
     }
 
 
+    //TODO broadcasts events every 1 second
+    private void broadcast() {
+        int i = 0;
+        while(true){
+            Event event = new Event(UUID.randomUUID(),i++,5,UUID.randomUUID());
+            HashMap<UUID, Event> output_hash = new HashMap<>();
+            output_hash.put(UUID.randomUUID(), event);
+            ByteArrayOutputStream byteOut = new ByteArrayOutputStream();
+            try {
+                ObjectOutputStream out = new ObjectOutputStream(byteOut);
+                out.writeObject(output_hash);
+                this.neem.write(ByteBuffer.wrap(byteOut.toByteArray()));
+            } catch (IOException e) {
+                e.printStackTrace();
+            }
+            try {
+                Thread.sleep(1000);
+            } catch (InterruptedException e) {
+                e.printStackTrace();
+            }
+        }
+    }
+
+
     public static void main(String[] args) {
         if (args.length < 1) {
             System.err.println("Usage: apps.App local peer1 ... peerN");
@@ -72,11 +96,12 @@ public class App implements Application {
                 neem.connect(Addresses.parse(args[i], false));
 
             app.start();
-
+            app.broadcast();
             neem.close();
 
         } catch (Exception e) {
             e.printStackTrace();
         }
     }
+
 }

### 2016-04-27 src/src/epto/utilities/Event.java

@@ -1,11 +1,12 @@
 package epto.utilities;
 
+import java.io.Serializable;
 import java.util.UUID;
 
 /**
  * Implementation of the events. This class implements the structure of an event as described in EpTO.
  */
-public class Event implements Comparable<Event> {
+    public class Event implements Comparable<Event>, Serializable {
     private UUID id;
     private long timeStamp;
     private int ttl;

### 2016-06-06 report/paper.tex
Add reference for BnB figure
Correct mistakes in report

@@
 \usepackage{csquotes}
-\usepackage[backend=biber,style=ieee,minbibnames=1,maxbibnames=3]{biblatex}
+\usepackage[backend=biber,style=ieee,minbibnames=1,maxbibnames=99]{biblatex}
+\usepackage{xpatch}
 \usepackage{url}
@@
 	Fribourg, Switzerland\\
-	Email:\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch}
+	\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch}
 	}
@@
 		Neuchâtel, Switzerland\\
-		Email:\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}
+		\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}
 	}
@@
 
+\xpatchbibdriver{online}
+{\printtext[parens]{\usebibmacro{date}}}
+{\iffieldundef{year}
+	{}
+	{\printtext[parens]{\usebibmacro{date}}}}
+{}
+{\typeout{There was an error patching biblatex-ieee (specifically, ieee.bbx's @online driver)}}
+
 \begin{document}
@@
 EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions.
+In this paper, we verify this claim by implementing EpTO\autocite{matos2015epto} and evaluate its reliability in real-world conditions.
 We then compare it to a deterministic total order algorithm named JGroups.
@@
 Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
-One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
-EpTO is an algorithm which claims to provide integrity, validity and total order in a large-scale distributed system.
+One of the recently designed algorithms on this topic is EpTO.
+EpTO is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
 In addition, EpTO is designed to work without a global clock, thus is well-suited for dynamic large-scale distributed systems.
@@
 There are some deterministic algorithms, which guarantee total order, agreement or other strong properties.
-Unfortunately, these types of algorithm are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
+Unfortunately, these types of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
 \par
@@
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, failure detector turns into a bottleneck for the system and thus, limits the scalability of the algorithm.
+Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the scalability of the algorithm.
 \par
@@
 This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable.
-Although as these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
+As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
 \par
@@
 To model the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
-The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin completely random and independent from other balls.
-In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball?
+The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls.
+In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?
 \begin{figure}
 \includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins}
+\caption{Balls-and-Bins \autocite{bnb}}
 \label{fig:balls-and-bins}
@@
 Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-In this work, We are going to implement EpTO in pure Java using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+In this work, we implement EpTO in pure Kotlin using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
 We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
@@
 A process or peer is defined as an actor in our system running the application that needs total order.
-Each process will communicate with other processes in the distributed systems and exchange events and order them together.
+Each process will communicate with other processes in the distributed systems, exchange events, and order them together.
 \par
-An event is defined as something that happens at a given time.
-For example, we could imagine a system where each process publish some sort of data.
+An event is defined as data sent at a given time by a peer.
+For example, we could imagine a system where each process publish some data to other peers.
 The moment where this data is published combined with the data is called an event.
@@
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''
+in the number of processes, ...''. The number of rounds to deliver must be low as well.
 
@@
 Since EpTO is using a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
-In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered event will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
+In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
 \subsection{Dissemination Component}
@@
 The Ordering component is responsible for ordering events before delivering them to the application.
-To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which are delivered.
+To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which were delivered.
 \par

### 2016-06-06 report/paper.tex
Fix sentences and grammar based on peer-review feedback

@@
 EpTO is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, EpTO is designed to work without a global clock, thus is well-suited for dynamic large-scale distributed systems.
+In addition, EpTO is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.
 \par 
@@
 \includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins \autocite{bnb}}
+\caption{Balls-and-Bins \autocite{bnb} \todo[inline] {FIX IMAGE}}
 \label{fig:balls-and-bins}
@@
 Based on this model we will calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability.
-Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Using these approaches, EpTO becomes highly scalable and resilient, while still providing total order.
+With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
+Thanks to these approaches, EpTO becomes highly scalable and resilient, while still providing total order.
 
@@
 We will implement EpTO using this assumption and test it against JGroups.
-In the second part of the tasks, we will assume unknown membership on all peers.
+In the second part of the tasks, we will assume unknown membership on all peers, meaning, peers never have a full view  of all the peers.
 We will update EpTO and use a PSS\footnote{Peer Sampling Service} to find peers to exchange with.
@@
 An event is defined as data sent at a given time by a peer.
-For example, we could imagine a system where each process publish some data to other peers.
-The moment where this data is published combined with the data is called an event.
+For example, we could imagine a system where each process publishes some data to other peers.
+The moment where we publish data combined with the data is called an event.
 \par

### 2016-06-06 report/paper.tex
Write Future Works and Methodology

@@
 \section{Performance Comparisons}
-\todo[inline]{Here we will present the different test we ran, the methodology used and the result we obtained}
+In the following subsections, we will present the results obtained for the different tests ran and explain their meanings.
 \subsection{Methodology}
-\todo[inline]{Here we explain how we ran our tests and on what type of machines}
+Our tests are run on a cluster provided by the University of Neuchâtel.
+We use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
+We make use of Docker \autocite{docker}.
+More specifically we use Docker-compose, which is a tool to run multi-container applications, coupled to docker-swarm, which lets us make a pool of Docker hosts appear as a single Docker host.
+\par
+Each peer is hosted in a Docker container.
+We want to test JGroups and EpTO and find at which point JGroups breaks latency and throughput wise.
+At this point EpTO should still be working.
+Under this treshold, JGroups should be faster than EpTO but EpTO should still hold its own.
 \subsection{Peer membership known}
 \todo[inline]{We will present the results obtained and try to explain the results}
-\subsection{Peer membership unknown}
-\todo[inline]{We will present the results obtained and try to explain the results}
 
+\section{Future Works}
+Future works will include testing, EpTO while having unknown peer membership.
+To account for that, we will need to modify NeEM to support the use of a Peer Sampling Service.
+We plan on implementing the CYCLON Peer Sampling Service \autocite{Voulgaris2005}.
+
+Our first implementation of EpTO focuses on correctness rather than fastness.
+Later work should optimize EpTO in regards to concurrency and network latency.
+To address the second point, we want to modify NeEM again to use UDP instead of TCP, to reduce latency and overhead on the network.
 \section{Conclusion}

### 2016-06-06 report/paper.tex
Rewrite a part of the contributions subsection

@@
 Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-In this work, we implement EpTO in pure Kotlin using the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+In this work, we implement EpTO in pure Kotlin using a modified version of the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+The modification to NeEM includes a dumbed-down gossip dissemination as seen in \autocite{matos2015epto}.
 We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
@@
 \par
-These tasks are split in two parts.
-The first part will assume that all peers are known by all peers.
-We will implement EpTO using this assumption and test it against JGroups.
-In the second part of the tasks, we will assume unknown membership on all peers, meaning, peers never have a full view  of all the peers.
-We will update EpTO and use a PSS\footnote{Peer Sampling Service} to find peers to exchange with.
-We will then update the underlying layer NeEM to use UDP instead of TCP and switch to NIO.2 \autocite{nio}.
+In our experiments we assume each peer knows every other peers.
+We implement EpTO using this assumption and test it against JGroups.
+The details concerning the methodology are explained in subsection \ref{sub:metho}.
+In future works, we want to test EpTO when not knowing the entire membership.
 
@@
 \subsection{Methodology}
+\label{sub:metho}
 Our tests are run on a cluster provided by the University of Neuchâtel.

### 2016-06-11 scripts/container-start-script.sh
Try to make tracker work

@@ -17,4 +17,4 @@ MY_NAME=(${TMP//./ })
 sleep 2m
 
 echo 'Starting epto peer'
-exec java -Xms50m -Xmx100m -cp ../build/libs/epto-1.0-SNAPSHOT-all.jar epto.utilities.App $MY_NAME "http://localhost:4321" | addtime > localhost.txt 2>&1
+exec java -Xms50m -Xmx100m -cp ../build/libs/epto-1.0-SNAPSHOT-all.jar epto.utilities.App $MY_NAME "http://eptoneem_tracker_1:4321" | addtime > localhost.txt 2>&1

### 2016-06-11 scripts/tracker.py

@@ -3,13 +3,14 @@ __author__ = "Sebastien Vaucher"
 '''
 
 import subprocess
-import threading
+import pydevd as pydevd
 from random import shuffle
 from http.server import HTTPServer, BaseHTTPRequestHandler
 
 
 def florida_string():
-    ps_output = subprocess.check_output('docker-compose ps epto', shell=True).decode().splitlines()
+    # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
+    ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
 
     nodes_names = [x.split(' ')[0] for x in ps_output if x.startswith('eptoneem')]
     shuffle(nodes_names)

### 2016-06-11 scripts/tracker.py
Try to expose some ports

@@ -3,21 +3,19 @@ __author__ = "Sebastien Vaucher"
 '''
 
 import subprocess
-import pydevd as pydevd
 from random import shuffle
 from http.server import HTTPServer, BaseHTTPRequestHandler
 
 
 def florida_string():
     # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
+    # TODO take IP from context
     ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
 
     nodes_names = [x.split(' ')[0] for x in ps_output if x.startswith('eptoneem')]
     shuffle(nodes_names)
     # TODO change 10 to k later
     return '|'.join(nodes_names[:10]).encode()
-
-
 
 
 class FloridaHandler(BaseHTTPRequestHandler):
@@ -34,10 +32,10 @@ class FloridaHandler(BaseHTTPRequestHandler):
             self.wfile.write(b"Nothing here, content is at /REST/v1/admin/get_view\n")
 
 
-class florida_server:
+class FloridaServer:
     def __init__(self):
         self.server = HTTPServer(('', 4321), FloridaHandler)
         self.server.serve_forever()
 
 
-florida_server()
+FloridaServer()

### 2016-06-11 report/paper.tex
chaning report title, add acknowledgement and some info to methodolgy

@@
 % can use linebreaks \\ within to get better formatting as desired
-\def\mytitle{EpTO: An Epidemic Total Order Algorithm for Large-Scale Distributed Systems}
+\def\mytitle{Practical Implementation of EpTO: An Epidemic Total Order Algorithm}
 \title{\mytitle}
@@
 Our tests are run on a cluster provided by the University of Neuchâtel.
-We use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
-We make use of Docker \autocite{docker}.
-More specifically we use Docker-compose, which is a tool to run multi-container applications, coupled to docker-swarm, which lets us make a pool of Docker hosts appear as a single Docker host.
+For setting up our testbed we use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
+We make use of Docker \autocite{docker}, More specifically we use Docker-compose.
+Docker-compose is a tool to run multi-container applications, coupled to docker-swarm, which lets us make a pool of Docker hosts appear as a single Docker host.
 \par
-Each peer is hosted in a Docker container.
-We want to test JGroups and EpTO and find at which point JGroups breaks latency and throughput wise.
+Each peer is hosted in a Docker container in order to be isolated from other peers.
+We want to test and compare JGroups and EpTO and find at which point JGroups breaks in terms of latency and throughput.
 At this point EpTO should still be working.
 Under this treshold, JGroups should be faster than EpTO but EpTO should still hold its own.
+\par
+To measure the throughput, we transmit a certain number of events by certain number of peers in a defined period of time, and then we measure number of packets transmitted over the network.
+\par
+To measure the latency of both protocols, we measure the average time difference between broadcasting an event by a peer, and delivering the same event by the same peers.
+In this way we can have an estimation of the approximate delay, and compare two protocols.
 \subsection{Peer membership known}
@@
 \section*{Acknowledgment}
-\todo[inline]{Write acknowledgment}
+This research was under supervision and help of one of the creators of EpTO algorithm, Dr.
+Hugues Mercier.
+We 	thank Dr.
+Miguel Matos, another creator of EpTO and one the creators of NeEM library, who provided us insight and expertise that greatly assisted us during the research.
+\par`
+We also like to thank Valerio Schiavoni to provide us the access to the cluster of University of Neuchâtel.
 

### 2016-06-12 scripts/tracker.py
another way to access docker daemon from the container

@@ -10,7 +10,11 @@ from http.server import HTTPServer, BaseHTTPRequestHandler
 def florida_string():
     # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
     # TODO take IP from context
-    ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
+    
+    #in this way docker-compose won't work in the container so instead of the command you wrote originaly, you should just do:
+    #ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
+    ps_output = subprocess.check_output('docker ps', shell=true).decode().splitlines()
+    #and then find the containers with eptoneem_epto_x name
 
     nodes_names = [x.split(' ')[0] for x in ps_output if x.startswith('eptoneem')]
     shuffle(nodes_names)

### 2016-06-12 scripts/container-start-script.sh
Make Dockerfile from Debian

@@ -17,4 +17,4 @@ MY_NAME=(${TMP//./ })
 sleep 2m
 
 echo 'Starting epto peer'
-exec java -Xms50m -Xmx100m -cp ../build/libs/epto-1.0-SNAPSHOT-all.jar epto.utilities.App $MY_NAME "http://eptoneem_tracker_1:4321" | addtime > localhost.txt 2>&1
+exec java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App $MY_NAME "http://eptoneem_tracker_1:4321" | addtime > localhost.txt 2>&1

### 2016-06-12 scripts/tracker.py

@@ -3,6 +3,7 @@ __author__ = "Sebastien Vaucher"
 '''
 
 import subprocess
+import pydevd
 from random import shuffle
 from http.server import HTTPServer, BaseHTTPRequestHandler
 
@@ -10,11 +11,7 @@ from http.server import HTTPServer, BaseHTTPRequestHandler
 def florida_string():
     # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
     # TODO take IP from context
-    
-    #in this way docker-compose won't work in the container so instead of the command you wrote originaly, you should just do:
-    #ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
-    ps_output = subprocess.check_output('docker ps', shell=true).decode().splitlines()
-    #and then find the containers with eptoneem_epto_x name
+    ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
 
     nodes_names = [x.split(' ')[0] for x in ps_output if x.startswith('eptoneem')]
     shuffle(nodes_names)

### 2016-06-12 scripts/container-start-script.sh
Fix scripts to be more efficient

@@ -7,14 +7,12 @@ addtime() {
     done
 }
 
-cd /code/scripts
-
 MY_IP_ADDR=$(/bin/hostname -i)
 TMP=$(dig -x $MY_IP_ADDR +short)
 MY_NAME=(${TMP//./ })
 
 # wait for all peers
-sleep 2m
+sleep 30s
 
 echo 'Starting epto peer'
 exec java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App $MY_NAME "http://eptoneem_tracker_1:4321" | addtime > localhost.txt 2>&1

### 2016-06-12 scripts/epto-benchmark.sh

@@ -1,33 +1,20 @@
 #!/bin/bash
 # for running the test and gathering the results
-# 
+#
 
-containsElement() {
-  local e
-  for e in "${@:2}"; do [[ "$e" == "$1" ]] && return 0; done
-  return 1
-}
+PEER_NUMBER=$1
 
 check_if_finish() {
-	mapfile -t lines < <(docker ps -a)
-	lines=("${lines[@]:1}")
-	number_of_nodes=${#lines[@]}
-	number_of_running_nodes=0
-	for i in "${lines[@]}"
-	do
-		read -a line <<<$i
-		if containsElement "Up" "${line[@]}"; then
-		   ((number_of_running_nodes++))
-		fi
-	done
-	finished_nodes=$(expr $number_of_nodes - $number_of_running_nodes)
-	if (( $finished_nodes > 3 )); then 
-		return 0
-	fi
-	return 1
+    if [ $(docker ps -aqf "status=exited" | wc -l) == "$PEER_NUMBER" ]
+        then
+            return 0
+        else
+            return 1
+    fi
+
 }
 
-if [ -z "$1" ]
+if [ -z "$PEER_NUMBER" ]
   then
     echo "you have to indicate number of peers"
     exit
@@ -37,17 +24,20 @@ fi
 echo "rebuild the image"
 docker-compose build
 echo "clean up previous containers"
-docker rm -f $(docker ps -a -q)
+if [ -n "$(docker ps -a -q)" ]
+    then
+        docker rm -f $(docker ps -a -q)
+fi
 echo "START..."
 COMPOSE_HTTP_TIMEOUT=200
 export COMPOSE_HTTP_TIMEOUT
 docker-compose up -d
-docker-compose scale epto=$1
+docker-compose scale epto=${PEER_NUMBER}
 
 #wait for apps to finish
-for i in 1 2 :
+for i in {1..20} :
 do
-	sleep 1m
+	sleep 10s
 	if check_if_finish;then
 		break
 	else
@@ -58,16 +48,7 @@ done
 #sleep 2m
 
 # collect logs
-mapfile -t lines < <(docker ps -a)
-lines=("${lines[@]:1}")
-for i in "${lines[@]}"
-do
-	read -a line <<<$i
-	LENGTH=${#line[@]}
-	LAST_POSITION=$((LENGTH - 1))
-
[…]

### 2016-06-12 scripts/tracker.py

@@ -11,7 +11,7 @@ from http.server import HTTPServer, BaseHTTPRequestHandler
 def florida_string():
     # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
     # TODO take IP from context
-    ps_output = subprocess.check_output('docker-compose -H tcp://192.168.1.201:2375 ps epto', shell=True).decode().splitlines()
+    ps_output = subprocess.check_output('docker-compose ps epto', shell=True).decode().splitlines()
 
     nodes_names = [x.split(' ')[0] for x in ps_output if x.startswith('eptoneem')]
     shuffle(nodes_names)

### 2016-06-12 report/paper.tex
Work on the paper

@@
 \includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins \autocite{bnb} \todo[inline] {FIX IMAGE}}
+\caption{Balls-and-Bins \autocite{bnb}}
 \label{fig:balls-and-bins}
@@
 For setting up our testbed we use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
-We make use of Docker \autocite{docker}, More specifically we use Docker-compose.
+We make use of Docker \autocite{docker}, more specifically we use Docker-compose.
 Docker-compose is a tool to run multi-container applications, coupled to docker-swarm, which lets us make a pool of Docker hosts appear as a single Docker host.
@@
 \section{Future Works}
-Future works will include testing, EpTO while having unknown peer membership.
+One of the student working on this project will do his master thesis on this project.
+
+Future works will include testing EpTO while having unknown peer membership.
 To account for that, we will need to modify NeEM to support the use of a Peer Sampling Service.
 We plan on implementing the CYCLON Peer Sampling Service \autocite{Voulgaris2005}.
+For now, we plan on using a tracker as in torrents to give each peer its initial view.
+This tracker is already implemented and works as a python web server that gets all alive peers and randomly select $K$ peers and sends this result to the EpTO peer.
 
@@
 \par`
-We also like to thank Valerio Schiavoni to provide us the access to the cluster of University of Neuchâtel.
+We would also like to thank Valerio Schiavoni to provide us the access to the cluster of University of Neuchâtel as well as Sebastien Vaucher for the help with setting up Docker and Docker Swarm.
 

### 2016-06-12 scripts/tracker.py
Use information gained from the tracker

@@ -3,7 +3,7 @@ __author__ = "Sebastien Vaucher"
 '''
 
 import subprocess
-import pydevd
+#import pydevd
 from random import shuffle
 from http.server import HTTPServer, BaseHTTPRequestHandler
 

### 2016-06-15 report/paper.tex
Fix typos and rewrite some sentences

@@
 EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO\autocite{matos2015epto} and evaluate its reliability in real-world conditions.
+In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions.
 We then compare it to a deterministic total order algorithm named JGroups.
@@
 Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
-One of the recently designed algorithms on this topic is EpTO.
+One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
 EpTO is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
@@
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the scalability of the algorithm.
+Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability.
 \par
@@
 The second part of EpTO is deterministic ordering.
-Once peers have received all events, they will deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.
+Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.
 \par
@@
 \includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins \autocite{bnb}}
+\caption{Balls-and-Bins \autocite{bnb}.}
 \label{fig:balls-and-bins}
@@
 \par
-Using the balls-and-bins approach we model processes as bins and events as balls.
-Based on this model we will calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability.
+Using the balls-and-bins approach we model processes as bins and events as balls and calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability.
 With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
@@
 Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-In this work, we implement EpTO in pure Kotlin using a modified version of the NeEM library \autocite{neem} and show that EpTO is actually suitable for real-world large-scale distributed systems.
+In this work, we implement EpTO in pure Kotlin using a modified version of the NeEM library \autocite{neem} and show that EpTO is suitable for real-world large-scale distributed systems.
 The modification to NeEM includes a dumbed-down gossip dissemination as seen in \autocite{matos2015epto}.
@@
 The details concerning the methodology are explained in subsection \ref{sub:metho}.
-In future works, we want to test EpTO when not knowing the entire membership.
+In future work, we want to test EpTO when the membership is not entirely known.
 
@@
 Partial order can be useful since it is less costly to implement.
-However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order, we have to use total order, otherwise we could end up in an inconsistent state.
+However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order.
+We then have to use total order, otherwise we can end up in an inconsistent state.
 
@@
 	\includegraphics[width=\linewidth]{figures/architecture.pdf}
-	\caption{EpTO architecture \autocite{matos2015epto}}
+	\caption{EpTO architecture \autocite{matos2015epto}.}
 	\label{fig:epto-architecture}
@@
 A process or peer is defined as an actor in our system running the application that needs total order.
-Each process will communicate with other processes in the distributed systems, exchange events, and order them together.
+Each process will communicate with other processes in the distributed system, exchange events, and order them together.
 \par
@@
 
-We define scalable as it was defined in  \autocite{matos2015epto}:  ``The number
+We define EpTO scaling well as it was defined in  \autocite{matos2015epto}:  ``The number
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds to deliver must be low as well.
+in the number of processes, ...''. The number of rounds needed to deliver an event stay low as well.
+\par
+Since EpTO uses a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
+In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by EpTO's deterministic ordering algorithm, thus the total order property is preserved.
+\subsection{Dissemination Component}
+The Dissemination component bridges EpTO with the rest of the network.
+As we can see in Figure \ref*{fig:epto-architecture}, it receives balls, opens them, passes them to the Ordering component, and stores their events.
+Then, in the next round, it sends all the events received in the previous round to K other processes, where K denotes the gossip fan-out.
 
-\par
-Since EpTO is using a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
-In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
-\subsection{Dissemination Component}
-The Dissemination component is the component that bridges EpTO with the rest of the network.
-As we can see in Figure \ref*{fig:epto-architecture}, it receives balls, it opens them, it passes them to the Ordering component and then it forwards them to \textit{K} other processes, where K denotes the gossip fan-out.
 \par
@@
 This is done to prevent delivering duplicate events.
-The remaining events in the received ball will be added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
+The remaining events in the received ball are added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
 \subsection{Stability Oracle}
@@
 At this point EpTO should still be working.
-Under this treshold, JGroups should be faster than EpTO but EpTO should still hold its own.
+Under this threshold, JGroups should be faster than EpTO but EpTO should still hold its own.
 \par
-To measure the throughput, we transmit a certain number of events by certain number of peers in a defined period of time, and then we measure number of packets transmitted over the network.
+To measure the throughput, we transmit a certain number of events by certain number of peers in a defined period of time, and then we measure the number of packets transmitted over the network.
 \par
-To measure the latency of both protocols, we measure the average time difference between broadcasting an event by a peer, and delivering the same event by the same peers.
-In this way we can have an estimation of the approximate delay, and compare two protocols.
+To measure the latency of both protocols, we measure the average time difference on a peer between the broadcast of an event and the time it delivers this same event.
+This way we can have an estimation of the approximate delay, and compare JGroups and EpTO.
 \subsection{Peer membership known}
@@
 
-\section{Future Works}
-One of the student working on this project will do his master thesis on this project.
+\section{Future Work}
+One of the students working on this project will do his master thesis on this project.
 
-Future works will include testing EpTO while having unknown peer membership.
+Future work will include testing EpTO while having unknown peer membership.
 To account for that, we will need to modify NeEM to support the use of a Peer Sampling Service.
@@
 Later work should optimize EpTO in regards to concurrency and network latency.
-To address the second point, we want to modify NeEM again to use UDP instead of TCP, to reduce latency and overhead on the network.
+To address the second point, we want to modify NeEM again to use UDP instead of TCP, to reduce latency and overhead in the network.
 \section{Conclusion}
@@
 \section*{Acknowledgment}
-This research was under supervision and help of one of the creators of EpTO algorithm, Dr.
-Hugues Mercier.
-We 	thank Dr.
-Miguel Matos, another creator of EpTO and one the creators of NeEM library, who provided us insight and expertise that greatly assisted us during the research.
-\par`
-We would also like to thank Valerio Schiavoni to provide us the access to the cluster of University of Neuchâtel as well as Sebastien Vaucher for the help with setting up Docker and Docker Swarm.
+This research was under supervision and help of two of the creators of EpTO, Dr.
+Hugues Mercier and Dr.
+Miguel Matos.
+We would like to thank them both for the help, insight and expertise they provided during this project.
+\par
+We would also like to thank Valerio Schiavoni for providing access to the cluster of University of Neuchâtel as well as Sebastien Vaucher for the help with setting up Docker, Docker-Compose and Docker Swarm.
 

### 2016-06-15 scripts/epto-benchmark-nodocker.rb
Add ugly script to start without docker

@@ -1,0 +1,23 @@
+#!/usr/bin/env ruby
+
+
+ports = [*10000..10050]
+
+puts 'compiling...'
+puts `cd ..; ./gradlew --daemon clean shadowJar && cd scripts/;`
+# Replace each of the dummy IP by the ips of the vms (don't forget the ':')
+ips = %w(1.1.1.1: 1.1.1.2: 1.1.1.3: 1.1.1.4:)
+# Here indicate the ip of the machine on which this script is started
+own_ip = '1.1.1.1:'
+
+args = ports.map { |port| [ips.map { |ip| ip + port.to_s }] }
+args.flatten!
+
+(0..50).each { |i|
+  complete_ip = own_ip + ports[i].to_s
+  args.delete(complete_ip)
+  args.insert(0, complete_ip)
+  args = args.join(' ')
+
+  puts `java -cp ../build/libs/epto-1.0-SNAPSHOT-all.jar epto.utilities.App #{args} > localhost#{i}.txt 2>&1 &`
+}

### 2016-06-16 scripts/nodocker.sh
adding shell script and minimizing the jvm

@@ -1,0 +1,62 @@
+#!/bin/bash	
+
+if [ -z "$1" ]
+  then
+    echo "No argument supplied"
+    exit
+fi
+
+addtime() {
+    while IFS= read -r line; do
+        echo "$(date +%s%N | cut -b1-13) $line"
+    done
+}
+
+cd ..; ./gradlew --daemon clean shadowJar && cd scripts/;
+# wait for all peers
+sleep 30
+
+
+
+MY_IP_ADDR=$(ifconfig eth0 | grep "inet addr" | cut -d ':' -f 2 | cut -d ' ' -f 1)
+
+number_of_peers=$1
+last_port=$((10000+number_of_peers))
+ips=(172.16.0.42: 172.16.0.43: 172.16.0.48: 172.16.0.49:)
+own_ip="$MY_IP_ADDR:"
+ports=()
+for ((p=10000; p<last_port; p++));do 
+	ports+=($p)
+done
+args=()
+for f1 in {0..3}
+do
+	for ((f2=0; f2<number_of_peers; f2++));do
+		args+=(${ips[$f1]}${ports[$f2]})
+	done
+done
+
+for ((i=0; i<number_of_peers;i++));do	
+	t_args=()
+	complete_ip="$own_ip${ports[$i]}"
+	for t in "${args[@]}"
+	do
+		if [ "$t" != "$complete_ip" ]; then
+			t_args+=("$t")
+		fi
+	done
+	final_args=("$complete_ip")
+	for m in "${t_args[@]}"
+	do
+		final_args+=("$m")
+	done
+	final_args_string=""
+
+	for n in "${final_args[@]}"
+	do
+		final_args_string+="$n "
+	done
+	#echo $final_args_string
+	echo 'Starting epto peer'
+	exec java -Xms20m -Xmx50m -cp ../build/libs/epto-1.0-SNAPSHOT-all.jar epto.utilities.App $final_args_string | addtime > localhost$i.txt 2>&1 &
+done

### 2016-06-16 report/paper.tex
Fix missing 's'
report: test

@@
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver an event stay low as well.
+in the number of processes, ...''. The number of rounds needed to deliver an event stays low as well.
 \par
@@
 For setting up our testbed we use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
-We make use of Docker \autocite{docker}, more specifically we use Docker-compose.
-Docker-compose is a tool to run multi-container applications, coupled to docker-swarm, which lets us make a pool of Docker hosts appear as a single Docker host.
+We were planning on using Docker \autocite{docker}, more specifically Docker-compose and docker-swarm in order to orchestrate and manage our cluster of peers on the OpenNebula but due to some issues with docker-swarm, we decided to deploy our application directly on the OpenNebula virtual machines, and control it using parallel ssh sessions.
 \par
-Each peer is hosted in a Docker container in order to be isolated from other peers.
-We want to test and compare JGroups and EpTO and find at which point JGroups breaks in terms of latency and throughput.
-At this point EpTO should still be working.
-Under this threshold, JGroups should be faster than EpTO but EpTO should still hold its own.
+We deployed our application on a small network of 4 virtual machines and we distribute EpTO and JGroups peers equally over the network.
 \par
@@
 This way we can have an estimation of the approximate delay, and compare JGroups and EpTO.
+We execute EpTO and JGroups with 60, 100, 120 and 160 peers each peer broadcasting 12 events for 12 seconds (one event per second).
 \subsection{Peer membership known}
-\todo[inline]{We will present the results obtained and try to explain the results}
-
+As we measured and compared the latency of EpTO and JGroups we realize that JGroups is performing better than EpTO with smaller number of peers.
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/Delay.png}
+\caption{Approximate delay of delivering each event.}
+\label{fig:delay}
+\end{figure}
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/total_time.png}
+\caption{total run time of EpTO peers in order to deliver all events.}
+\label{fig:total_time}
+\end{figure}
+As we increased the number of peers, our results showed that JGroups is not as scalable as EpTO, but considering our limited resources we couldn't fully observe the difference between JGroup and EpTO on this matter.
+Although we should bear in mind that in this experiment EpTO has a relatively large delay due to ordering of the events without a central node (as in JGroups).
+But despite of this delay, we expected that EpTO performs much better and our analysis is that due the insufficient resource on the testbed.
+Epto could not show its advantages.
+\par
+We perceive this issue again when we measured the total time of execution of EpTO for delivering a certain number of events.
+\par
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/throughput.png}
+\caption{number of events transmitted packets over the network in 10 minutes.}
+\label{fig:throughput}
+\end{figure}
+For the throughput of both protocols we used a network sniffing tool, TCPdump.
+TCPdump is a tool for observing the network traffic.
+Using this tool we count number of packets (TCP packet for EpTO and UDP packet for JGroups) transmitting over the network for 10 minutes.
+And again, number of packets sent by EpTO peers are much higher than JGroups.
+This issue can be explain due the differences of the using protocols, TCP and UDP.
+TCP uses much more packets and data over the network for a connection to secure and guarantee the delivery of data.
+that's why in a perfect and lossless network, JGroups has better performance from perspective of network throughput.
 \section{Future Work}

### 2016-06-16 report/paper.tex
Write the conclusion

@@
 The modification to NeEM includes a dumbed-down gossip dissemination as seen in \autocite{matos2015epto}.
-We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
+We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in stable (no-churn) systems.
 These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
-\todo[inline]{Here we will present our conclusion briefly}
 \par
@@
 \end{figure}
-As we increased the number of peers, our results showed that JGroups is not as scalable as EpTO, but considering our limited resources we couldn't fully observe the difference between JGroup and EpTO on this matter.
+When we increase the number of peers, our results show that JGroups is not as scalable as EpTO, but considering our limited resources we could not fully observe the difference between JGroup and EpTO on this matter.
 Although we should bear in mind that in this experiment EpTO has a relatively large delay due to ordering of the events without a central node (as in JGroups).
-But despite of this delay, we expected that EpTO performs much better and our analysis is that due the insufficient resource on the testbed.
-Epto could not show its advantages.
+But despite this delay, we expected EpTO to perform much better.
+We infer the problem might be related more to the way the experiments were run and not directly to EpTO.
+This needs further testing and will be done during the master thesis on this subject.
 \par
-We perceive this issue again when we measured the total time of execution of EpTO for delivering a certain number of events.
+This issue is also present when measuring the total time of execution of EpTO for delivering a certain number of events.
 \par
@@
 TCPdump is a tool for observing the network traffic.
-Using this tool we count number of packets (TCP packet for EpTO and UDP packet for JGroups) transmitting over the network for 10 minutes.
-And again, number of packets sent by EpTO peers are much higher than JGroups.
-This issue can be explain due the differences of the using protocols, TCP and UDP.
+Using this tool we count the number of packets (TCP packets for EpTO and UDP packets for JGroups) transmitted over the network for 10 minutes.
+Again, the number of packets sent by EpTO peers are much higher than JGroups.
+This issue can be explain due to the differences between TCP and UDP.
 TCP uses much more packets and data over the network for a connection to secure and guarantee the delivery of data.
 that's why in a perfect and lossless network, JGroups has better performance from perspective of network throughput.
+We expect the number of packets sent by EpTO to drastically diminish once we have made the switch to UDP.
 \section{Future Work}
@@
 \section{Conclusion}
-\todo[inline]{He we will summarize the results we found and present some future tasks that could be accomplished on EpTO}
-
-% conference papers do not normally have an appendix
+In conclusion, EpTO was implemented using Kotlin with a focus on correctness instead of fastness.
+There are still many optimizations left to be done on the code which will be implemented during the master thesis following this project.
+\par
+Tests were ran evaluating  the average delay for an event to be delivered, the total time of execution and the number of packets transmitted over the network.
+These tests showed that EpTO was not behaving as we expected.
+Unfortunately as we encountered many problems during the tests we could not test the algorithm the way we wanted and these results are only a start and will be more conclusive in the future.
+\par
+At the moment, the tests must be run somewhat manually. We want to use Docker and Docker-swarm in the future to enhance the reproducibility of the tests on any cluster.
 

### 2016-06-16 scripts/eprun.sh
adding test results

@@ -1,0 +1,9 @@
+#!/bin/bash
+# run this script to log the packets sending to specific address of epto nodes
+# and the node loobback (for epto peers running on the same node)
+# also it uses ./mem.sh to log the memory status
+
+tcpdump -i eth0 tcp and "(dst host 172.16.0.43 or dst host 172.16.0.48 or dst host 172.16.0.49)" > ep$1-eth0.txt &
+tcpdump -i lo tcp > ep$1-lo.txt &
+./mem.sh &
+

### 2016-06-16 scripts/mem.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#using to log memory status of the machin.
+
+addtime() {
+    while IFS= read -r line; do
+        echo "$(date +%s%N | cut -b1-13) $line"
+    done
+}
+
+while true; do
+	mapfile -t lines < <(free -m)
+	lines=("${lines[@]:1}")
+	echo "${lines[0]}" | addtime >> $(hostname).txt 2>&1
+	sleep 1
+done

### 2016-06-16 scripts/nodocker.sh

@@ -1,8 +1,13 @@
-#!/bin/bash	
+#!/bin/bash
+
+# this script will run the epto peers and deploying them on specific machins with known IP adresses. you have to provide IP address of all working node: ips
+# to run this just run the ./nodocker.sh N 
+# in epto-neem/scripts/ folder
+# N is the number of peers per node.
 
 if [ -z "$1" ]
   then
-    echo "No argument supplied"
+    echo "indicate number of peers per node"
     exit
 fi
 

### 2016-06-16 scripts/stop-mon.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+#stop all the logging services
+
+pkill -f 'tcpdump'
+pkill -f 'mem.sh'

### 2016-06-16 test_results/jg_100/delivered_events_checked.py

@@ -1,0 +1,27 @@
+import os
+import re
+
+number_1 = []
+with open("node1_localhost1.txt") as f:
+    for line in f:
+        l = line.split()
+        if len(l) <= 2:
+            continue
+        if l[1] == 'Delivered:':
+            number_1.append(l[-1])
+
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    tmp = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                l = line.split()
+                if len(l) <= 2:
+                    continue
+                if l[1] == 'Delivered:':
+                    tmp.append(l[-1])
+        if tmp != number_1:
+            print("error")
+            break;

### 2016-06-16 test_results/jg_120/total_time.py

@@ -1,0 +1,19 @@
+import os
+import re
+
+overall_times = []
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    ff = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                ff .append(line.split())
+        overall_times.append(int(ff[-1][0]) - int(ff[0][0]))
+
+average_time = 0
+for i in overall_times:
+    average_time += i
+average_time /= len(overall_times)
+print(average_time)

### 2016-06-16 test_results/jg_60/jgroup-evaluation.py

@@ -1,0 +1,31 @@
+import os
+import re
+
+overall_delay = []
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    sent = []
+    delivered = []
+    event_diffs = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                l = line.split()
+                if len(l) == 1:
+                    continue
+                if l[1] == 'Sent:':
+                    sent.append([l[0], l[-1]])
+                if l[1] == 'Delivered:':
+                    delivered.append([l[0], l[-1]])
+
+        for j in sent:
+            for k in delivered:
+                if j[1] == k[1]:
+                    event_diffs.append(int(k[0]) - int(j[0]))
+        if len(event_diffs) > 1:
+            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
+            overall_delay.append(avg_event_diff)
+
+total_average = sum(overall_delay) / float(len(overall_delay))
+print(total_average)

### 2016-06-16 test_results/jg_60/jgroup-packet-counter.py

@@ -1,0 +1,18 @@
+import os
+import re
+
+count = 0
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    fileReGex = re.compile("node\d_jg\d*-lo\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                count += 1
+    fileReGex = re.compile("node\d_jg\d*-eth0\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                count += 1
+
+print(count)

### 2016-06-16 test_results/jg_60/ram_mon.py

@@ -1,0 +1,19 @@
+import os
+import re
+
+node = []
+
+for i in range(1, 5):
+    n = []
+    with open("node" + str(1) + "_node" + str(1) + ".txt") as f:
+        for line in f:
+            l = line.split()
+            n.append(l[3])
+    node.append(n)
+
+avg = []
+for i in range(0, len(node[1])):
+    avg.append((int(node[0][i]) + int(node[1][i]) + int(node[2][i]) + int(node[3][i]))/4)
+
+for i in avg:
+    print(i)

### 2016-06-16 test_results/scripts/checking_delivered_events.py

@@ -1,0 +1,30 @@
+#!/usr/bin/python
+#running this script inside the folder of logs will check if the delivered events on all peers are the same. If there's an error, it will print error.
+
+import os
+import re
+
+number_1 = []
+with open("node1_localhost1.txt") as f:
+    for line in f:
+        l = line.split()
+        if len(l) <= 2:
+            continue
+        if l[1] == 'Delivered:':
+            number_1.append(l[-1])
+
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    tmp = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                l = line.split()
+                if len(l) <= 2:
+                    continue
+                if l[1] == 'Delivered:':
+                    tmp.append(l[-1])
+        if tmp != number_1:
+            print("error")
+            break;

### 2016-06-16 test_results/scripts/epto-average-latency.py

@@ -1,0 +1,33 @@
+#!/usr/bin/python
+#running this script inside the folder of Epto logs will calculate the average latency.
+import os
+import re
+
+overall_delay = []
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    sent = []
+    delivered = []
+    event_diffs = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                l = line.split()
+                if len(l) <= 2:
+                    continue
+                if l[1] == 'sent:':
+                    sent.append([l[0], l[-1]])
+                if l[1] == 'Delivered:':
+                    delivered.append([l[0], l[-1]])
+
+        for j in sent:
+            for k in delivered:
+                if j[1] == k[1]:
+                    event_diffs.append(int(k[0]) - int(j[0]))
+        if len(event_diffs) > 1:
+            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
+            overall_delay.append(avg_event_diff)
+
+total_average = sum(overall_delay) / float(len(overall_delay))
+print(total_average)

### 2016-06-16 test_results/scripts/jgroup-average-latency.py

@@ -1,0 +1,34 @@
+#!/usr/bin/python
+#running this script inside the folder of JGroup logs will calculate the average latency.
+
+import os
+import re
+
+overall_delay = []
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    sent = []
+    delivered = []
+    event_diffs = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                l = line.split()
+                if len(l) == 1:
+                    continue
+                if l[1] == 'Sent:':
+                    sent.append([l[0], l[-1]])
+                if l[1] == 'Delivered:':
+                    delivered.append([l[0], l[-1]])
+
+        for j in sent:
+            for k in delivered:
+                if j[1] == k[1]:
+                    event_diffs.append(int(k[0]) - int(j[0]))
+        if len(event_diffs) > 1:
+            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
+            overall_delay.append(avg_event_diff)
+
+total_average = sum(overall_delay) / float(len(overall_delay))
+print(total_average)

### 2016-06-16 test_results/scripts/packet-counter.py

@@ -1,0 +1,21 @@
+#!/usr/bin/python
+#running this script inside the folder of logs will calculate the number of packets transmitted over the network (based on logs from TCPdump).
+
+import os
+import re
+
+count = 0 ()
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    fileReGex = re.compile("node\d_ep\d*-lo\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                count += 1
+    fileReGex = re.compile("node\d_ep\d*-eth0\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                count += 1
+
+print(count)

### 2016-06-16 test_results/scripts/ram_mon.py

@@ -1,0 +1,22 @@
+#!/usr/bin/python
+#running this script inside the folder of logs will show the memory load.
+
+import os
+import re
+
+node = []
+
+for i in range(1, 5):
+    n = []
+    with open("node" + str(1) + "_node" + str(1) + ".txt") as f:
+        for line in f:
+            l = line.split()
+            n.append(l[3])
+    node.append(n)
+
+avg = []
+for i in range(0, len(node[1])):
+    avg.append((int(node[0][i]) + int(node[1][i]) + int(node[2][i]) + int(node[3][i]))/4)
+
+for i in avg:
+    print(i)

### 2016-06-16 test_results/scripts/total_time.py

@@ -1,0 +1,22 @@
+#!/usr/bin/python
+#running this script inside the folder of logs will calculate the average total time for executing the Peers.
+
+import os
+import re
+
+overall_times = []
+files = [f for f in os.listdir('.') if os.path.isfile(f)]
+for file in files:
+    ff = []
+    fileReGex = re.compile("node\d_localhost\d*\.txt")
+    if fileReGex.match(file):
+        with open(file) as f:
+            for line in f:
+                ff .append(line.split())
+        overall_times.append(int(ff[-1][0]) - int(ff[10][0]))
+
+average_time = 0
+for i in overall_times:
+    average_time += i
+average_time /= len(overall_times)
+print(average_time)

### 2016-06-16 test_results/jg_100/jgroup-evaluation.py
deleteing test results

@@ -1,31 +1,0 @@
-import os
-import re
-
-overall_delay = []
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    sent = []
-    delivered = []
-    event_diffs = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                l = line.split()
-                if len(l) == 1:
-                    continue
-                if l[1] == 'Sent:':
-                    sent.append([l[0], l[-1]])
-                if l[1] == 'Delivered:':
-                    delivered.append([l[0], l[-1]])
-
-        for j in sent:
-            for k in delivered:
-                if j[1] == k[1]:
-                    event_diffs.append(int(k[0]) - int(j[0]))
-        if len(event_diffs) > 1:
-            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
-            overall_delay.append(avg_event_diff)
-
-total_average = sum(overall_delay) / float(len(overall_delay))
-print(total_average)

### 2016-06-16 test_results/jg_100/ram_mon.py

@@ -1,19 +1,0 @@
-import os
-import re
-
-node = []
-
-for i in range(1, 5):
-    n = []
-    with open("node" + str(1) + "_node" + str(1) + ".txt") as f:
-        for line in f:
-            l = line.split()
-            n.append(l[3])
-    node.append(n)
-
-avg = []
-for i in range(0, len(node[1])):
-    avg.append((int(node[0][i]) + int(node[1][i]) + int(node[2][i]) + int(node[3][i]))/4)
-
-for i in avg:
-    print(i)

### 2016-06-16 test_results/jg_100/total_time.py

@@ -1,19 +1,0 @@
-import os
-import re
-
-overall_times = []
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    ff = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                ff .append(line.split())
-        overall_times.append(int(ff[-1][0]) - int(ff[0][0]))
-
-average_time = 0
-for i in overall_times:
-    average_time += i
-average_time /= len(overall_times)
-print(average_time)

### 2016-06-16 test_results/jg_120/packet-counter.py

@@ -1,18 +1,0 @@
-import os
-import re
-
-count = 0
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    fileReGex = re.compile("node\d_jg\d*-lo\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                count += 1
-    fileReGex = re.compile("node\d_jg\d*-eth0\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                count += 1
-
-print(count)

### 2016-06-16 test_results/jg_60/delivered_events_checked.py

@@ -1,27 +1,0 @@
-import os
-import re
-
-number_1 = []
-with open("node1_localhost1.txt") as f:
-    for line in f:
-        l = line.split()
-        if len(l) <= 2:
-            continue
-        if l[1] == 'Delivered:':
-            number_1.append(l[-1])
-
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    tmp = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                l = line.split()
-                if len(l) <= 2:
-                    continue
-                if l[1] == 'Delivered:':
-                    tmp.append(l[-1])
-        if tmp != number_1:
-            print("error")
-            break;

### 2016-06-16 test_results/scripts/checking_delivered_events.py

@@ -1,30 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of logs will check if the delivered events on all peers are the same. If there's an error, it will print error.
-
-import os
-import re
-
-number_1 = []
-with open("node1_localhost1.txt") as f:
-    for line in f:
-        l = line.split()
-        if len(l) <= 2:
-            continue
-        if l[1] == 'Delivered:':
-            number_1.append(l[-1])
-
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    tmp = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                l = line.split()
-                if len(l) <= 2:
-                    continue
-                if l[1] == 'Delivered:':
-                    tmp.append(l[-1])
-        if tmp != number_1:
-            print("error")
-            break;

### 2016-06-16 test_results/scripts/epto-average-latency.py

@@ -1,33 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of Epto logs will calculate the average latency.
-import os
-import re
-
-overall_delay = []
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    sent = []
-    delivered = []
-    event_diffs = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                l = line.split()
-                if len(l) <= 2:
-                    continue
-                if l[1] == 'sent:':
-                    sent.append([l[0], l[-1]])
-                if l[1] == 'Delivered:':
-                    delivered.append([l[0], l[-1]])
-
-        for j in sent:
-            for k in delivered:
-                if j[1] == k[1]:
-                    event_diffs.append(int(k[0]) - int(j[0]))
-        if len(event_diffs) > 1:
-            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
-            overall_delay.append(avg_event_diff)
-
-total_average = sum(overall_delay) / float(len(overall_delay))
-print(total_average)

### 2016-06-16 test_results/scripts/jgroup-average-latency.py

@@ -1,34 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of JGroup logs will calculate the average latency.
-
-import os
-import re
-
-overall_delay = []
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    sent = []
-    delivered = []
-    event_diffs = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                l = line.split()
-                if len(l) == 1:
-                    continue
-                if l[1] == 'Sent:':
-                    sent.append([l[0], l[-1]])
-                if l[1] == 'Delivered:':
-                    delivered.append([l[0], l[-1]])
-
-        for j in sent:
-            for k in delivered:
-                if j[1] == k[1]:
-                    event_diffs.append(int(k[0]) - int(j[0]))
-        if len(event_diffs) > 1:
-            avg_event_diff = sum(event_diffs) / float(len(event_diffs))
-            overall_delay.append(avg_event_diff)
-
-total_average = sum(overall_delay) / float(len(overall_delay))
-print(total_average)

### 2016-06-16 test_results/scripts/packet-counter.py

@@ -1,21 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of logs will calculate the number of packets transmitted over the network (based on logs from TCPdump).
-
-import os
-import re
-
-count = 0 ()
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    fileReGex = re.compile("node\d_ep\d*-lo\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                count += 1
-    fileReGex = re.compile("node\d_ep\d*-eth0\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                count += 1
-
-print(count)

### 2016-06-16 test_results/scripts/ram_mon.py

@@ -1,22 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of logs will show the memory load.
-
-import os
-import re
-
-node = []
-
-for i in range(1, 5):
-    n = []
-    with open("node" + str(1) + "_node" + str(1) + ".txt") as f:
-        for line in f:
-            l = line.split()
-            n.append(l[3])
-    node.append(n)
-
-avg = []
-for i in range(0, len(node[1])):
-    avg.append((int(node[0][i]) + int(node[1][i]) + int(node[2][i]) + int(node[3][i]))/4)
-
-for i in avg:
-    print(i)

### 2016-06-16 test_results/scripts/total_time.py

@@ -1,22 +1,0 @@
-#!/usr/bin/python
-#running this script inside the folder of logs will calculate the average total time for executing the Peers.
-
-import os
-import re
-
-overall_times = []
-files = [f for f in os.listdir('.') if os.path.isfile(f)]
-for file in files:
-    ff = []
-    fileReGex = re.compile("node\d_localhost\d*\.txt")
-    if fileReGex.match(file):
-        with open(file) as f:
-            for line in f:
-                ff .append(line.split())
-        overall_times.append(int(ff[-1][0]) - int(ff[10][0]))
-
-average_time = 0
-for i in overall_times:
-    average_time += i
-average_time /= len(overall_times)
-print(average_time)

### 2016-08-01 scripts/container-start-script.sh
Modify script to have tracker in a container

@@ -15,5 +15,5 @@ sleep 30s
 echo 'Starting epto peer'
 echo "${MY_IP_ADDR}"
 MY_IP_ADDR=($MY_IP_ADDR)
-echo "${MY_IP_ADDR[1]}"
-java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[1]}" "http://192.168.1.201:4321" | addtime > "${MY_IP_ADDR[1]}.txt" 2>&1
+echo "${MY_IP_ADDR[0]}"
+java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://epto_tracker:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1

### 2016-08-01 scripts/epto-benchmark.sh

@@ -20,6 +20,7 @@ docker network create -d overlay epto-network
 TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
+docker service create --name epto_tracker --network epto-network --replicas 1 --limit-memory 180m tracker
 docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
 
 #wait for apps to finish

### 2016-08-01 scripts/container-start-script.sh
Fix address of the tracker

@@ -16,4 +16,4 @@ echo 'Starting epto peer'
 echo "${MY_IP_ADDR}"
 MY_IP_ADDR=($MY_IP_ADDR)
 echo "${MY_IP_ADDR[0]}"
-java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://epto_tracker:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1
+java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://${MY_IP_ADDR[1]}:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1

### 2016-08-01 scripts/container-start-script.sh
Use hostname provided by Docker for the tracker

@@ -16,4 +16,4 @@ echo 'Starting epto peer'
 echo "${MY_IP_ADDR}"
 MY_IP_ADDR=($MY_IP_ADDR)
 echo "${MY_IP_ADDR[0]}"
-java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://${MY_IP_ADDR[1]}:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1
+java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://epto-tracker:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1

### 2016-08-01 scripts/epto-benchmark.sh
Change '_' to '-'

@@ -20,7 +20,7 @@ docker network create -d overlay epto-network
 TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
-docker service create --name epto_tracker --network epto-network --replicas 1 --limit-memory 180m tracker
+docker service create --name epto-tracker --network epto-network --replicas 1 --limit-memory 180m tracker
 docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
 
 #wait for apps to finish

### 2016-08-01 scripts/epto-benchmark.sh
Add sleep between service creation to ensure tracker is already running

@@ -21,6 +21,7 @@ TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
 docker service create --name epto-tracker --network epto-network --replicas 1 --limit-memory 180m tracker
+sleep 15s
 docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
 
 #wait for apps to finish

### 2016-08-01 scripts/epto-benchmark.sh
Use ingress as network

@@ -16,13 +16,11 @@ fi
 
 echo "START..."
 docker swarm init
-docker network create -d overlay epto-network
 TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
-docker service create --name epto-tracker --network epto-network --replicas 1 --limit-memory 180m tracker
-sleep 15s
-docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
+docker service create --name epto-tracker --network ingress --replicas 1 --limit-memory 180m tracker
+docker service create --name epto-service --network ingress --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
 
 #wait for apps to finish
 for i in {1..40} :
@@ -34,6 +32,7 @@ done
 #sleep 2m
 
 docker service rm epto-service
+docker service rm epto-tracker
 # collect logs
 #for i in $(docker ps -aqf "ancestor=epto");do  docker cp ${i}:/opt/epto/localhost.txt ./${i}_log.txt; done
 parallel-ssh -h hosts "docker swarm leave"

### 2016-08-01 scripts/tracker.py
Remove Heartbeat checking for now
* Will have to be done in each peer

@@ -16,18 +16,17 @@ K = 16
 
 def florida_string(ip):
     available_peers[ip] = int(time.time())
-    for ip, timestamp in available_peers.items():
-        new_timestamp = int(time.time())
-        if new_timestamp - timestamp > 180:
-            to_keep = os.system("ping -c 1 -W 1 " + ip) == 0
+#    for ip, timestamp in available_peers.items():
+#        new_timestamp = int(time.time())
+#        if new_timestamp - timestamp > 180:
+#            to_keep = os.system("ping -c 1 -W 1 " + ip) == 0
 
-            if not to_keep:
-                available_peers.pop(ip)
-            else:
-                available_peers[ip] = new_timestamp
+#            if not to_keep:
+#                available_peers.pop(ip)
+#            else:
+#                available_peers[ip] = new_timestamp
 
-
-    # pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
+#   pydevd.settrace('192.168.1.201', port=9292, stdoutToServer=True, stderrToServer=True)
 
     if len(available_peers) > K:
         to_send = random.sample(available_peers.keys(), K)

### 2016-08-01 scripts/container-start-script.sh
Fix log path

@@ -16,4 +16,4 @@ echo 'Starting epto peer'
 echo "${MY_IP_ADDR}"
 MY_IP_ADDR=($MY_IP_ADDR)
 echo "${MY_IP_ADDR[0]}"
-java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://epto-tracker:4321" | addtime > "${MY_IP_ADDR[0]}.txt" 2>&1
+java -Xms50m -Xmx100m -cp ./epto-1.0-SNAPSHOT-all.jar epto.utilities.App "${MY_IP_ADDR[0]}" "http://epto-tracker:4321" | addtime > "/data/${MY_IP_ADDR[0]}.txt" 2>&1

### 2016-08-01 scripts/epto-benchmark.sh
Create own network instead of using ingress

@@ -19,8 +19,11 @@ docker swarm init
 TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
-docker service create --name epto-tracker --network ingress --replicas 1 --limit-memory 180m tracker
-docker service create --name epto-service --network ingress --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
+# If networking doesn't work use ingress
+docker network create -d overlay --subnet=10.0.9.0/24 epto-network
+
+docker service create --name epto-tracker --network epto-network --replicas 1 --limit-memory 180m tracker
+docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
 
 #wait for apps to finish
 for i in {1..40} :
@@ -33,6 +36,7 @@ done
 
 docker service rm epto-service
 docker service rm epto-tracker
+docker network rm epto-network
 # collect logs
 #for i in $(docker ps -aqf "ancestor=epto");do  docker cp ${i}:/opt/epto/localhost.txt ./${i}_log.txt; done
 parallel-ssh -h hosts "docker swarm leave"

### 2016-08-01 report/paper.tex

@@
 Tests were ran evaluating  the average delay for an event to be delivered, the total time of execution and the number of packets transmitted over the network.
-These tests showed that EpTO was not behaving as we expected.
-Unfortunately as we encountered many problems during the tests we could not test the algorithm the way we wanted and these results are only a start and will be more conclusive in the future.
+They show that EpTO was not behaving as we expected.
+Unfortunately, as we encountered many problems during the tests we could not test the algorithm the way we wanted and these results are only a start and will be more conclusive in the future.
+We feel the problems appearing in the tests can be fixed by having a better testing environment as well as a more optimized code.
 \par

### 2016-08-02 scripts/epto-benchmark.sh
Make adjustements to use script on a remote cluster

@@ -1,6 +1,5 @@
 #!/bin/bash
-# for running the test and gathering the results
-#
+# This scripts runs the benchmarks on a remote cluster
 
 MANAGER_IP=192.168.1.40
 PEER_NUMBER=$1
@@ -15,15 +14,20 @@ fi
 # TODO have a repo and pull from it for the image
 
 echo "START..."
+
+# Clean everything at Ctrl+C
+trap 'docker network rm epto-network && parallel-ssh -h hosts "docker swarm leave" \
+&& docker swarm leave --force && exit' TERM INT
+
 docker swarm init
 TOKEN=$(docker swarm join-token -q worker)
 parallel-ssh -h hosts "docker swarm join --token ${TOKEN} ${MANAGER_IP}:2377"
 
 # If networking doesn't work use ingress
-docker network create -d overlay --subnet=10.0.9.0/24 epto-network
+docker network create -d overlay --subnet=10.0.93.0/24 epto-network
 
 docker service create --name epto-tracker --network epto-network --replicas 1 --limit-memory 180m tracker
-docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 180m --mount type=bind,source=/data,target=/data epto
+docker service create --name epto-service --network epto-network --replicas ${PEER_NUMBER} --limit-memory 200m --mount type=bind,source=/home/debian/data,target=/data epto
 
 #wait for apps to finish
 for i in {1..40} :
@@ -38,10 +42,11 @@ docker service rm epto-service
 docker service rm epto-tracker
 docker network rm epto-network
 # collect logs
-#for i in $(docker ps -aqf "ancestor=epto");do  docker cp ${i}:/opt/epto/localhost.txt ./${i}_log.txt; done
 parallel-ssh -h hosts "docker swarm leave"
 docker swarm leave --force
 
 
-#analyze results
+for i (40 42); do
+    rsync -a -v "debian@192.168.1.${i}:~/data/" .
+done
 echo "finished"

### 2016-12-01 report/evaluation.tex
Add EpTO figures

\section{Evaluation}
\subsection{Bandwidth}
\jt{\epto was run with $c=4$ to prevent any holes and a $\delta$ period of 100ms}
\begin{figure}[h]
	\centering
	\input{figures/bandwidth-figure.tex}
	\vspace{-2mm} 
	\caption{Throughput percentiles of a node during an experiment}
	\vspace{-2mm} 
	\label{fig:bandwidth}
\end{figure}
In \autoref{fig:bandwidth} we can see \epto has a worse baseline compared to \jgroups. It uses a median bandwidth of approximately \SI{1}{\mbps} for the smallest benchmark whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}. However, in \jgroups most of the work is done solely by the coordinator. We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.5}{\mbps}.

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of events sent per second, we can see the bandwidth doubling in both cases. In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups has the upper hand. Since \epto has a worse baseline we will reach the maximum bandwidth possible much quicker when increasing the event throughput.

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto is performing better than \jgroups. Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it by 50\% or less. \jt{logarithmic I think as per the formula}. Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum bandwidth.

\begin{figure}
	\centering
	\input{figures/bandwidth-figure-churn.tex}
	\vspace{-2mm} 
	\caption{Throughput percentiles of a node during an experiment with churn}
	\vspace{-2mm} 
	\label{fig:bandwidth-churn}
\end{figure}

In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns. In the first one we kill one node per minute. In the second one, we still kill one node per minute, but we immediately create a new one. For \jgroups we ran the benchmarks once without killing the coordinator and once killing it.

All tests were run using 100 peers and a global throughput of 50 events per second (100-1sec). \jt{I need to rephrase 50-1sec, 100-1sec etc. to be clearer}

We can see that the churn doesn't affect  \epto at all when there are only nodes leaving. We have small peaks when adding a node to \SI{3}{\mbps} or less. Probably due to running the PSS initialization method on top of having one more node spreading rumors in the system. This is confirmed at the end of the plot where \epto goes back to a normal Bandwidth after stabilization.

On the other hand, when killing the coordinator in \jgroups we can see a huge spike in bandwidth, going from \SI{1}{\mbps} to more than \SI{15}{\mbps}. This is due to how \jgroups operates when selecting a new coordinator.

Even when not killing the coordinator, \jgroups suffers from the churn. We can see that each time the view changes, it generates an almost 100\% increase in bandwidth usage. This is due to \jgroups having to update the view and propagate it to every peer.

\newpage

\subsection{Total Bytes sent/received}
\begin{figure}[h]
	\centering
	\input{figures/total-bytes.tex}
	\vspace{-2mm} 
	\caption{Total bytes sent/received during an average experiment}
	\vspace{-2mm} 
	\label{fig:total-bandwidth}
\end{figure}
In \autoref{fig:total-bandwidth}, \epto has a worse baseline than \jgroups. This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at minimum $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect. Here we are well within this ratio.

\begin{figure}[h]
	\centering
	\input{figures/total-bytes-churn.tex}
	\vspace{-2mm} 
	\caption{Total bytes sent/received during an average experiment with churn}
	\vspace{-2mm} 
	\label{fig:total-bandwidth-churn}
\end{figure}
In \autoref{fig:total-bandwidth-churn} we can see that \jgroups total bandwidth usage doesn't increase significantly compared to \epto with churn. An explanation for this phenomenon is that \jgroups takes more than a minute to find out the coordinator is effectively dead. During this time the bandwidth is practically not used. The big spike when the new coordinator is chosen compensates for this hole thus making the total bandwidth used appear as to not have changed.

The fact that killing the coordinator or not is of no effect on the total bandwidth used tends to show the hypothesis to be correct.
\subsection{Local Times}
\label{sub:local-times}
\jt{Having a table for each percentile figure might be a good idea to put numbers for key percentiles (min,50th,max) for example}
\begin{figure}[h]
	\centering
	\input{figures/local-times.tex}
	\vspace{-2mm} 
	\caption{Local dissemination times}
	\vspace{-2mm}
	\label{fig:local-times} 
\end{figure}

\begin{figure}[h]
	\centering
	\input{figures/local-times-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination times with churn}
	\vspace{-2mm} 
	\label{fig:local-times-churn} 
\end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
\newpage
\subsection{Global Times}
\begin{figure}[h]
	\centering
	\input{figures/global-times.tex}
	\vspace{-2mm} 
	\caption{Global dissemination times}
	\vspace{-2mm}
	\label{fig:global-times}  
\end{figure}

\begin{figure}[h]
	\centering
	\input{figures/global-times-churn.tex}
	\vspace{-2mm} 
	\caption{Global dissemination times with churn}
	\vspace{-2mm} 
	\label{fig:global-times-churn} 
\end{figure}
We computed global times as well. They are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}. These global times are of less interest than their local counterpart as they differences in clocks between hosts can skew this measurement.

Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times}.
\newpage
\subsection{Local Dissemination stretch}
\begin{figure}[h]
	\centering
	\input{figures/local-delta.tex}
	\vspace{-2mm} 
	\caption{Local dissemination stretch}
	\vspace{-2mm}
	\label{fig:local-delta}  
\end{figure}
\jt{I wonder if we should split the plots (3 by 2) having a row for EpTO and a row for JGroups. This way we could show more for JGroups local dissemination stretch}In \autoref{fig:local-delta}, We can see the percentiles of the local dissemination stretch. The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.

\jgroups is usually much faster than \epto in a perfect environment. This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minmimal latency. The median dissemination stretch of \jgroups is around \SI{8}{\milli\second} where as the median dissemiantion stretch of \epto is around \SI{685}{\milli\second}. When increasing the number of peers, JGroups starts to have long delivery times

\begin{figure}[h]
	\centering
	\input{figures/local-delta-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination stretch with churn}
	\vspace{-2mm}
	\label{fig:local-delta-churn}   
\end{figure}
In \autoref{fig:local-delta-churn} We can see a completely different picture. When under churn, the 90th percentile of \jgroups is at \SI{447}{\milli\second} and the highest percentiles are at more than \SI{20}{\second} with the highest dissemination stretch being \SI{67.5}{\second}. This effect is due to the coordinator dying as we clearly see that it does not happen when we do not kill it.

The median is bigger at around \SI{11}{\milli\second}, whether we kill the coordinator or not. This shows that there are some degradation in \jgroups local dissemination stretch when under churn.

On the contrary, \epto performs very well under churn. The median stays really close at \SI{686}{\milli\second} with the 99.9th percentile being at \SI{1647}{\milli\second} compared to \SI{1366}{\milli\second} when no churn is happening.
\newpage
\subsection{Events sent}
\begin{figure}[h]
	\centering
	\input{figures/total-events-sent.tex}
	\vspace{-2mm} 
	\caption{Total events sent per experiment on average}
	\vspace{-2mm}
	\label{fig:total-events}   
\end{figure}
In \autoref{fig:total-events} we can see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
\begin{figure}[h]
	\centering
	\input{figures/total-events-sent-churn.tex}
	\vspace{-2mm} 
	\caption{Total events sent per experiment during churn on average}
	\vspace{-2mm}
	\label{fig:total-events-churn}  
\end{figure}

In \autoref{fig:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.

However when killing and adding nodes, we notice that \jgroups delivers more than 2000 less events than \epto, even when the coordinator is not killed.

We are not sure why this is happening, but one hypothesis is that \jgroups peers take a longer time to join the cluster and thus start sending events later. Since the benchmarks start and stop at a given time, they send less events in total.

### 2016-12-01 report/figures/bandwidth-figure-churn.tex
Add EpTO figures

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth / 2.6,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=2 by 3,
	vertical sep=8mm,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	yticklabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\second}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(1.25,1.5)},anchor=north, font=\small, draw=none},]
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-suspend-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-add-suspend-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},
yticklabels={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-suspend-coord-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-add-suspend-coord-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},
yticklabels={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-suspend-nocoord-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\end{groupplot}
\node[anchor=south] at (group c1r1.north) {1 kill};
\node[anchor=south] at (group c2r1.north) {1 kill/minute, 1 add/minute};
\node[anchor=south, rotate=-90] at (group c2r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c2r2.east){JGroups coord};
\node[anchor=south, rotate=-90] at (group c2r3.east){JGroups nocoord};
\end{tikzpicture}

### 2016-12-01 report/figures/bandwidth-figure.tex
Add EpTO figures

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth / 2.6,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=3 by 2,
	vertical sep=8mm,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	yticklabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
ymax=4.5,
ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\second}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(1.9,1.5)},anchor=north, font=\small, draw=none},]
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-50-1sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-50-2sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-100-1sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},yticklabels={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-50-1sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-50-2sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-100-1sec-0.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.250000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.500000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.750000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-1.000000, col sep=comma]{csv-data/average-bytes-sent-recv.csv} \closedcycle;
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\node[anchor=south, rotate=-90] at (group c3r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c3r2.east){JGroups};
\end{tikzpicture}

### 2016-12-01 report/figures/global-times-churn.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
group size=2 by 1,
vertical sep=0pt,
horizontal sep=1cm,
xlabels at=edge bottom,
ylabels at=edge left,
xticklabels at=edge bottom,
yticklabels at=edge left,
},
every axis plot/.append style={very thick},
height=5cm, width=\linewidth / 1.75,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
%
\nextgroupplot
\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}

### 2016-12-01 report/figures/global-times.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.34,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.8,1.4)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {csv-data/plot-global-time-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}

### 2016-12-01 report/figures/local-delta-churn.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
height=5cm, width=\linewidth / 1.75,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
scaled x ticks = false,
x tick label style={/pgf/number format/fixed},
xlabel={Time $\left[\si{\second}\right]$},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\legend{EpTO, JGroups-coord, JGroups-nocoord}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}

### 2016-12-01 report/figures/local-delta.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=3.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.7,1.4)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}

### 2016-12-01 report/figures/local-times-churn.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
group size=2 by 1,
vertical sep=0pt,
horizontal sep=6mm,
xlabels at=edge bottom,
ylabels at=edge left,
xticklabels at=edge bottom,
yticklabels at=edge left,
},
every axis plot/.append style={very thick},
height=5cm, width=\linewidth / 1.75,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
%
\nextgroupplot
\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}

### 2016-12-01 report/figures/local-times.tex
Add EpTO figures

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.7,1.4)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}

### 2016-12-01 report/figures/total-bytes-churn.tex
Add EpTO figures

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=0,
ymax=35,
width=\linewidth / 1.75,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
\nextgroupplot
% 100-1sec 1kill/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 21.3437665258) +- (0, 0.302297622101786)
	(JGroups-nocoord, 1.3338886813) +- (0, 0.00853400183682598)
	(JGroups-coord, 1.3440551501) +- (0, 0.0164341768113215)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 21.4504884001) +- (0, 0.0157182231373203)
	(JGroups-nocoord, 1.285075604) +- (0, 0.00872334968295833)
	(JGroups-coord, 1.2945710642) +- (0, 0.0158581561021704)
};
\legend{receive, send}
\nextgroupplot
% 100-1sec 1kill, 1add/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 27.0797668262) +- (0, 0.300652054138594)
	(JGroups-nocoord, 1.6071496365) +- (0, 0.00972747186106245)
	(JGroups-coord, 1.6039433255) +- (0, 0.0084962148465832)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 26.8856375946) +- (0, 0.0157182231373203)
	(JGroups-nocoord, 1.5459290852) +- (0, 0.00937891633596378)
	(JGroups-coord, 1.5426003209) +- (0, 0.00829044483237941)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 kill/minute, 1 add/minute};
\end{tikzpicture}

### 2016-12-01 report/figures/total-bytes.tex
Add EpTO figures

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=0,
ymax=35,
width=\linewidth / 2.5,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.3,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups},
]
\nextgroupplot
% 50-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 10.8956669416) +- (0, 0.112678096321221)
	(JGroups, 0.6420599937) +- (0, 0.0164341768113215)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 10.8952655703) +- (0, 0.112677849128705)
	(JGroups, 0.628793349) +- (0, 0.0164560780074577)
};
\legend{receive, send}
\nextgroupplot
% 50-2sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 22.1900560357) +- (0, 0.513548491076014)
	(JGroups, 1.1888473068) +- (0, 0.0236261886904278)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 22.1896583376) +- (0, 0.513548672987696)
	(JGroups, 1.1756330318) +- (0, 0.0236412207168411)
};
\nextgroupplot
% 100-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 25.9895900054) +- (0, 0.285368750128887)
	(JGroups, 1.6842673321) +- (0, 0.014352534074787)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 25.9864516777) +- (0, 0.285365521649361)
	(JGroups, 1.6291726389) +- (0, 0.0124299318567327)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}

### 2016-12-01 report/figures/total-events-sent-churn.tex
Add EpTO figures

\begin{tikzpicture}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=53000,
ymax=65000,
width=\linewidth / 1.75,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
scaled ticks=false, tick label style={/pgf/number format/fixed},
ylabel={Events sent},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
\nextgroupplot
% 100-1sec 1kill/min
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 54169.6) +- (0, 158.689214084218)
	(JGroups-nocoord, 53981.9) +- (0, 102.155056436554)
	(JGroups-coord, 53953.4) +- (0, 180.331300050151)
};
%
\nextgroupplot
% 100-1sec 1kill, 1add/min
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59853.9) +- (0, 228.339148344445)
	(JGroups-nocoord, 57813.8) +- (0, 128.414087146925)
	(JGroups-coord, 57873.3) +- (0, 150.893671172785)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}

### 2016-12-01 report/figures/total-events-sent.tex
Add EpTO figures

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=40000,
ymax=155000,
width=\linewidth / 2.5,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.3,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
ylabel={Events sent},
scaled ticks=false, tick label style={/pgf/number format/fixed},
symbolic x coords={EpTO, JGroups},
]
\nextgroupplot
% 50-1sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59994.9) +- (0, 2.76686746259348)
	(JGroups, 59993) +- (0, 15.1217282962848)
};
%
\nextgroupplot
% 50-2sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 119908.5) +- (0, 8.04501225742594)
	(JGroups, 119924.7) +- (0, 5.69697775628018)
};
\nextgroupplot
% 100-1sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59993) +- (0, 237.827201509368)
	(JGroups, 59974.6) +- (0, 216.060279654648)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}

### 2016-12-02 report/evaluation.tex
Fix small typos

@@
 
-Comparing \epto and \jgroups in terms of bandwidth when we increase the number of events sent per second, we can see the bandwidth doubling in both cases.
-In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups has the upper hand.
+Comparing \epto and \jgroups in terms of bandwidth when we increase the number of events sent per second, we can see the bandwidth doubling in both cases.In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups is clearly at an advantage.
 Since \epto has a worse baseline we will reach the maximum bandwidth possible much quicker when increasing the event throughput.
@@
 Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it by 50\% or less.
-\jt{logarithmic I think as per the formula}.
+\jt{logarithmic I think as was expected}.
 Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum bandwidth.
@@
 In \autoref{fig:total-bandwidth}, \epto has a worse baseline than \jgroups.
-This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at minimum $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect.
+This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect.
 Here we are well within this ratio.
@@
 
-The fact that killing the coordinator or not is of no effect on the total bandwidth used tends to show the hypothesis to be correct.
+The fact that killing the coordinator or not has no effect on the total bandwidth used tends to show the hypothesis to be correct.
 \subsection{Local Times}
@@
 They are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}.
-These global times are of less interest than their local counterpart as they differences in clocks between hosts can skew this measurement.
+These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.
 
@@
 
-However when killing and adding nodes, we notice that \jgroups delivers more than 2000 less events than \epto, even when the coordinator is not killed.
+However when killing and adding nodes \epto is clearly superior, delivering more than 2000 more events than \jgroups, even when \jgroups coordinator is not killed.
 

### 2016-12-02 results/check_order.py
Fix small stuff in log scripts

@@ -125,7 +125,9 @@ if len(complete_list) == sent_events:
 # There might be a problem
 churn_problem = False
 if not no_problem:
-    for event in sent_events:
+    logging.info('Checking for possible churn problem...')
+    bar = progressbar.ProgressBar()
+    for event in bar(sent_events):
         if not any(True for event_list in events.values() if event in event_list):
             churn_problem = True
             logging.info('Event {:s} was never sent due to churn. Please remove it.'.format(event))

### 2016-12-02 results/epto-tester.py

@@ -296,4 +296,4 @@ with open('event-sent-stats.csv', 'w', newline='') as csvfile:
     print('Writing events sent to csv file...')
     bar = progressbar.ProgressBar()
     for event_sent in stats_events_sent:
-        writer.writerows({'events-sent': event_sent})
+        writer.writerow({'events-sent': event_sent})

### 2016-12-02 results/epto-tester.py
Make scripts compatible with nanoTime

@@ -35,7 +35,7 @@ CHURN_NUMBER = 17
 expected_ratio = 1 - (1 / (PEER_NUMBER ** args.constant))
 k = ttl = delta = 0
 local_deltas = []
-events_sent = {}
+# events_sent = {}
 events_delivered = {}
 
 
@@ -70,7 +70,7 @@ def extract_stats(file):
     delta = match_line(r'\d+ - Delta: (\d+)')
     start_at = match_line(r'(\d+) - Sending:')
     file.seek(0)  # Start again
-    events_sent = {}
+    # events_sent = {}
     events_delivered = {}
     local_deltas = []
 
@@ -86,15 +86,16 @@ def extract_stats(file):
                 time = int(match.group(1))
                 event = match.group(2)
                 events_delivered[event] = time
-                # Compute local deltas
-                if event in events_sent:
-                    local_deltas.append(events_delivered[event] - events_sent[event])
+                local_match = re.match(r'\d+ - Delivered: .+ -- Local Delta: (\d+)', line)
+                if local_match:
+                    delta = int(local_match.group(1)) / (10**6)
+                    local_deltas.append(delta)
                 result = int(match.group(1))
                 pos = file.tell()
                 continue
             match = re.match(r'(\d+) - Sending: (\[.+\])', line)
             if match:
-                events_sent[match.group(2)] = int(match.group(1))
+                # events_sent[match.group(2)] = int(match.group(1))
                 events_sent_count += 1
                 continue
             if re.match(r'.+ - Time given was smaller than current time', line):
@@ -111,13 +112,13 @@ def extract_stats(file):
     # Only count complete peers
     if not balls_sent:
         return Stats(State.dead, start_at, end_at, end_at - start_at, evts_sent,
-                     None, None, None), events_sent, events_delivered, local_deltas
+                     None, None, None), events_delivered, local_deltas
     balls_received = match_line(r'\d+ - Balls received: (\d+)')
     messages_sent = match_line(r'\d+ - Events sent: (\d+)')
     messages_received = match_line(r'\d+ - Events received: (\d+)')
 
     return Stats(state, start_at, end_at, end_at - start_at, messages_sent,
-                 messages_received, balls_sent, balls_received), events_sent, events_delivered, local_deltas
+                 messages_received, balls_sent, balls_received), events_delivered, local_deltas
 
 
 def all_stats(files):
@@ -125,20 +126,20 @@ def all_stats(files):
[…]

### 2016-12-12 results/epto-tester.py
Don't sum balls equal to None

@@ -246,8 +246,8 @@ print("Population std fo the time to deliver: %f ms" % statistics.pstdev(global_
 print("-------------------------------------------")
 messages_sent = [stat.msg_sent for stat in stats]
 messages_received = [stat.msg_received for stat in perfect_stats]
-balls_sent = [stat.balls_sent for stat in stats]
-balls_received = [stat.balls_received for stat in stats]
+balls_sent = [stat.balls_sent for stat in stats if stat.balls_sent]
+balls_received = [stat.balls_received for stat in stats if stat.balls_received]
 print(len(stats))
 print(len(balls_sent))
 print(len(balls_received))

### 2016-12-22 report/abstract.tex
Start working on report

\begin{abstract}
\epto is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time. In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics.\jt{Speak about results}
\end{abstract}

### 2016-12-22 report/definitions.tex
Start working on report

\section{Definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process will communicate with other processes in the distributed system, exchange events, and order them together.
\par
An event is defined as data sent at a given time by a peer. For example, we could imagine a system where each process publishes some data to other peers. The moment where we publish data combined with the data is called an event.
\par
A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic and make it scalable in terms of processes and events.

We define \epto scaling well as it was defined in  \autocite{matos2015epto}:  ``The number
of messages transmitted per process per round is logarithmic
in the number of processes, ...''. The number of rounds needed to deliver an event stays low as well.
\par
Since \epto uses a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability). In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
\par
We write the cluster parameters as $(p,e)$, where $p$ designates the total number of peers and $e$ designates the global event throughput. We tested \epto and \jgroups with three different cluster parameters:
\begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
	\item[\textbf{$(100,100)$}:] 50 peers with a global event throughput of 50 events per seconds.
	\item[\textbf{$(100,100)$}:] 50 peers with a global event throughput of 100 events per seconds.
	\item[\textbf{$(100,100)$}:] 100 peers with a global event throughput of 50 events per seconds. This parameter was also used for all synthetic churns.
\end{description}
\par

### 2016-12-22 report/evaluation.tex
Start working on report

@@
 \section{Evaluation}
+\label{sec:evaluation}
+\pgfplotstableread[col sep=comma]{csv-data/average-bytes-sent-recv.csv}\tableaverage
+\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
+\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
+\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
+In this section we present the results of the benchmarks.
+\subsection{Testbed}
+\jt{Write about benchmarks}
+\subsection{Parameter}
+\epto uses a constant $c$ to set the probability of a hole appearing.
+We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions.
+We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
+\par 
+We used specific settings recommended for a big cluster available in the \jgroups library.
 \subsection{Bandwidth}
-\jt{\epto was run with $c=4$ to prevent any holes and a $\delta$ period of 100ms}
-\begin{figure}[h]
+\jt{\epto was run with $c=4$ to prevent any holes and a $\delta$ period of 100ms.
+Churn was run for 17minutes.
+3min at the end with no churn, to see stabilization.
+Churn starts 30seconds after protocol}
+\jt{We should define acronyms for the different benchmark variants}
+
+\begin{figure}[htp]
 	\centering
-	\input{figures/bandwidth-figure.tex}
+	\input{figures/no-churn/bandwidth-figure.tex}
 	\vspace{-2mm} 
@@
 In \autoref{fig:bandwidth} we can see \epto has a worse baseline compared to \jgroups.
-It uses a median bandwidth of approximately \SI{1}{\mbps} for the smallest benchmark whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}.
+It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}.
 However, in \jgroups most of the work is done solely by the coordinator.
-We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.5}{\mbps}.
+We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.6}{\mbps}.
 
@@
 
-\begin{figure}
+\begin{figure}[htp]
 	\centering
-	\input{figures/bandwidth-figure-churn.tex}
+	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
 	\vspace{-2mm} 
@@
 
-All tests were run using 100 peers and a global throughput of 50 events per second (100-1sec). \jt{I need to rephrase 50-1sec, 100-1sec etc. to be clearer}
-
 We can see that the churn doesn't affect  \epto at all when there are only nodes leaving.
@@
 
-On the other hand, when killing the coordinator in \jgroups we can see a huge spike in bandwidth, going from \SI{1}{\mbps} to more than \SI{15}{\mbps}.
+On the other hand, when killing the coordinator in \jgroups we can see a huge spike in bandwidth, going from \SI{1.2}{\mbps} to more than \SI{15}{\mbps}.
 This is due to how \jgroups operates when selecting a new coordinator.
@@
 
-\subsection{Total Bytes sent/received}
-\begin{figure}[h]
+\subsection{Total GigaBytes sent/received}
+\begin{table}[htp]
 	\centering
-	\input{figures/total-bytes.tex}
-	\vspace{-2mm} 
-	\caption{Total bytes sent/received during an average experiment}
-	\vspace{-2mm} 
-	\label{fig:total-bandwidth}
-\end{figure}
-In \autoref{fig:total-bandwidth}, \epto has a worse baseline than \jgroups.
+	\caption{Total \si{\giga\byte} sent/received in a stable system}
+	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\begin{tabular}{lSSSS}
+		\toprule
+		&& \multicolumn{3}{c}{Cluster parameters} \\
+		\cmidrule{3-5}
+		{Protocol}&& {50P50E} & {50P50E} & {100P50E} \\
+		\midrule
+		\multirow{2}{*}{\epto}&{Receive}& 10.85(016) & 22.31(039) & 26.01(028) \\
+						    &{Sending}& 10.85(016) & 22.31(039) & 26.01(028)\\
+		\midrule
+		\multirow{2}{*}{\jgroups}&{Receive}& 0.81(003) & 1.48(001) & 2.00(001)\\
+							   &{Sending}& 0.80(003) & 1.47(001) & 1.95(001)\\
+		\bottomrule
+	\end{tabular}
+	\label{table:total-bandwidth} 
+\end{table}
+
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/no-churn/total-bytes.tex}
+%	\vspace{-2mm} 
+%	\caption{Total bytes sent/received during an average experiment}
+%	\vspace{-2mm} 
+%	\label{fig:total-bandwidth}
+%\end{figure}
+In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups.
 This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect.
 Here we are well within this ratio.
+\begin{table}[htp]
+	\centering
+	\caption{Total \si{\giga\byte} sent/received with a synthetic churn}
+	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\begin{tabular}{lSSS}
+		\toprule
+		&& \multicolumn{2}{c}{Churn parameters} \\
+		\cmidrule{3-4}
+		{Protocol}&& \{1kill\}/minute & \{1kill,1add\}/minute \\
+		\midrule
+		\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
+							&{Sending}& 21.21(025) & 26.57(032)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.55(002) & 1.85(002)\\
+									 &{Sending}& 1.51(002) & 1.79(002)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.54(001) & 1.83(002)\\
+									   &{Sending}& 1.49(001) & 1.78(002)\\
+		\bottomrule
+	\end{tabular}
+	\label{table:total-bandwidth-churn} 
+\end{table}
 
-\begin{figure}[h]
-	\centering
-	\input{figures/total-bytes-churn.tex}
-	\vspace{-2mm} 
-	\caption{Total bytes sent/received during an average experiment with churn}
-	\vspace{-2mm} 
-	\label{fig:total-bandwidth-churn}
-\end{figure}
-In \autoref{fig:total-bandwidth-churn} we can see that \jgroups total bandwidth usage doesn't increase significantly compared to \epto with churn.
-An explanation for this phenomenon is that \jgroups takes more than a minute to find out the coordinator is effectively dead.
-During this time the bandwidth is practically not used.
-The big spike when the new coordinator is chosen compensates for this hole thus making the total bandwidth used appear as to not have changed.
-
-The fact that killing the coordinator or not has no effect on the total bandwidth used tends to show the hypothesis to be correct.
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/synthetic-churn/total-bytes-churn.tex}
+%	\vspace{-2mm} 
+%	\caption{Total bytes sent/received during an average experiment with churn}
+%	\vspace{-2mm} 
+%	\label{fig:total-bandwidth-churn}
+%\end{figure}
+In \autoref{table:total-bandwidth-churn} we can see that \jgroups total bandwidth usage is smaller when there is churn.
+One hypothesis for this is that a JGroups replica takes a longer time to start up thus the overall benchmark has a longer time with less than 100 replicas.
+We also do not see a difference whether we kill the coordinator or not.
+This can be explained by the fact that before the detection of the faulty coordinator \jgroups is forced to a stop for time up to \SI{20}{\second}.
+The big spike afterwards compensates for this hole.
 \subsection{Local Times}
@@
 \jt{Having a table for each percentile figure might be a good idea to put numbers for key percentiles (min,50th,max) for example}
-\begin{figure}[h]
+\begin{figure*}[htp]
 	\centering
-	\input{figures/local-times.tex}
+	\input{figures/no-churn/local-times.tex}
 	\vspace{-2mm} 
@@
 	\label{fig:local-times} 
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure}[htp]
 	\centering
-	\input{figures/local-times-churn.tex}
+	\input{figures/synthetic-churn/local-times-churn.tex}
 	\vspace{-2mm} 
@@
 \subsection{Global Times}
-\begin{figure}[h]
+\begin{figure*}[htp]
 	\centering
-	\input{figures/global-times.tex}
+	\input{figures/no-churn/global-times.tex}
 	\vspace{-2mm} 
@@
 	\label{fig:global-times}  
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure}[htp]
 	\centering
-	\input{figures/global-times-churn.tex}
+	\input{figures/synthetic-churn/global-times-churn.tex}
 	\vspace{-2mm} 
@@
 \subsection{Local Dissemination stretch}
-\begin{figure}[h]
+\begin{figure*}[htp]
 	\centering
-	\input{figures/local-delta.tex}
+	\input{figures/no-churn/local-delta.tex}
 	\vspace{-2mm} 
@@
 	\label{fig:local-delta}  
-\end{figure}
-\jt{I wonder if we should split the plots (3 by 2) having a row for EpTO and a row for JGroups.
-This way we could show more for JGroups local dissemination stretch}In \autoref{fig:local-delta}, We can see the percentiles of the local dissemination stretch.
+\end{figure*}
+In \autoref{fig:local-delta}, We can see the percentiles of the local dissemination stretch.
 The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
@@
 This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minmimal latency.
-The median dissemination stretch of \jgroups is around \SI{8}{\milli\second} where as the median dissemiantion stretch of \epto is around \SI{685}{\milli\second}.
-When increasing the number of peers, JGroups starts to have long delivery times
+The median dissemination stretch of \jgroups is around \SI{7}{\milli\second} where as the median dissemiantion stretch of \epto is around \SI{630}{\milli\second} for $(100,50)$.
+When increasing the number of peers, \jgroups starts to have long delivery times for some outliers.
 
-\begin{figure}[h]
+\begin{figure}[htp]
 	\centering
-	\input{figures/local-delta-churn.tex}
+	\input{figures/synthetic-churn/local-delta-churn.tex}
 	\vspace{-2mm} 
@@
 In \autoref{fig:local-delta-churn} We can see a completely different picture.
-When under churn, the 90th percentile of \jgroups is at \SI{447}{\milli\second} and the highest percentiles are at more than \SI{20}{\second} with the highest dissemination stretch being \SI{67.5}{\second}.
+When under churn, the 95th percentile of \jgroups is at \SI{31}{\milli\second} compared to \SI{14}{\milli\second} when there is no churn.
+The highest percentiles are at more than \SI{10}{\second}.
 This effect is due to the coordinator dying as we clearly see that it does not happen when we do not kill it.
 
-The median is bigger at around \SI{11}{\milli\second}, whether we kill the coordinator or not. This shows that there are some degradation in \jgroups local dissemination stretch when under churn.
+The median is bigger at around \SI{9}{\milli\second}, whether we kill the coordinator or not. This shows that there are some degradation in \jgroups local dissemination stretch when under churn.
 
 On the contrary, \epto performs very well under churn.
-The median stays really close at \SI{686}{\milli\second} with the 99.9th percentile being at \SI{1647}{\milli\second} compared to \SI{1366}{\milli\second} when no churn is happening.
+The median degrade a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
 \newpage
 \subsection{Events sent}
-\begin{figure}[h]
+\jt{Representing as a table might be better}
+\begin{table}[htp]
 	\centering
-	\input{figures/total-events-sent.tex}
+	\caption{Total events sent in a stable environment}
+\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+\begin{tabular}{lSSS}
+	\toprule
+	& \multicolumn{3}{c}{Cluster parameters} \\
+	\cmidrule{2-4}
+	Protocol & {50P50E} & {50P50E} & {100P50E} \\
+	\midrule
+	\epto & 59993.8(33) & 119898.2(97) & 59913.0(1643) \\
+	\jgroups & 59961.9(109) & 119885.7(50) & 60023.1(2871) \\
+	\bottomrule
+\end{tabular}
+\label{table:total-events}  
+\end{table}
+
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/no-churn/total-events-sent.tex}
+%	\vspace{-2mm} 
+%	\caption{Total events sent per experiment on average}
+%	\vspace{-2mm}
+%	\label{fig:total-events}   
+%\end{figure}
+In \autoref{table:total-events} we can see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
+\begin{table}[htp]
+	\centering
+	\caption{Total events sent with a synthetic churn}
+\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+\begin{tabular}{lSSS}
+	\toprule
+	& \multicolumn{2}{c}{Cluster parameters} \\
+	\cmidrule{2-3}
+	Protocol & \{1kill\}/minute & \{1kill,1add\}/minute \\
+	\midrule
+	\epto & 53898.5(1339) & 59798.6(1401) \\
+	\jgroups-coord & 53834.7(1755) & 59507.9(2409) \\
+	\jgroups-nocoord & 53830.5(2003) & 59450.5(1751) \\
+	\bottomrule
+\end{tabular}
+    \label{table:total-events-churn}
+\end{table}
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/synthetic-churn/total-events-sent-churn.tex}
+%	\vspace{-2mm} 
+%	\caption{Total events sent per experiment during churn on average}
+%	\vspace{-2mm}
+%	\label{fig:total-events-churn}  
+%\end{figure}
+
+In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
+When killing and adding nodes, JGroups seems to deliver a smaller amount of nodes, however it does not look significant \jt{I didn't run any statistical analysis}
+\section{Real Trace}
+\subsection{Bandwidth}
+
+\begin{figure}[htp]
+	\centering
+	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
 	\vspace{-2mm} 
-	\caption{Total events sent per experiment on average}
+	\caption{Throughput percentiles of a node during an experiment with churn}
 	\vspace{-2mm} 
-	\label{fig:total-events}   
-\end{figure}
-In \autoref{fig:total-events} we can see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
-\begin{figure}[h]
-	\centering
-	\input{figures/total-events-sent-churn.tex}
-	\vspace{-2mm} 
-	\caption{Total events sent per experiment during churn on average}
-	\vspace{-2mm}
-	\label{fig:total-events-churn}  
+	\label{fig:bandwidth-real-churn}
 \end{figure}
 
-In \autoref{fig:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
+\newpage
+\subsection{Total Bytes sent/received}
+\begin{table}[htp]
+	\centering
+	\caption{Total \si{\giga\byte} sent/received}
+	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\begin{tabular}{lSS}
+		\toprule
+		&& \multicolumn{1}{c}{Churn parameters} \\
+		\cmidrule{3-3}
+		{Protocol}&& {Real Trace} \\
+		\midrule
+		\multirow{2}{*}{\epto}&{Receive}& 81.41(108)\\
+		&{Sending}& 82.67(108)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.61(008)\\
+		&{Sending}& 5.45(008)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.63(005)\\
+		&{Sending}& 5.48(005)\\
+		\bottomrule
+	\end{tabular}
+	\label{table:total-bandwidth-real-churn} 
+\end{table}
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/real-churn/total-bytes-real-churn.tex}
+%	\vspace{-2mm} 
+%	\caption{Total bytes sent/received during an average experiment with churn}
+%	\vspace{-2mm} 
+%	\label{fig:total-bandwidth-real-churn}
+%\end{figure}
 
-However when killing and adding nodes \epto is clearly superior, delivering more than 2000 more events than \jgroups, even when \jgroups coordinator is not killed.
+\subsection{Local Times}
+\label{sub:local-times}
+\begin{figure}[htp]
+	\centering
+	\input{figures/real-churn/local-times-real-churn.tex}
+	\vspace{-2mm} 
+	\caption{Local dissemination times}
+	\vspace{-2mm}
+	\label{fig:local-times-real-churn} 
+\end{figure}
 
-We are not sure why this is happening, but one hypothesis is that \jgroups peers take a longer time to join the cluster and thus start sending events later.
-Since the benchmarks start and stop at a given time, they send less events in total.
+\newpage
+\subsection{Global Times}
+
+\begin{figure}[htp]
+	\centering
+	\input{figures/real-churn/global-times-real-churn.tex}
+	\vspace{-2mm} 
+	\caption{Global dissemination times with churn}
+	\vspace{-2mm} 
+	\label{fig:global-times-real-churn} 
+\end{figure}
+\newpage
+\subsection{Local Dissemination stretch}
+
+\begin{figure}[htp]
+	\centering
+	\input{figures/real-churn/local-delta-real-churn.tex}
+	\vspace{-2mm} 
+	\caption{Local dissemination stretch with churn}
+	\vspace{-2mm}
+	\label{fig:local-delta-real-churn}   
+\end{figure}
+\newpage
+\subsection{Events sent}
+\begin{table}[htp]
+	\centering
+	\caption{Total events sent during a real trace}
+\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+\begin{tabular}{lS}
+	\toprule
+	Protocol & \\
+	\midrule
+	\epto & 165844.2(2102) \\
+	\jgroups-coord & 166183.0(13681) \\
+	\jgroups-nocoord & 166585.8(8249) \\
+	\bottomrule
+\end{tabular}
+    \label{table:total-sent-real-churn}
+\end{table}
+%\begin{figure}[h]
+%	\centering
+%	\input{figures/real-churn/total-events-sent-real-churn.tex}
+%	\vspace{-2mm} 
+%	\caption{Total events sent per experiment during churn on average}
+%	\vspace{-2mm}
+%	\label{fig:total-events-real-churn}  
+%\end{figure}

### 2016-12-22 report/figures/no-churn/bandwidth-figure.tex
Start working on report

\begin{tikzpicture}[domain=0:1800]
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth / 2.6,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=3 by 2,
	vertical sep=8mm,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	yticklabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
ymax=4.5,
ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(1.9,1.5)},anchor=north, font=\small, draw=none},]
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-50-1sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-1sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-50-2sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-50-2sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-100-1sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-100-1sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},yticklabels={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-50-1sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-1sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-50-2sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-50-2sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=1.5,ytick={0,0.5,1,1.5},]
\addplot table[x=time,y=JGroups-100-1sec-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-100-1sec-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\node[anchor=south, rotate=-90] at (group c3r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c3r2.east){JGroups};
\end{tikzpicture}

### 2016-12-22 report/figures/no-churn/global-times.tex
Start working on report

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 3,
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.6,1.4)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tableglobaltime};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tableglobaltime};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tableglobaltime};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}

### 2016-12-22 report/figures/no-churn/local-delta.tex
Start working on report

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=3.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.7,1.7)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tablelocaldeltas};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}

### 2016-12-22 report/figures/no-churn/local-times.tex
Start working on report

--- 2016-12-22 report/figures/no-churn/global-times.tex
@@
 \nextgroupplot
-\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaltime};
 \legend{EpTO, JGroups}
@@
 \nextgroupplot
-\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tablelocaltime};
 %
 \nextgroupplot
-\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tablelocaltime};
 \end{groupplot}

### 2016-12-22 report/figures/real-churn/bandwidth-figure-real-churn.tex
Start working on report

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=1 by 4,
	vertical sep=6mm,
	horizontal sep=0mm,
	xlabels at=edge bottom,
	xticklabels at=edge bottom,
	ylabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
xmax=61,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(0.2,2)},anchor=north, font=\small, draw=none},]
\nextgroupplot[height=2.2cm, 
ylabel={Nodes}, 
ymin=80,ymax=120,
ytick={80,120}, 
yticklabels={80,...,120},
xmajorgrids,
yminorgrids,
tick label style={font=\footnotesize},
label style={font=\tiny}]
\addplot[const plot, color=blue, mark=none] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4}]
\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\end{groupplot}
%
\node[anchor=south, rotate=-90] at (group c1r2.east) {EpTO};
\node[anchor=south, rotate=-90] at (group c1r3.east) {JGroups coord};
\node[anchor=south, rotate=-90] at (group c1r4.east) {JGroups nocoord};
\end{tikzpicture}

### 2016-12-22 report/figures/real-churn/global-times-real-churn.tex
Start working on report

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tableglobaltime};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
\end{axis}
\end{tikzpicture}

### 2016-12-22 report/figures/real-churn/local-delta-real-churn.tex
Start working on report

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
scaled x ticks = false,
x tick label style={/pgf/number format/fixed},
xlabel={Time $\left[\si{\second}\right]$},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaldeltas};
\legend{EpTO, JGroups-coord, JGroups-nocoord}
\end{axis}
\end{tikzpicture}

### 2016-12-22 report/figures/real-churn/local-times-real-churn.tex
Start working on report

--- 2016-12-22 report/figures/real-churn/global-times-real-churn.tex
@@
 ]
-\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaltime};
 \legend{EpTO, JGroups-nocoord, JGroups-coord}

### 2016-12-22 report/figures/real-churn/total-bytes-real-churn.tex
Start working on report

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{axis}[
ymin=0,
ymax=85,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
% 100-1sec 1kill/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 81.4132665062) +- (0, 1.0822819001435)
	(JGroups-nocoord, 5.6382174422) +- (0, 0.058029870294168)
	(JGroups-coord, 5.6118450728) +- (0, 0.089291024953546)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 82.6741897528) +- (0, 1.08368411575119)
	(JGroups-nocoord, 5.4807749566) +- (0, 0.056977818312573)
	(JGroups-coord, 5.4548362802) +- (0, 0.086273738554255)
};
\legend{receive, send}
\end{axis}
\end{tikzpicture}

### 2016-12-22 report/figures/real-churn/total-events-sent-real-churn.tex
Start working on report

\begin{tikzpicture}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{axis}[
ymin=160000,
ymax=170000,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
scaled ticks=false, tick label style={/pgf/number format/fixed},
ylabel={Events sent},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
% 100-1sec real trace
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Events sent
	(EpTO, 165844.2) +- (0, 210.222976860287)
	(JGroups-nocoord, 166585.8) +- (0, 824.998303028558)
	(JGroups-coord, 166183) +- (0, 1368.95507596122)
};
\end{axis}
\end{tikzpicture}

### 2016-12-22 report/figures/synthetic-churn/bandwidth-figure-churn.tex
Start working on report

\begin{tikzpicture}[domain=0:1800]
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth / 2.5,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=2 by 3,
	vertical sep=8mm,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	yticklabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(1.25,1.5)},anchor=north, font=\small, draw=none},]
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-suspend-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-add-suspend-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},
yticklabels={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-suspend-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-add-suspend-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},
yticklabels={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-suspend-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\end{groupplot}
\node[anchor=south] at (group c1r1.north) {1 kill};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\node[anchor=south, rotate=-90] at (group c2r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c2r2.east){JGroups coord};
\node[anchor=south, rotate=-90] at (group c2r3.east){JGroups nocoord};
\end{tikzpicture}

### 2016-12-22 report/figures/synthetic-churn/global-times-churn.tex
Start working on report

--- 2016-12-01 report/figures/local-times-churn.tex
@@
 \nextgroupplot
-\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
-\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
-\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
+\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tableglobaltime};
 \legend{EpTO, JGroups-nocoord, JGroups-coord}
@@
 \nextgroupplot
-\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
-\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
-\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
+\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {\tableglobaltime};
 \end{groupplot}

### 2016-12-22 report/figures/synthetic-churn/local-delta-churn.tex
Start working on report

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
height=5cm, width=\linewidth / 1.75,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=25,
scaled x ticks = false,
x tick label style={/pgf/number format/fixed},
xlabel={Time $\left[\si{\second}\right]$},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tablelocaldeltas};
\legend{EpTO, JGroups-coord, JGroups-nocoord}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {\tablelocaldeltas};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}

### 2016-12-22 report/figures/synthetic-churn/local-times-churn.tex
Start working on report

--- 2016-12-22 report/figures/synthetic-churn/global-times-churn.tex
@@
 \nextgroupplot
-\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tablelocaltime};
 \legend{EpTO, JGroups-nocoord, JGroups-coord}
@@
 \nextgroupplot
-\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {\tableglobaltime};
-\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {\tableglobaltime};
+\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {\tablelocaltime};
+\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {\tablelocaltime};
 \end{groupplot}

### 2016-12-22 report/framework.tex
Start working on report

\section{Framework}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
\subsection{Architecture}
Figure \ref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can broadcast and delivers events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every events received from the network are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to gather a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not obligatory. In practice it works well, but using DHTs is certainly a possibility.
\subsection{Implementation}
We implemented the data sent as randomly generated UUIDs. We implemented our own PSS CYCLON operating on its own port. When implementing \jgroups we had to use TCPGOSSIP instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. The framework is open-sourced on Github.
\subsection{Deployment}
To deploy \epto effectively we effectively need two different services. One containing the tracker and one containing all \epto replicas. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal tweaks. Every benchmark is launched through a Python script available on the master mode. This script handles everything, from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. \jt{I should create a figure showcasing the complete architecture}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}, although the synthetic churn was improved to support adding and removing nodes at the same time.

### 2016-12-22 report/introduction.tex
Start working on report

--- 2016-06-16 report/paper.tex
@@
-\documentclass[10pt,conference,a4paper]{IEEEtran}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-% Correct date format in references but with american hyphenation, quotes, ...
-% Trick from http://tex.stackexchange.com/a/129209
-\usepackage[australian,american]{babel}
+\section{Introduction}
 
-% *** GRAPHICS RELATED PACKAGES ***
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-\else
-  \usepackage[dvips]{graphicx}
-\fi
-
-% *** MATH PACKAGES ***
-\usepackage{amssymb}
-\usepackage[cmex10]{amsmath}
-
-% Figure captions in small font
-\makeatletter
-\let\MYcaption\@makecaption
-\makeatother
-\usepackage[font=footnotesize]{subcaption}
-\makeatletter
-\let\@makecaption\MYcaption
-\makeatother
-
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.9}
-%TODO minted ?
-\usepackage[binary-units]{siunitx}
-\usepackage{booktabs}
-\usepackage{pifont}
-\usepackage{tikz}
-\usepackage{csquotes}
-\usepackage[backend=biber,style=ieee,minbibnames=1,maxbibnames=99]{biblatex}
-\usepackage{xpatch}
-\usepackage{url}
-\usepackage[backgroundcolor=lightgray]{todonotes}
-\usepackage[hidelinks]{hyperref}
-
-\bibliography{paper}
-% Small font for references
-\renewcommand*{\bibfont}{\small}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\def\mytitle{Practical Implementation of EpTO: An Epidemic Total Order Algorithm}
-\title{\mytitle}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\def\jocelyn{Jocelyn Thode}
-\def\ehsan{Ehsan Farhadi}
-\author{
-	\IEEEauthorblockN{\jocelyn}
-	\IEEEauthorblockA{Université de Fribourg\\
-	Fribourg, Switzerland\\
-	\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch}
-	}
-	\and
-	\IEEEauthorblockN{\ehsan}
-	\IEEEauthorblockA{Université de Neuchâtel\\
-		Neuchâtel, Switzerland\\
-		\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}
-	}
-
-}
-
-% Set PDF file properties
-\hypersetup{
-	pdftitle=\mytitle,
-}
-
-% Tick marks
-\newcommand{\cmark}{\ding{51}}
-\newcommand{\xmark}{\ding{55}}
-
-% Columns balancing
-\renewbibmacro{finentry}{%
-  \iffieldequalstr{entrykey}{jujjuri2010virtfs}%<- key after which you want the break
-   {\finentry\newpage}
-   {\finentry}}
-
-\xpatchbibdriver{online}
-{\printtext[parens]{\usebibmacro{date}}}
-{\iffieldundef{year}
-	{}
-	{\printtext[parens]{\usebibmacro{date}}}}
-{}
-{\typeout{There was an error patching biblatex-ieee (specifically, ieee.bbx's @online driver)}}
-
-\begin{document}
-\graphicspath{{figures/}}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-Total ordering has always been of interest in large-scale distributed systems.
-EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions.
-We then compare it to a deterministic total order algorithm named JGroups.
-\end{abstract}
-\section{Introduction}
 Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
-One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
-EpTO is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, EpTO is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.
+One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}.
+\epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
+In addition, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.
 \par 
@@
 \par
-This is where EpTO comes into light. EpTO, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
-EpTO consists of two distinct parts.
+This is where \epto comes into light. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
+\epto consists of two distinct parts.
 The first part is probabilistic dissemination.
-EpTO guarantees that all peers will receive an event with arbitrarily high probability.
-The second part of EpTO is deterministic ordering.
-Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.
+\epto guarantees that all peers will receive an event with arbitrarily high probability.
+The second part of \epto is deterministic ordering.
+Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster id of the events.
 \par
-To model the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
-The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls.
+To model the first part, \epto is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls.
 In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?
@@
 With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Thanks to these approaches, EpTO becomes highly scalable and resilient, while still providing total order.
-
-\subsection{Contribution}
-Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-In this work, we implement EpTO in pure Kotlin using a modified version of the NeEM library \autocite{neem} and show that EpTO is suitable for real-world large-scale distributed systems.
-The modification to NeEM includes a dumbed-down gossip dissemination as seen in \autocite{matos2015epto}.
-We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in both stable (no-churn) and unstable systems.
-These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
-\todo[inline]{Here we will present our conclusion briefly}
+Thanks to these approaches, \epto becomes highly scalable and resilient, while still providing total order.
 \par
-In our experiments we assume each peer knows every other peers.
-We implement EpTO using this assumption and test it against JGroups.
-The details concerning the methodology are explained in subsection \ref{sub:metho}.
-In future work, we want to test EpTO when the membership is not entirely known.
-
-\section{Ordering Algorithms}
-Distributed systems, like centralized systems, need to preserve the temporal order of events produced by concurrent processes in the system.
-When there are separated processes that can only communicate through messages, you can’t easily order these messages.
-Therefore we need ordering algorithms to overcome this problem.
+Until now, the creators of \epto have only tested this algorithm in a simulated environment.
+In this work, we implement \epto in pure Kotlin and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems.
+We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
+\epto uses a PSS CYCLON to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.
+These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
+Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases} in the system.
 \par
-We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
-\subsection{Partial Order Algorithms}
-Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
-\begin{itemize}
-	\item Reflexivity: $a \leq a$ for all $a \in S$.
-	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
-	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
-\end{itemize}
-
-\subsection{Total Order Algorithms}
-A totally ordered set of events is a partially ordered set which satisfies one additional property:
-\begin{itemize}
-	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
-\end{itemize}
-\par
-In other words, total order is an ordering that defines the exact order of every event in the system.
-On the other hand, partial ordering only defines the order between certain key events that depend on each other.
-Partial order can be useful since it is less costly to implement.
-However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order.
-We then have to use total order, otherwise we can end up in an inconsistent state.
-
-\section{EpTO}
-\begin{figure}
-	\includegraphics[width=\linewidth]{figures/architecture.pdf}
-	\caption{EpTO architecture \autocite{matos2015epto}.}
-	\label{fig:epto-architecture}
-\end{figure}
-\subsection{Definitions}
-A process or peer is defined as an actor in our system running the application that needs total order.
-Each process will communicate with other processes in the distributed system, exchange events, and order them together.
-\par
-An event is defined as data sent at a given time by a peer.
-For example, we could imagine a system where each process publishes some data to other peers.
-The moment where we publish data combined with the data is called an event.
-\par
-A ball is a set of events bundled together and sent as one package. We use balls in EpTO to reduce network traffic and make it scalable in terms of processes and events.
-
-We define EpTO scaling well as it was defined in  \autocite{matos2015epto}:  ``The number
-of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver an event stays low as well.
-\par
-Since EpTO uses a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
-In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by EpTO's deterministic ordering algorithm, thus the total order property is preserved.
-\subsection{Dissemination Component}
-The Dissemination component bridges EpTO with the rest of the network.
-As we can see in Figure \ref*{fig:epto-architecture}, it receives balls, opens them, passes them to the Ordering component, and stores their events.
-Then, in the next round, it sends all the events received in the previous round to K other processes, where K denotes the gossip fan-out.
-
-\par
-When an application wants to publish an event, it will broadcast this event to the Dissemination component.
-
-\subsection{Ordering Component}
-The Ordering component is responsible for ordering events before delivering them to the application.
-To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which were delivered.
-\par
-In brief, the Ordering component first increments the timestamp of all the events which have been received in previous rounds to indicate the start of a new round.
-Then, it processes new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event).
-This is done to prevent delivering duplicate events.
-The remaining events in the received ball are added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
-\subsection{Stability Oracle}
-The Stability oracle is the component that outputs timestamps.
-It will increment its local clock every round as well as synchronize itself using timestamps of newly received events, to make sure the local clock does not drift too much.
-\par
-This Stability oracle offers an API to the Ordering component letting it know when an event is mature enough to be delivered to the application.
-\par
-As this component is local and only corrects itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
-\section{Performance Comparisons}
-In the following subsections, we will present the results obtained for the different tests ran and explain their meanings.
-\subsection{Methodology}
-\label{sub:metho}
-Our tests are run on a cluster provided by the University of Neuchâtel.
-For setting up our testbed we use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
-We were planning on using Docker \autocite{docker}, more specifically Docker-compose and docker-swarm in order to orchestrate and manage our cluster of peers on the OpenNebula but due to some issues with docker-swarm, we decided to deploy our application directly on the OpenNebula virtual machines, and control it using parallel ssh sessions.
-\par
-We deployed our application on a small network of 4 virtual machines and we distribute EpTO and JGroups peers equally over the network.
-\par
-To measure the throughput, we transmit a certain number of events by certain number of peers in a defined period of time, and then we measure the number of packets transmitted over the network.
-\par
-To measure the latency of both protocols, we measure the average time difference on a peer between the broadcast of an event and the time it delivers this same event.
-This way we can have an estimation of the approximate delay, and compare JGroups and EpTO.
-We execute EpTO and JGroups with 60, 100, 120 and 160 peers each peer broadcasting 12 events for 12 seconds (one event per second).
-\subsection{Peer membership known}
-As we measured and compared the latency of EpTO and JGroups we realize that JGroups is performing better than EpTO with smaller number of peers.
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/Delay.png}
-\caption{Approximate delay of delivering each event.}
-\label{fig:delay}
-\end{figure}
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/total_time.png}
-\caption{total run time of EpTO peers in order to deliver all events.}
-\label{fig:total_time}
-\end{figure}
-As we increased the number of peers, our results showed that JGroups is not as scalable as EpTO, but considering our limited resources we couldn't fully observe the difference between JGroup and EpTO on this matter.
-Although we should bear in mind that in this experiment EpTO has a relatively large delay due to ordering of the events without a central node (as in JGroups).
-But despite of this delay, we expected that EpTO performs much better and our analysis is that due the insufficient resource on the testbed.
-Epto could not show its advantages.
-\par
-We perceive this issue again when we measured the total time of execution of EpTO for delivering a certain number of events.
-\par
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/throughput.png}
-\caption{number of events transmitted packets over the network in 10 minutes.}
-\label{fig:throughput}
-\end{figure}
-For the throughput of both protocols we used a network sniffing tool, TCPdump.
-TCPdump is a tool for observing the network traffic.
-Using this tool we count number of packets (TCP packet for EpTO and UDP packet for JGroups) transmitting over the network for 10 minutes.
-And again, number of packets sent by EpTO peers are much higher than JGroups.
-This issue can be explain due the differences of the using protocols, TCP and UDP.
-TCP uses much more packets and data over the network for a connection to secure and guarantee the delivery of data.
-that's why in a perfect and lossless network, JGroups has better performance from perspective of network throughput.
-\section{Future Work}
-One of the students working on this project will do his master thesis on this project.
-
-Future work will include testing EpTO while having unknown peer membership.
-To account for that, we will need to modify NeEM to support the use of a Peer Sampling Service.
-We plan on implementing the CYCLON Peer Sampling Service \autocite{Voulgaris2005}.
-For now, we plan on using a tracker as in torrents to give each peer its initial view.
-This tracker is already implemented and works as a python web server that gets all alive peers and randomly select $K$ peers and sends this result to the EpTO peer.
-
-Our first implementation of EpTO focuses on correctness rather than fastness.
-Later work should optimize EpTO in regards to concurrency and network latency.
-To address the second point, we want to modify NeEM again to use UDP instead of TCP, to reduce latency and overhead in the network.
-\section{Conclusion}
-\todo[inline]{He we will summarize the results we found and present some future tasks that could be accomplished on EpTO}
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-This research was under supervision and help of two of the creators of EpTO, Dr.
-Hugues Mercier and Dr.
-Miguel Matos.
-We would like to thank them both for the help, insight and expertise they provided during this project.
-\par
-We would also like to thank Valerio Schiavoni for providing access to the cluster of University of Neuchâtel as well as Sebastien Vaucher for the help with setting up Docker, Docker-Compose and Docker Swarm.
-
-% references section, with correct date format
-\begin{otherlanguage}{australian}
-\printbibliography
-\end{otherlanguage}
-
-% that's all folks
-\end{document}
+The next sections are dedicated to present our benchmarking solution and to present the results obtained.
+Section \ref{sec:ordering} presents the different kind of ordering.
+Section \ref{sec:epto} presents \epto and the architecture of our implementation.
+Section \ref{sec:evaluation} shows and explains the results obtained.
+We then argue about future work in section \ref{sec:future-work}.
+We finally draw conclusions in section \ref{sec:conclusion}.

### 2016-12-22 report/main.tex
Start working on report

% !TeX program = lualatex
\documentclass[10pt,conference,a4paper]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[rm={oldstyle=false}]{cfr-lm}
\usepackage[australian,american]{babel}

\usepackage[backend=biber,style=ieee,bibencoding=utf8,sorting=none,doi=false,isbn=false,url=false,date=short]{biblatex}
\usepackage{amssymb}
\usepackage[cmex10]{amsmath}
\usepackage{csquotes}
\usepackage[nolist]{acronym}
\usepackage{float}
\usepackage{todonotes}
\usepackage{pgfplots}
\usepackage{pgf}
\usepackage[binary-units]{siunitx}
\usepackage{pgfplotstable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lipsum}
\usepackage[hidelinks]{hyperref}
\usepackage{xspace}
\usepackage[caption=false]{subfig}

\usetikzlibrary{patterns}

\usepgfplotslibrary{external, groupplots, units}
\tikzexternalize
\tikzsetexternalprefix{generated-figures/}

\sisetup{
	round-mode          = places,
	round-precision     = 2,
}
\DeclareSIUnit\mbps{MB/s}

\pgfplotscreateplotcyclelist{my colors}{%
solid, color=black\\%
densely dotted, color=blue\\%
densely dashed, color=red\\%
}
\makeatother
%% Environment for comments: Set the boolean to false to produce a comment-free version.
\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
{ \newcommand{\mynote}[3]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\small$\blacktriangleright$\textsf{\textit{\color{#3}{#2}}}$\blacktriangleleft$}}}
{ \newcommand{\mynote}[3]{}}
% One command per author:
\newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}

\newcommand{\epto}{\textsc{EpTO}\xspace}
\newcommand{\jgroups}{\textsc{JGroups}\xspace}

\newif\ifjmcs
% Change false to true to have the JMCS cover page
\jmcstrue

\pgfplotsset{
	compat=1.14,
	tick label style={font=\small},
	label style={font=\small},
	legend style={font=\small},
	/pgfplots/short line legend/.style={
		legend image code/.code={
			\draw[mark repeat=2,mark phase=2,##1]
			plot coordinates {
				(0cm,0cm)
				(0.2cm,0cm)
				(0.4cm,0cm)
			};%
		}
	},
	/pgfplots/xtick pos=left,
}

\def\ieeetitle{EpTO TO DEFINE}
\def\ieeeauthor{X,X,X}

\def\thesistitle{EpTO TO DEFINE}
\def\thesissubtitle{X}
\def\thesisauthor{Jocelyn Thode}

\author{\IEEEauthorblockN{\ieeeauthor}
	\IEEEauthorblockA{University of Fribourg, Switzerland\\
		\href{mailto:first.last@unifr.ch}{first.last@unifr.ch}}
}
\title{\ieeetitle}

\addbibresource{references.bib}
\renewcommand*{\bibfont}{\small}

\hypersetup{
	pdftitle=\thesistitle,
	pdfauthor=\thesisauthor
}

% \noautocite{*}

\begin{document}

\ifjmcs
\begin{titlepage}
	\begin{otherlanguage}{australian}
		\begin{center}
			\begin{figure}[t]
				\center{\includegraphics[scale=0.2]{logos/MSc_quer.png}}
				\vspace{0.4in}
			\end{figure}

			{\bfseries\Huge \thesistitle \par
				\Large \vspace{0.1in} \thesissubtitle \par}

			\vspace{0.3in}
			\LARGE{\textbf{Master Thesis} \\}
			\vspace{0.4in}

			{\Large \thesisauthor}

			\vspace{0.3in}
			{\Large Université de Fribourg \par}
			\vfill
			{\Large \today \par}

			\vspace{0.9in}

			% === Logos ==============================================
			\begin{figure}[htp]
				\centering
				\subfloat{\includegraphics[scale=0.60]{logos/UNI_Bern.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.54]{logos/UNI_Neuchatel.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.81]{logos/UNI_Fribourg.pdf}}
			\end{figure}
			% === // Logos ===========================================

		\end{center}
	\end{otherlanguage}
\end{titlepage}

\pagebreak
\fi

%	\maketitle

\input{abstract.tex}
\input{introduction.tex}
\input{ordering.tex}
\input{definitions.tex}
\input{framework.tex}
\input{evaluation.tex}
\input{future-work.tex}
\input{conclusion.tex}

\printbibliography
\end{document}

### 2016-12-22 report/old_report/paper.tex
Add EpTO figures
Start working on report

--- 2016-08-01 report/paper.tex
@@
 {\typeout{There was an error patching biblatex-ieee (specifically, ieee.bbx's @online driver)}}
+\newdimen\tempcolwidth \tempcolwidth0pt
+\newwrite\widthfile
+\AtBeginDocument{\checkcolwidth}
+\newcommand\checkcolwidth{%
+	\InputIfFileExists{\jobname.width}{}{}%
+	\ifdim\columnwidth=\tempcolwidth\else
+	\openout\widthfile=\jobname.width
+	\write\widthfile{\tempcolwidth=\the\columnwidth}%
+	\closeout\widthfile
+	\fi
+}
+
+\newif\ifjmcs
+% Change false to true to have the JMCS cover page
+\jmcsfalse
+
+\pgfplotsset{
+	compat=1.12,
+	tick label style={font=\small},
+	label style={font=\small},
+	legend style={font=\small},
+	/pgfplots/short line legend/.style={
+		legend image code/.code={
+			\draw[mark repeat=2,mark phase=2,##1]
+			plot coordinates {
+				(0cm,0cm)
+				(0.2cm,0cm)
+				(0.4cm,0cm)
+			};%
+		}
+	},
+	/pgfplots/xtick pos=left,
+}
+
+\newif\ifjmcs
+% Change false to true to have the JMCS cover page
+\jmcsfalse
+
+def\ieeetitle{Have a Seat on the ErasureBench:\\Easy Evaluation of Erasure Coding Libraries\\ for Distributed Storage Systems}
+
+%\def\ieeetitle{ErasureBench: an Experimental Testbed to Compare\\ Locally Repairable and Erasure Coding Libraries}
+
+\def\ieeeauthor{Sébastien Vaucher, Hugues Mercier, Valerio Schiavoni}
+
+\def\thesistitle{ErasureBench}
+\def\thesissubtitle{An experimental testbed to evaluate erasure-coding libraries}
+\def\thesisauthor{Sébastien Vaucher}
+
+\author{\IEEEauthorblockN{\ieeeauthor}
+	\IEEEauthorblockA{University of Neuchâtel, Switzerland\\
+		\href{mailto:first.last@unine.ch}{first.last@unine.ch}}
+}
+\title{\ieeetitle}
+
+\addbibresource{references.bib}
+\renewcommand*{\bibfont}{\small}
+
+\hypersetup{
+	pdftitle=\thesistitle,
+	pdfauthor=\thesisauthor
+}
 
@@
 \begin{abstract}
-Total ordering has always been of interest in large-scale distributed systems.
 EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
@@
 \section{Introduction}
+
 Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.

### 2016-12-22 report/ordering.tex
Start working on report

--- 2016-06-16 report/paper.tex
@@
-\documentclass[10pt,conference,a4paper]{IEEEtran}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-% Correct date format in references but with american hyphenation, quotes, ...
-% Trick from http://tex.stackexchange.com/a/129209
-\usepackage[australian,american]{babel}
-
-% *** GRAPHICS RELATED PACKAGES ***
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-\else
-  \usepackage[dvips]{graphicx}
-\fi
-
-% *** MATH PACKAGES ***
-\usepackage{amssymb}
-\usepackage[cmex10]{amsmath}
-
-% Figure captions in small font
-\makeatletter
-\let\MYcaption\@makecaption
-\makeatother
-\usepackage[font=footnotesize]{subcaption}
-\makeatletter
-\let\@makecaption\MYcaption
-\makeatother
-
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.9}
-%TODO minted ?
-\usepackage[binary-units]{siunitx}
-\usepackage{booktabs}
-\usepackage{pifont}
-\usepackage{tikz}
-\usepackage{csquotes}
-\usepackage[backend=biber,style=ieee,minbibnames=1,maxbibnames=99]{biblatex}
-\usepackage{xpatch}
-\usepackage{url}
-\usepackage[backgroundcolor=lightgray]{todonotes}
-\usepackage[hidelinks]{hyperref}
-
-\bibliography{paper}
-% Small font for references
-\renewcommand*{\bibfont}{\small}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\def\mytitle{Practical Implementation of EpTO: An Epidemic Total Order Algorithm}
-\title{\mytitle}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\def\jocelyn{Jocelyn Thode}
-\def\ehsan{Ehsan Farhadi}
-\author{
-	\IEEEauthorblockN{\jocelyn}
-	\IEEEauthorblockA{Université de Fribourg\\
-	Fribourg, Switzerland\\
-	\href{mailto:jocelyn.thode@unifr.ch}{jocelyn.thode@unifr.ch}
-	}
-	\and
-	\IEEEauthorblockN{\ehsan}
-	\IEEEauthorblockA{Université de Neuchâtel\\
-		Neuchâtel, Switzerland\\
-		\href{mailto:ehsan.farhadi@unine.ch}{ehsan.farhadi@unine.ch}
-	}
-
-}
-
-% Set PDF file properties
-\hypersetup{
-	pdftitle=\mytitle,
-}
-
-% Tick marks
-\newcommand{\cmark}{\ding{51}}
-\newcommand{\xmark}{\ding{55}}
-
-% Columns balancing
-\renewbibmacro{finentry}{%
-  \iffieldequalstr{entrykey}{jujjuri2010virtfs}%<- key after which you want the break
-   {\finentry\newpage}
-   {\finentry}}
-
-\xpatchbibdriver{online}
-{\printtext[parens]{\usebibmacro{date}}}
-{\iffieldundef{year}
-	{}
-	{\printtext[parens]{\usebibmacro{date}}}}
-{}
-{\typeout{There was an error patching biblatex-ieee (specifically, ieee.bbx's @online driver)}}
-
-\begin{document}
-\graphicspath{{figures/}}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-Total ordering has always been of interest in large-scale distributed systems.
-EpTO is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions.
-We then compare it to a deterministic total order algorithm named JGroups.
-\end{abstract}
-\section{Introduction}
-Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
-One of the recently designed algorithms on this topic is EpTO \autocite{matos2015epto}.
-EpTO is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, EpTO is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.
-\par 
-There are many other algorithms for disseminating and ordering events in a distributed system.
-There are some deterministic algorithms, which guarantee total order, agreement or other strong properties.
-Unfortunately, these types of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
-\par
-The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the system.
-This causes a massive amount of network traffic and overhead on the system.
-Moreover, an agreement feature for an asynchronous system requires to
-explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability.
-\par
-As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
-These algorithms guarantee the dissemination of events in the system with high probability.
-This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable.
-As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
-\par
-This is where EpTO comes into light. EpTO, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
-EpTO consists of two distinct parts.
-The first part is probabilistic dissemination.
-EpTO guarantees that all peers will receive an event with arbitrarily high probability.
-The second part of EpTO is deterministic ordering.
-Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the event's broadcaster id.
-\par
-To model the first part, EpTO is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
-The balls-and-bins problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls.
-In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/BnB.jpeg}
-\caption{Balls-and-Bins \autocite{bnb}.}
-\label{fig:balls-and-bins}
-\end{figure}
-\par
-Using the balls-and-bins approach we model processes as bins and events as balls and calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability.
-With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Thanks to these approaches, EpTO becomes highly scalable and resilient, while still providing total order.
-
-\subsection{Contribution}
-Until now, the creators of EpTO have only tested this algorithm in a simulated environment.
-In this work, we implement EpTO in pure Kotlin using a modified version of the NeEM library \autocite{neem} and show that EpTO is suitable for real-world large-scale distributed systems.
-The modification to NeEM includes a dumbed-down gossip dissemination as seen in \autocite{matos2015epto}.
-We then evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups  \autocite{jgroups} in stable (no-churn) systems.
-These comparisons help us verify if EpTO is actually performing as expected in a real-world scenario.
-\par
-In our experiments we assume each peer knows every other peers.
-We implement EpTO using this assumption and test it against JGroups.
-The details concerning the methodology are explained in subsection \ref{sub:metho}.
-In future work, we want to test EpTO when the membership is not entirely known.
-
 \section{Ordering Algorithms}
+\label{sec:ordering}
+\jt{I'm not sure we still need this section}
 Distributed systems, like centralized systems, need to preserve the temporal order of events produced by concurrent processes in the system.
-When there are separated processes that can only communicate through messages, you can’t easily order these messages.
+When there are separated processes that can only communicate through messages, you cannot easily order these messages.
 Therefore we need ordering algorithms to overcome this problem.
@@
 We then have to use total order, otherwise we can end up in an inconsistent state.
-
-\section{EpTO}
-\begin{figure}
-	\includegraphics[width=\linewidth]{figures/architecture.pdf}
-	\caption{EpTO architecture \autocite{matos2015epto}.}
-	\label{fig:epto-architecture}
-\end{figure}
-\subsection{Definitions}
-A process or peer is defined as an actor in our system running the application that needs total order.
-Each process will communicate with other processes in the distributed system, exchange events, and order them together.
-\par
-An event is defined as data sent at a given time by a peer.
-For example, we could imagine a system where each process publishes some data to other peers.
-The moment where we publish data combined with the data is called an event.
-\par
-A ball is a set of events bundled together and sent as one package. We use balls in EpTO to reduce network traffic and make it scalable in terms of processes and events.
-
-We define EpTO scaling well as it was defined in  \autocite{matos2015epto}:  ``The number
-of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver an event stays low as well.
-\par
-Since EpTO uses a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
-In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by EpTO's deterministic ordering algorithm, thus the total order property is preserved.
-\subsection{Dissemination Component}
-The Dissemination component bridges EpTO with the rest of the network.
-As we can see in Figure \ref*{fig:epto-architecture}, it receives balls, opens them, passes them to the Ordering component, and stores their events.
-Then, in the next round, it sends all the events received in the previous round to K other processes, where K denotes the gossip fan-out.
-
-\par
-When an application wants to publish an event, it will broadcast this event to the Dissemination component.
-
-\subsection{Ordering Component}
-The Ordering component is responsible for ordering events before delivering them to the application.
-To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which were delivered.
-\par
-In brief, the Ordering component first increments the timestamp of all the events which have been received in previous rounds to indicate the start of a new round.
-Then, it processes new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event).
-This is done to prevent delivering duplicate events.
-The remaining events in the received ball are added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
-\subsection{Stability Oracle}
-The Stability oracle is the component that outputs timestamps.
-It will increment its local clock every round as well as synchronize itself using timestamps of newly received events, to make sure the local clock does not drift too much.
-\par
-This Stability oracle offers an API to the Ordering component letting it know when an event is mature enough to be delivered to the application.
-\par
-As this component is local and only corrects itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
-\section{Performance Comparisons}
-In the following subsections, we will present the results obtained for the different tests ran and explain their meanings.
-\subsection{Methodology}
-\label{sub:metho}
-Our tests are run on a cluster provided by the University of Neuchâtel.
-For setting up our testbed we use OpenNebula \autocite{opennebula} to manage our cluster of virtual machines.
-We were planning on using Docker \autocite{docker}, more specifically Docker-compose and docker-swarm in order to orchestrate and manage our cluster of peers on the OpenNebula but due to some issues with docker-swarm, we decided to deploy our application directly on the OpenNebula virtual machines, and control it using parallel ssh sessions.
-\par
-We deployed our application on a small network of 4 virtual machines and we distribute EpTO and JGroups peers equally over the network.
-\par
-To measure the throughput, we transmit a certain number of events by certain number of peers in a defined period of time, and then we measure the number of packets transmitted over the network.
-\par
-To measure the latency of both protocols, we measure the average time difference on a peer between the broadcast of an event and the time it delivers this same event.
-This way we can have an estimation of the approximate delay, and compare JGroups and EpTO.
-We execute EpTO and JGroups with 60, 100, 120 and 160 peers each peer broadcasting 12 events for 12 seconds (one event per second).
-\subsection{Peer membership known}
-As we measured and compared the latency of EpTO and JGroups we realize that JGroups is performing better than EpTO with smaller number of peers.
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/Delay.png}
-\caption{Approximate delay of delivering each event.}
-\label{fig:delay}
-\end{figure}
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/total_time.png}
-\caption{total run time of EpTO peers in order to deliver all events.}
-\label{fig:total_time}
-\end{figure}
-When we increase the number of peers, our results show that JGroups is not as scalable as EpTO, but considering our limited resources we could not fully observe the difference between JGroup and EpTO on this matter.
-Although we should bear in mind that in this experiment EpTO has a relatively large delay due to ordering of the events without a central node (as in JGroups).
-But despite this delay, we expected EpTO to perform much better.
-We infer the problem might be related more to the way the experiments were run and not directly to EpTO.
-This needs further testing and will be done during the master thesis on this subject.
-\par
-This issue is also present when measuring the total time of execution of EpTO for delivering a certain number of events.
-\par
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/throughput.png}
-\caption{number of events transmitted packets over the network in 10 minutes.}
-\label{fig:throughput}
-\end{figure}
-For the throughput of both protocols we used a network sniffing tool, TCPdump.
-TCPdump is a tool for observing the network traffic.
-Using this tool we count the number of packets (TCP packets for EpTO and UDP packets for JGroups) transmitted over the network for 10 minutes.
-Again, the number of packets sent by EpTO peers are much higher than JGroups.
-This issue can be explain due to the differences between TCP and UDP.
-TCP uses much more packets and data over the network for a connection to secure and guarantee the delivery of data.
-that's why in a perfect and lossless network, JGroups has better performance from perspective of network throughput.
-We expect the number of packets sent by EpTO to drastically diminish once we have made the switch to UDP.
-\section{Future Work}
-One of the students working on this project will do his master thesis on this project.
-
-Future work will include testing EpTO while having unknown peer membership.
-To account for that, we will need to modify NeEM to support the use of a Peer Sampling Service.
-We plan on implementing the CYCLON Peer Sampling Service \autocite{Voulgaris2005}.
-For now, we plan on using a tracker as in torrents to give each peer its initial view.
-This tracker is already implemented and works as a python web server that gets all alive peers and randomly select $K$ peers and sends this result to the EpTO peer.
-
-Our first implementation of EpTO focuses on correctness rather than fastness.
-Later work should optimize EpTO in regards to concurrency and network latency.
-To address the second point, we want to modify NeEM again to use UDP instead of TCP, to reduce latency and overhead in the network.
-\section{Conclusion}
-In conclusion, EpTO was implemented using Kotlin with a focus on correctness instead of fastness.
-There are still many optimizations left to be done on the code which will be implemented during the master thesis following this project.
-\par
-Tests were ran evaluating  the average delay for an event to be delivered, the total time of execution and the number of packets transmitted over the network.
-These tests showed that EpTO was not behaving as we expected.
-Unfortunately as we encountered many problems during the tests we could not test the algorithm the way we wanted and these results are only a start and will be more conclusive in the future.
-\par
-At the moment, the tests must be run somewhat manually. We want to use Docker and Docker-swarm in the future to enhance the reproducibility of the tests on any cluster.
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-This research was under supervision and help of two of the creators of EpTO, Dr.
-Hugues Mercier and Dr.
-Miguel Matos.
-We would like to thank them both for the help, insight and expertise they provided during this project.
-\par
-We would also like to thank Valerio Schiavoni for providing access to the cluster of University of Neuchâtel as well as Sebastien Vaucher for the help with setting up Docker, Docker-Compose and Docker Swarm.
-
-% references section, with correct date format
-\begin{otherlanguage}{australian}
-\printbibliography
-\end{otherlanguage}
-
-% that's all folks
-\end{document}

### 2016-12-22 projects/EpTOTester/results/epto-tester.py
Fix small errors

@@ -234,9 +234,9 @@ logging.info("K={:d} / TTL={:d} / Delta={:d}ms".format(k, ttl, delta))
 logging.info("-------------------------------------------")
 logging.info("Least time to deliver in total : {:d} ms".format(mininum))
 logging.info("Most time to deliver in total : {:d} ms".format(maximum))
-logging.info("Average time to deliver per peer in total: {:d} ms".format(average))
+logging.info("Average time to deliver per peer in total: {:f} ms".format(average))
 logging.info("Population std fo the time to deliver: {:f} ms" .format(statistics.pstdev(durations, average)))
-logging.info("Average global time to deliver on all peers per experiment: {:d} ms".format(global_average))
+logging.info("Average global time to deliver on all peers per experiment: {:f} ms".format(global_average))
 logging.info("Population std fo the time to deliver: {:f} ms".format(statistics.pstdev(global_times, global_average)))
 logging.info("-------------------------------------------")
 balls_sent_sum = sum(balls_sent)
@@ -255,8 +255,8 @@ stats_events_sent = []
 for experiment_nb, the_stats in stats.items():
     sent_sum = sum(messages_sent[experiment_nb])
     stats_events_sent.append(sent_sum)
-    logging.info("Experiment {:d}:" % experiment_nb)
-    logging.info("Total events sent: {:d}" % sent_sum)
+    logging.info("Experiment {:d}:".format(experiment_nb))
+    logging.info("Total events sent: {:d}".format(sent_sum))
     logging.info("Total events received on average: {:f}".format(
         sum(messages_received[experiment_nb]) / len(messages_received[experiment_nb])))
     logging.info("-----------")

### 2016-12-22 projects/JGroupsTester/results/jgroups-tester.py

@@ -191,9 +191,9 @@ logging.info("JGroups")
 logging.info("-------------------------------------------")
 logging.info("Least time to deliver in total : {:d} ms".format(mininum))
 logging.info("Most time to deliver in total : {:d} ms".format(maximum))
-logging.info("Average time to deliver per peer in total: {:d} ms".format(average))
+logging.info("Average time to deliver per peer in total: {:f} ms".format(average))
 logging.info("Population std fo the time to deliver: {:f} ms" .format(statistics.pstdev(durations, average)))
-logging.info("Average global time to deliver on all peers per experiment: {:d} ms".format(global_average))
+logging.info("Average global time to deliver on all peers per experiment: {:f} ms".format(global_average))
 logging.info("Population std fo the time to deliver: {:f} ms".format(statistics.pstdev(global_times, global_average)))
 logging.info("-------------------------------------------")
 
@@ -203,10 +203,10 @@ for experiment_nb, the_stats in stats.items():
     sent_sum = sum(messages_sent[experiment_nb])
     received_sum = sum(messages_received[experiment_nb])
     stats_events_sent.append(sent_sum)
-    logging.info("Experiment {:d}:" % experiment_nb)
-    logging.info("Total events sent: {:d}" % sent_sum)
+    logging.info("Experiment {:d}:".format(experiment_nb))
+    logging.info("Total events sent: {:d}".format(sent_sum))
     logging.info("Total events received on average: {:f}".format(
-        sum(received_sum) / len(messages_received[experiment_nb])))
+        received_sum / len(messages_received[experiment_nb])))
     logging.info("-----------")
     ratios = [(msg_received / sent_sum) for msg_received in messages_received[experiment_nb]]
     logging.info("Best ratio events received/sent: {:.10g}".format(max(ratios)))

### 2016-12-23 report/conclusion.tex
Continue work on thesis

\section{Conclusion}
\label{sec:conclusion}
\jt{Do later}

### 2016-12-23 report/definitions.tex
Continue work on thesis

@@
 \section{Definitions}
+\label{sec:definitions}
 A process or peer is defined as an actor in our system running the application that needs total order.
@@
 \par
-We write the cluster parameters as $(p,e)$, where $p$ designates the total number of peers and $e$ designates the global event throughput.
+We write the cluster parameters as $(p,e)$, where $p$ designates the total number of peers and $e$ designates the global event throughput per second.
 We tested \epto and \jgroups with three different cluster parameters:
 \begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
-	\item[\textbf{$(100,100)$}:] 50 peers with a global event throughput of 50 events per seconds.
-	\item[\textbf{$(100,100)$}:] 50 peers with a global event throughput of 100 events per seconds.
-	\item[\textbf{$(100,100)$}:] 100 peers with a global event throughput of 50 events per seconds. This parameter was also used for all synthetic churns.
+	\item[\textbf{$(50,50)$}:] 50 peers with a global event throughput of 50 events per second.
+	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
+	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. This parameter was also used for all synthetic churns.
 \end{description}
-\par

### 2016-12-23 report/evaluation.tex
Continue work on thesis

@@
 \label{sec:evaluation}
+\jt{specify somewhere the problems encountered with JGroups}
 \pgfplotstableread[col sep=comma]{csv-data/average-bytes-sent-recv.csv}\tableaverage
@@
 \pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
-In this section we present the results of the benchmarks.
-\subsection{Testbed}
-\jt{Write about benchmarks}
-\subsection{Parameter}
-\epto uses a constant $c$ to set the probability of a hole appearing.
+In this section we present our results.
+In the following benchmarks \epto uses a constant $c$ to set the probability of a hole appearing.
 We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions.
@@
 \par 
-We used specific settings recommended for a big cluster available in the \jgroups library.
+We used specific settings recommended by \jgroups for a big cluster.
+\par
+Every \jgroups test run with churn is run once killing the coordinator and once not killing it. We use the following nomenclature to differentiate both tests:
+\begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
+	\item[\textbf{\jgroups-coord}:] The coordinator is killed once
+	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive
+\end{description}
+\par
+\textbf{Testbed.} Our cluster is comprised of... \jt{Ask Valerio+Seb and speak about the different software needed}
+\par
+\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}.
+When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}.
+The benchmarks following a real trace are run 5 times.
+The trace is speed-up 2x, which means we follow \SI{2}{\hour} in \SI{1}{\hour}.
+Every benchmark run with churn is run with the $(100,50)$ parameters.
 \subsection{Bandwidth}
-\jt{\epto was run with $c=4$ to prevent any holes and a $\delta$ period of 100ms.
-Churn was run for 17minutes.
-3min at the end with no churn, to see stabilization.
-Churn starts 30seconds after protocol}
-\jt{We should define acronyms for the different benchmark variants}
-
-\begin{figure}[htp]
+\begin{figure}[hpt]
 	\centering
@@
 
-\begin{figure}[htp]
+\begin{figure}[hpt]
 	\centering
@@
 
-\newpage
-
 \subsection{Total GigaBytes sent/received}
-\begin{table}[htp]
+\begin{table*}[hpt]
 	\centering
@@
 	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
-	\begin{tabular}{lSSSS}
+	\begin{tabular}{llSSS}
 		\toprule
@@
 		\cmidrule{3-5}
-		{Protocol}&& {50P50E} & {50P50E} & {100P50E} \\
+		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
 		\midrule
@@
 	\label{table:total-bandwidth} 
-\end{table}
+\end{table*}
 
@@
 Here we are well within this ratio.
-\begin{table}[htp]
+\begin{table*}[hpt]
 	\centering
@@
 		\midrule
-		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.55(002) & 1.85(002)\\
-									 &{Sending}& 1.51(002) & 1.79(002)\\
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.55(002) & 1.85(002)\\&{Sending}& 1.51(002) & 1.79(002)\\
 		\midrule
-		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.54(001) & 1.83(002)\\
-									   &{Sending}& 1.49(001) & 1.78(002)\\
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.54(001) & 1.83(002)\\&{Sending}& 1.49(001) & 1.78(002)\\
 		\bottomrule
@@
 	\label{table:total-bandwidth-churn} 
-\end{table}
+\end{table*}
 
@@
 \label{sub:local-times}
-\jt{Having a table for each percentile figure might be a good idea to put numbers for key percentiles (min,50th,max) for example}
-\begin{figure*}[htp]
+\begin{figure*}[hpt]
 	\centering
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
-\newpage
 \subsection{Global Times}
-\begin{figure*}[htp]
+\begin{figure*}[hpt]
 	\centering
@@
 Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times}.
-\newpage
 \subsection{Local Dissemination stretch}
-\begin{figure*}[htp]
+\begin{figure*}[hpt]
 	\centering
@@
 The median degrade a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
-\newpage
 \subsection{Events sent}
 \jt{Representing as a table might be better}
-\begin{table}[htp]
+\begin{table}[hpt]
 	\centering
@@
 In \autoref{table:total-events} we can see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
-\begin{table}[htp]
+\begin{table}[hpt]
 	\centering
@@
 When killing and adding nodes, JGroups seems to deliver a smaller amount of nodes, however it does not look significant \jt{I didn't run any statistical analysis}
-\section{Real Trace}
-\subsection{Bandwidth}
-
-\begin{figure}[htp]
+\subsection{Real Trace}
+\subsubsection{Bandwidth}
+\begin{figure}[hpt]
 	\centering
@@
 \end{figure}
-
-\newpage
-\subsection{Total Bytes sent/received}
-\begin{table}[htp]
+\subsubsection{Total Bytes sent/received}
+\begin{table}[hpt]
 	\centering
@@
 
-\subsection{Local Times}
-\label{sub:local-times}
-\begin{figure}[htp]
+\subsubsection{Local Times}
+\begin{figure}[hpt]
 	\centering
@@
 
-\newpage
-\subsection{Global Times}
+\subsubsection{Global Times}
 
-\begin{figure}[htp]
+\begin{figure}[hpt]
 	\centering
@@
 \end{figure}
-\newpage
-\subsection{Local Dissemination stretch}
+\subsubsection{Local Dissemination stretch}
 
-\begin{figure}[htp]
+\begin{figure}[hpt]
 	\centering
@@
 \end{figure}
-\newpage
-\subsection{Events sent}
-\begin{table}[htp]
+\subsubsection{Events sent}
+\begin{table}[hpt]
 	\centering
 	\caption{Total events sent during a real trace}
-\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 6, table-align-uncertainty}
 \begin{tabular}{lS}
 	\toprule
-	Protocol & \\
+	Protocol &\\
 	\midrule
-	\epto & 165844.2(2102) \\
-	\jgroups-coord & 166183.0(13681) \\
-	\jgroups-nocoord & 166585.8(8249) \\
+	\epto & 165844.2(2102)\\
+	\jgroups-coord & 166183.0(13681)\\
+	\jgroups-nocoord & 166585.8(8249)\\
 	\bottomrule

### 2016-12-23 report/figures/architecture.tex
Continue work on thesis

	\begin{tikzpicture}[transform shape,scale=0.6]
\node[draw, text width=2.1cm, text centered] (bench) at (2.1, 2.6) {Benchmarks runner};
\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python script}};
\draw[dotted,red,thick] (0.8,2) rectangle (3.4,3.4);
%
\draw[loosely dashed,black] (3.7,4.3) rectangle (14.8,-0.8);
\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay:} \texttt{epto containers communicate between each other using UDP}};
%
\node[draw] (tck) at (12.4, 2.5) {Tracker};
\draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto-tracker:}};
\draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);
\draw[dotted,red,thick] (11.6,2) rectangle (14.4,3.25);
%
\node[draw] (app1) at (7.1, 2.5) {EpTOTester 1};
\draw (5.7,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (5.8,3.25) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (5.7,2) rectangle (8.5,3.65);
\draw[dotted,red,thick] (5.8,2.1) rectangle (8.4,3.25);
%
\node[draw] (app2) at (5.4, 0) {EpTOTester 2};
\draw (4,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (4,-0.5) rectangle (6.8,1.15);
\draw[dotted,red,thick] (4.1,-0.4) rectangle (6.7,.75);
%
\node[draw] (appn) at (9.4, 0) {EpTOTester N};
\draw (8,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (8,-0.5) rectangle (10.8,1.15);
\draw[dotted,red,thick] (8.1,-0.4) rectangle (10.7,.75);
%
\draw[<->,thick,>=stealth] (app1.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (app2.east) to[out=50,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (appn.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
%
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app1.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app2.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (appn.west);
\end{tikzpicture}

### 2016-12-23 report/figures/no-churn/global-times.tex
Continue work on thesis

@@
 legend cell align=left,
-legend style={at={(1.6,1.4)},anchor=north, font=\small, draw=none},
+legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none},
 cycle list name=my colors,

### 2016-12-23 report/figures/no-churn/local-delta.tex
Continue work on thesis

@@
 every axis plot/.append style={very thick},
-width=\linewidth / 2.5,
+width=\linewidth / 3,
+height=5cm,
 grid=major,
@@
 legend cell align=left,
-legend style={at={(1.7,1.7)},anchor=north, font=\small, draw=none},
+legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none},
 cycle list name=my colors,

### 2016-12-23 report/figures/real-churn/bandwidth-figure-real-churn.tex
Continue work on thesis

@@
 legend cell align=left,
-legend style={at={(0.2,2)},anchor=north, font=\small, draw=none},]
-\nextgroupplot[height=2.2cm, 
+legend style={at={(0.4,1.9)},anchor=north, font=\small, draw=none},]
+\nextgroupplot[height=2.3cm, 
 ylabel={Nodes}, 
-ymin=80,ymax=120,
-ytick={80,120}, 
-yticklabels={80,...,120},
+ymin=75,ymax=120,
+ytick={80,100,115}, 
+yticklabels={80,100,115},
 xmajorgrids,

### 2016-12-23 report/figures/real-churn/global-times-real-churn.tex
Continue work on thesis

@@
 every axis plot/.append style={very thick},
+height=5cm,
 grid=major,
@@
 legend cell align=left,
-legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
+legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none},
 cycle list name=my colors,

### 2016-12-23 report/framework.tex
Continue work on thesis

@@
 Events broadcasted to the Dissemination component are sent over the network every $\delta$ period.
-Every events received from the network are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
+Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
+They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
 The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie.
@@
 To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system.
-We want to emphasize that the tracker is not obligatory.
-In practice it works well, but using DHTs is certainly a possibility.
+We want to emphasize that the tracker is not required.
+In practice it works well, but using a DHT is certainly a possibility.
 \subsection{Implementation}
-We implemented the data sent as randomly generated UUIDs.
-We implemented our own PSS CYCLON operating on its own port.
-When implementing \jgroups we had to use TCPGOSSIP instead of the traditional MULTICAST option to coordinate peers.
+We implement the data sent as randomly generated UUIDs.
+We implement our own PSS CYCLON operating on its own port.
+When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
-The framework is open-sourced on Github.
+Finally, The framework is open-sourced on Github.
 \subsection{Deployment}
-To deploy \epto effectively we effectively need two different services.
+\begin{figure}[htp]
+	\centering
+	\input{figures/architecture.tex}
+	\vspace{-2mm} 
+	\caption{EpTOTester architecture}
+	\vspace{-2mm} 
+	\label{fig:complete-architecture}
+\end{figure}
+To deploy \epto effectively we need two different docker services.
 One containing the tracker and one containing all \epto replicas.
 This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
-This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal tweaks.
-Every benchmark is launched through a Python script available on the master mode.
-This script handles everything, from starting benchmarks, scaling the cluster during a churn and collecting results.
+This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
+Every benchmark is started through a Python script available on the master mode.
+This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
 All \epto parameters are customizable by using options provided through this script.
@@
 \jgroups deployment is much the same.
-\jt{I should create a figure showcasing the complete architecture}
+The entire EpTOTester architecture is shown in Figure \ref{fig:complete-architecture}.
+\jt{The figure is heavily inspired from the one in ErasureBench.
+How should I indicate this?}
 \subsection{Fault Injection}

### 2016-12-23 report/future-work.tex
Continue work on thesis

\section{Future Work}
\label{sec:future-work}
\jt{Speak about running test on a bigger infrastructure, because the small ammount of nodes does not let us see the full potential of \epto. Then speak about the ongoing work of Push-Pull}

### 2016-12-23 report/introduction.tex
Continue work on thesis

@@
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
-\epto uses a PSS CYCLON to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.
+\epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.
 These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
@@
 Section \ref{sec:ordering} presents the different kind of ordering.
+Section \ref{sec:definitions} defines the terms used in the paer.
 Section \ref{sec:epto} presents \epto and the architecture of our implementation.

### 2016-12-27 report/abstract.tex
WIP report

@@
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.
-We then compare them using a synthetic churn trace and a real one using the same metrics.\jt{Speak about results}
+We then compare them using a synthetic churn trace and a real one using the same metrics.
+Our results shows \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
 \end{abstract}

### 2016-12-27 report/definitions.tex
WIP report

@@
 \par
-A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic and make it scalable in terms of processes and events.
+A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic.
 
-We define \epto scaling well as it was defined in  \autocite{matos2015epto}:  ``The number
+We define \epto scaling well as it is defined in  \autocite{matos2015epto}:  ``The number
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver an event stays low as well.
+in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well.
 \par
-Since \epto uses a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability).
+Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a very low arbitrary probability a peer does not receive an event.
 In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
 \par
-We write the cluster parameters as $(p,e)$, where $p$ designates the total number of peers and $e$ designates the global event throughput per second.
+We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second.
 We tested \epto and \jgroups with three different cluster parameters:
@@
 	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
-	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. This parameter was also used for all synthetic churns.
+	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
 \end{description}

### 2016-12-27 report/evaluation.tex
WIP report

@@
 \jt{specify somewhere the problems encountered with JGroups}
-\pgfplotstableread[col sep=comma]{csv-data/average-bytes-sent-recv.csv}\tableaverage
+\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
+\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
 \pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
@@
 \begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
-	\item[\textbf{\jgroups-coord}:] The coordinator is killed once
-	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive
+	\item[\textbf{\jgroups-coord}:] The coordinator is killed once.
+	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive throughout the experiment.
 \end{description}
 \par
-\textbf{Testbed.} Our cluster is comprised of... \jt{Ask Valerio+Seb and speak about the different software needed}
+\textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well.
+Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\giga\bit\per\second} switched network.
+Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM.
+We deploy 12 virtual machines on these hosts.
+Each virtual machine has acces to 4 CPUs and 4 VCPUs as well as \SI{7}{\giga\byte} of RAM.
+We use KVM as our hypervisor.
+Each VM uses Debian as its O/S.
+EpTOTester is packaged as docker images.
+We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services.
+Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
 \par
@@
 The benchmarks following a real trace are run 5 times.
-The trace is speed-up 2x, which means we follow \SI{2}{\hour} in \SI{1}{\hour}.
+The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}.
 Every benchmark run with churn is run with the $(100,50)$ parameters.
+When we have a global even throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput.
+For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
+\par
+We compared \epto and \jgroups based on their bandwidth, the time to deliver every expected events, their local dissemination stretch and finally the number of events sent.
 \subsection{Bandwidth}
+The initial bandwidth peaks observed for \epto are due to the PSS initialization.
+When a node starts, it received an initial view and then quickly run the PSS four times to obtain a random enough view.
 \begin{figure}[hpt]
@@
 	\vspace{-2mm} 
-	\caption{Throughput percentiles of a node during an experiment}
+	\caption{Bandwidth percentiles of a node during an experiment}
 	\vspace{-2mm} 
@@
 \end{figure}
-In \autoref{fig:bandwidth} we can see \epto has a worse baseline compared to \jgroups.
+In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups.
 It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}.
@@
 
-Comparing \epto and \jgroups in terms of bandwidth when we increase the number of events sent per second, we can see the bandwidth doubling in both cases.In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups is clearly at an advantage.
+Comparing \epto and \jgroups in terms of bandwidth when we increase the number of events sent per second, we can see the bandwidth doubling in both cases.
+In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups is clearly at an advantage.
 Since \epto has a worse baseline we will reach the maximum bandwidth possible much quicker when increasing the event throughput.
 
-Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto is performing better than \jgroups.
-Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it by 50\% or less.
-\jt{logarithmic I think as was expected}.
+Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups.
+Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}.
 Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum bandwidth.
@@
 	\vspace{-2mm} 
-	\caption{Throughput percentiles of a node during an experiment with churn}
+	\caption{Bandwidth percentiles of a node during an experiment with churn}
 	\vspace{-2mm} 
@@
 \subsection{Total GigaBytes sent/received}
+In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.
+\par
 \begin{table*}[hpt]
@@
 		\midrule
-		\multirow{2}{*}{\epto}&{Receive}& 10.85(016) & 22.31(039) & 26.01(028) \\
-						    &{Sending}& 10.85(016) & 22.31(039) & 26.01(028)\\
+		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
+						    &{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
 		\midrule
-		\multirow{2}{*}{\jgroups}&{Receive}& 0.81(003) & 1.48(001) & 2.00(001)\\
-							   &{Sending}& 0.80(003) & 1.47(001) & 1.95(001)\\
+		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
+							   &{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
 		\bottomrule
@@
 		\cmidrule{3-4}
-		{Protocol}&& \{1kill\}/minute & \{1kill,1add\}/minute \\
+		{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
 		\midrule
@@
 		\midrule
-		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.55(002) & 1.85(002)\\&{Sending}& 1.51(002) & 1.79(002)\\
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
 		\midrule
-		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.54(001) & 1.83(002)\\&{Sending}& 1.49(001) & 1.78(002)\\
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
 		\bottomrule
@@
 %\end{figure}
-In \autoref{table:total-bandwidth-churn} we can see that \jgroups total bandwidth usage is smaller when there is churn.
-One hypothesis for this is that a JGroups replica takes a longer time to start up thus the overall benchmark has a longer time with less than 100 replicas.
+In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn.
+One hypothesis for this is that a JGroups replica takes a longer time to start up compared to stopping a replica.
+Therefore the overall benchmark has a longer time with less than 100 replicas.
 We also do not see a difference whether we kill the coordinator or not.
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
+We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system.
 \subsection{Global Times}
@@
 \end{figure*}
-In \autoref{fig:local-delta}, We can see the percentiles of the local dissemination stretch.
+In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch.
 The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
-
+\par
 \jgroups is usually much faster than \epto in a perfect environment.
-This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minmimal latency.
-The median dissemination stretch of \jgroups is around \SI{7}{\milli\second} where as the median dissemiantion stretch of \epto is around \SI{630}{\milli\second} for $(100,50)$.
-When increasing the number of peers, \jgroups starts to have long delivery times for some outliers.
+This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minimal latency.
+The median dissemination stretch of \jgroups is around \SI{7}{\milli\second} where as the median dissemination stretch of \epto is around \SI{630}{\milli\second} for $(100,50)$.
+When increasing the number of peers, \jgroups starts to have longer delivery times for some outliers.
+A small portion of the local dissemination stretches are really fast (\SI{1}{\milli\second} or lower).
+This happens when the coordinator itself sends an event.
+Since it has to deliver it to himself the local dissemination stretch is intra-process and thus extremely fast.
 
@@
 \end{figure}
-In \autoref{fig:local-delta-churn} We can see a completely different picture.
+In \autoref{fig:local-delta-churn} We see a completely different picture.
 When under churn, the 95th percentile of \jgroups is at \SI{31}{\milli\second} compared to \SI{14}{\milli\second} when there is no churn.
@@
 \subsection{Events sent}
-\jt{Representing as a table might be better}
+The variation observed in the different tests are due to the fact that we run an experiment for a period of time and that in some configurations a randomness decides whether an event is sent or not.
 \begin{table}[hpt]
@@
 	\cmidrule{2-4}
-	Protocol & {50P50E} & {50P50E} & {100P50E} \\
+	Protocol & {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
 	\midrule
@@
 %\end{figure}
-In \autoref{table:total-events} we can see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
+In \autoref{table:total-events} we see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
 \begin{table}[hpt]
@@
 	\cmidrule{2-3}
-	Protocol & \{1kill\}/minute & \{1kill,1add\}/minute \\
+	Protocol & {1 kill/minute} & {1\{kill,add\}/minute} \\
 	\midrule
@@
 In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
-When killing and adding nodes, JGroups seems to deliver a smaller amount of nodes, however it does not look significant \jt{I didn't run any statistical analysis}
+When killing and adding nodes, JGroups seems to deliver a smaller amount of nodes, however it does not look significant enough to draw any conclusion.
+\jt{I didn't run any statistical analysis}
 \subsection{Real Trace}
-\subsubsection{Bandwidth}
 \begin{figure}[hpt]
@@
 \end{figure}
-\subsubsection{Total Bytes sent/received}
+\autoref{fig:bandwidth-real-churn} confirms the results from the synthetic churn.
+\epto is not affected by the churn.
+As before, \jgroups is affected in the same situations as when following a synthetic churn.
+The spikes visible on \jgroups-coord and \jgroups-nocoord show the different view changes occuring.
+\par
 \begin{table}[hpt]
@@
 		\midrule
-		\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.61(008)\\
-		&{Sending}& 5.45(008)\\
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.56(008)\\
+		&{Sending}& 5.40(008)\\
 		\midrule
-		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.63(005)\\
-		&{Sending}& 5.48(005)\\
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.58(005)\\
+		&{Sending}& 5.43(005)\\
 		\bottomrule
@@
 \end{table}
+\autoref{table:total-bandwidth-real-churn} also confirms our earlier results.
+We run the experiment for approximately 3 times the synthetic period and we get approximately 3 times more \SI{}{\giga\byte} used.
+\par
 %\begin{figure}[h]
@@
 %\end{figure}
-
-\subsubsection{Local Times}
 \begin{figure}[hpt]
@@
 \end{figure}
-
-\subsubsection{Global Times}
-
 \begin{figure}[hpt]
@@
 \end{figure}
-\subsubsection{Local Dissemination stretch}
-
+The local and global dissemination times shown in respectively \autoref{fig:local-times-real-churn} and \autoref{fig:global-times-real-churn} still show \jgroups to outperform \epto in this scenario.
+Again we want to emphasize that \epto would more than probably outperform \jgroups in a bigger size cluster.
+\par
 \begin{figure}[hpt]
@@
 \end{figure}
-\subsubsection{Events sent}
+In \autoref{fig:local-delta-real-churn}, \epto as worse outliers than before and so does \jgroups-coord.
 \begin{table}[hpt]
@@
 \end{table}
+\autoref{table:total-bandwidth-real-churn} is interesting.
+We see that on average \jgroups seems to deliver more events than \epto, but its standard deviation is way higher than that of \epto, suggesting \epto to be more stable and consistent.
 %\begin{figure}[h]
@@
 %\end{figure}
+\subsection{Problems encountered}
+The \epto simulation and theory is assuming we can have balls of infinite size.
+In practice, packet loss happens.
+We must therefore limit the ball size to a practical number.
+We use balls with a maximum size of \SI{1432}{\byte}.
+We choose this number as the default MTU on Ethernet is \SI{1500}{\byte}.
+We then subtract the bytes required for the IPv4 and UDP headers.
+\par
+When running \jgroups we sometimes get failed runs, where either some peers will experience holes or peer will deliver out of order events.
+This only happens when there is churn happening and when the coordinator dies.
+We have not investigated these problems further.
+This points to either \jgroups failing when there is too much churn or a bug present in the SEQUENCER implementation.

### 2016-12-27 report/figures/architecture.tex
WIP report

@@
 \draw[loosely dashed,black] (3.7,4.3) rectangle (14.8,-0.8);
-\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay:} \texttt{epto containers communicate between each other using UDP}};
+\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay}};
 %
@@
 \draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
-\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto-tracker:}};
+\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto-tracker:}};
 \draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);
@@
 %
+\draw[loosely dashed,blue] (4,-0.5) rectangle (10.8,3.65);
+\draw (4,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto}};
+%
 \node[draw] (app1) at (7.1, 2.5) {EpTOTester 1};
-\draw (5.7,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
 \draw (5.8,3.25) node[below right,red,scale=0.8] {\textit{Java program}};
-\draw[loosely dashed,blue] (5.7,2) rectangle (8.5,3.65);
 \draw[dotted,red,thick] (5.8,2.1) rectangle (8.4,3.25);
@@
 \node[draw] (app2) at (5.4, 0) {EpTOTester 2};
-\draw (4,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
-\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
-\draw[loosely dashed,blue] (4,-0.5) rectangle (6.8,1.15);
 \draw[dotted,red,thick] (4.1,-0.4) rectangle (6.7,.75);
@@
 \node[draw] (appn) at (9.4, 0) {EpTOTester N};
-\draw (8,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
-\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
-\draw[loosely dashed,blue] (8,-0.5) rectangle (10.8,1.15);
 \draw[dotted,red,thick] (8.1,-0.4) rectangle (10.7,.75);
@@
 \draw[<->,thick,>=stealth] (app1.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
-\draw[<->,thick,>=stealth] (app2.east) to[out=50,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
+\draw[<->,thick,>=stealth] (app2.east) to[out=55,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
 \draw[<->,thick,>=stealth] (appn.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
 %
-\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app1.west);
-\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app2.west);
-\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (appn.west);
+\draw[->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Docker API calls} (4,1);
+%
+\draw[<->,thick,>=stealth] (app1.south) -- node[sloped,midway,scale=0.6,above]{UDP} (app2.north);
+\draw[<->,thick,>=stealth] (app1.south) -- node[sloped,midway,scale=0.6,above]{UDP} (appn.north);
+\draw[<->,thick,>=stealth] (app2.east) -- node[sloped,midway,scale=0.6,above]{UDP} (appn.west);
+%
+\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
+\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
 \end{tikzpicture}

### 2016-12-27 report/figures/real-churn/bandwidth-figure-real-churn.tex
WIP report

--- 2016-12-22 report/figures/real-churn/bandwidth-figure-real-churn.tex
@@
 legend cell align=left,
-legend style={at={(0.2,2)},anchor=north, font=\small, draw=none},]
-\nextgroupplot[height=2.2cm, 
+legend style={at={(0.6,1.9)},anchor=north, font=\small, draw=none},]
+\nextgroupplot[height=2.3cm, 
 ylabel={Nodes}, 
-ymin=80,ymax=120,
-ytick={80,120}, 
-yticklabels={80,...,120},
+ymin=75,ymax=120,
+ytick={80,100,120}, 
+yticklabels={80,100,120},
 xmajorgrids,
@@
 label style={font=\tiny}]
-\addplot[const plot, color=blue, mark=none] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
+\addplot[const plot, color=blue, mark=none,set layers,
+axis lines = middle,] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
 \nextgroupplot[ymax=4,ytick={0,1,2,3,4},
 yticklabels={0,1,2,3,4}]
-\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaverage} \closedcycle;
+\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
 \legend{0, 0.25, 0.5, 0.75, 1}
@@
 \nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
-\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
+\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
 %
 \nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
-\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
-\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
+\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
+\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
 \end{groupplot}
 %
-\node[anchor=south, rotate=-90] at (group c1r2.east) {EpTO};
-\node[anchor=south, rotate=-90] at (group c1r3.east) {JGroups coord};
-\node[anchor=south, rotate=-90] at (group c1r4.east) {JGroups nocoord};
+\node[anchor=south, rotate=-90] at (group c1r2.east) {\epto};
+\node[anchor=south, rotate=-90] at (group c1r3.east) {\jgroups-coord};
+\node[anchor=south, rotate=-90] at (group c1r4.east) {\jgroups-nocoord};
 \end{tikzpicture}

### 2016-12-27 report/figures/real-churn/global-times-real-churn.tex
WIP report

--- 2016-12-22 report/figures/real-churn/global-times-real-churn.tex
@@
 yticklabels={0,0.2,0.4,0.6,0.8,1},
+xmin=3649,
+xmax=3663,
+xtick={3650,3655,3660},
+xticklabels={3650,3655,3660},
 ylabel=Percentiles,
@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tableglobaltime};
-\legend{EpTO, JGroups-nocoord, JGroups-coord}
+\legend{\epto, \jgroups-nocoord, \jgroups-coord}
 \end{axis}

### 2016-12-27 report/figures/real-churn/local-times-real-churn.tex
WIP report

--- 2016-12-22 report/figures/real-churn/local-times-real-churn.tex
@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaltime};
-\legend{EpTO, JGroups-nocoord, JGroups-coord}
+\legend{\epto, \jgroups-nocoord, \jgroups-coord}
 \end{axis}

### 2016-12-27 report/figures/synthetic-churn/bandwidth-figure-churn.tex
Continue work on thesis
WIP report

@@
-\begin{tikzpicture}[domain=0:1800]
+\begin{tikzpicture}
 \usetikzlibrary{plotmarks}
@@
 	height=4cm,
-	width=\linewidth / 2.5,
+	width=\linewidth / 2,
 	every axis plot post/.append style={
@@
 legend cell align=left,
-legend style={at={(1.25,1.5)},anchor=north, font=\small, draw=none},]
+legend style={at={(1.15,1.5)},anchor=north, font=\small, draw=none},]
 \nextgroupplot[ymax=4,ytick={0,1,2,3,4},
@@
 \node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
-\node[anchor=south, rotate=-90] at (group c2r1.east){EpTO};
-\node[anchor=south, rotate=-90] at (group c2r2.east){JGroups coord};
-\node[anchor=south, rotate=-90] at (group c2r3.east){JGroups nocoord};
+\node[anchor=south, rotate=-90] at (group c2r1.east){\epto};
+\node[anchor=south, rotate=-90] at (group c2r2.east){\jgroups-coord};
+\node[anchor=south, rotate=-90] at (group c2r3.east){\jgroups-nocoord};
 \end{tikzpicture}

### 2016-12-27 report/framework.tex
WIP report

@@
-\section{Framework}
+\section{EpToTester}
 \label{sec:epto}
@@
 \subsection{Architecture}
-Figure \ref{fig:epto-architecture} illustrates the architecture of a single replica.
-An application extending our Application class can broadcast and delivers events.
+\autoref{fig:epto-architecture} illustrates the architecture of a single replica.
+An application extending our Application class can \epto-broadcast and \epto-deliver events.
 Events broadcasted to the Dissemination component are sent over the network every $\delta$ period.
@@
 The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie.
-The network layer communication is done using UDP and a PSS CYCLON to gather a random view of peers.
+The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers.
 To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system.
@@
 We implement our own PSS CYCLON operating on its own port.
+We use \jgroups 3.6.11.
 When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
-Finally, The framework is open-sourced on Github.
+Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
 \subsection{Deployment}
@@
 \end{figure}
-To deploy \epto effectively we need two different docker services.
+To deploy \epto effectively we need two different Docker Swarm services.
 One containing the tracker and one containing all \epto replicas.
+Both of these services use the same network overlay to communicate.
 This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
@@
 \jgroups deployment is much the same.
-The entire EpTOTester architecture is shown in Figure \ref{fig:complete-architecture}.
-\jt{The figure is heavily inspired from the one in ErasureBench.
+The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}.
+\jt{The figure is inspired from the one in ErasureBench.
 How should I indicate this?}
 \subsection{Fault Injection}
-Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}, although the synthetic churn was improved to support adding and removing nodes at the same time.
+Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}.
+The synthetic churn provided by it was improved to support adding and removing nodes at the same time.

### 2016-12-27 report/introduction.tex
WIP report

@@
 \par
-This is where \epto comes into light. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
+This is where \epto shines. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
 \epto consists of two distinct parts.
@@
 The second part of \epto is deterministic ordering.
-Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster id of the events.
+Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events.
 \par
@@
 The next sections are dedicated to present our benchmarking solution and to present the results obtained.
-Section \ref{sec:ordering} presents the different kind of ordering.
-Section \ref{sec:definitions} defines the terms used in the paer.
-Section \ref{sec:epto} presents \epto and the architecture of our implementation.
-Section \ref{sec:evaluation} shows and explains the results obtained.
-We then argue about future work in section \ref{sec:future-work}.
-We finally draw conclusions in section \ref{sec:conclusion}.
+\autoref{sec:ordering} presents the different kinds of ordering.
+\autoref{sec:definitions} defines the terms used in the paper.
+\autoref{sec:epto} presents \epto and the architecture of our implementation.
+\autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced.
+We then present possible future work in \autoref{sec:future-work}.
+We finally conclude in \autoref{sec:conclusion}.

### 2016-12-27 report/ordering.tex
WIP report

@@
 \label{sec:ordering}
-\jt{I'm not sure we still need this section}
-Distributed systems, like centralized systems, need to preserve the temporal order of events produced by concurrent processes in the system.
+\jt{Do we still want this section?}
+Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.
 When there are separated processes that can only communicate through messages, you cannot easily order these messages.

### 2016-12-27 report/evaluation.tex
WIP report

@@
 \textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well.
-Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\giga\bit\per\second} switched network.
+Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\gbps} switched network.
 Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM.

### 2016-12-28 report/evaluation.tex
WIP

@@
 \label{sec:evaluation}
-\jt{specify somewhere the problems encountered with JGroups}
 \pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
@@
 \par
-\begin{table*}[hpt]
+\begin{table}[hpt]
 	\centering
@@
 	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\resizebox{\linewidth}{!}{
 	\begin{tabular}{llSSS}
@@
 	\end{tabular}
+	}
 	\label{table:total-bandwidth} 
-\end{table*}
+\end{table}
 
@@
 Here we are well within this ratio.
-\begin{table*}[hpt]
+\begin{table}[hpt]
 	\centering
@@
 	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\resizebox{\linewidth}{!}{
 	\begin{tabular}{lSSS}
@@
 	\end{tabular}
+	}
 	\label{table:total-bandwidth-churn} 
-\end{table*}
+\end{table}
 
@@
 \subsection{Local Dissemination stretch}
-\begin{figure*}[hpt]
+\begin{figure}[hpt]
 	\centering
@@
 	\label{fig:local-delta}  
-\end{figure*}
+\end{figure}
 In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch.
@@
 	\vspace{-2mm} 
-	\caption{Throughput percentiles of a node during an experiment with churn}
+	\caption{Bandwidth percentiles of a node during an experiment with a real churn}
 	\vspace{-2mm} 
@@
 	\vspace{-2mm} 
-	\caption{Global dissemination times with churn}
+	\caption{Global dissemination times with a real churn}
 	\vspace{-2mm} 
@@
 	\vspace{-2mm} 
-	\caption{Local dissemination stretch with churn}
+	\caption{Local dissemination stretch with a real churn}
 	\vspace{-2mm}

### 2016-12-28 report/figures/no-churn/local-delta.tex
WIP report
WIP

--- 2016-12-22 report/figures/no-churn/local-delta.tex
@@
 width=\linewidth / 2.5,
+height=5cm,
 grid=major,
@@
 \addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaldeltas};
-\legend{EpTO, JGroups}
+\legend{\epto, \jgroups}
 %

### 2016-12-28 report/future-work.tex
WIP

@@
 \label{sec:future-work}
-\jt{Speak about running test on a bigger infrastructure, because the small ammount of nodes does not let us see the full potential of \epto. Then speak about the ongoing work of Push-Pull}
+\jt{Do we have Push-Pull references?}
+Future work should focus on testing \epto on a bigger cluster, preferably at a point where \jgroups can no longer function, as \epto is designed to work on very large number of peers.
+As it is, \epto does indeed work as intended in our infrastructure and under churn, but a larger test would showcase \epto's ability to handle large peer numbers.
+\par
+Future work also includes working on an alternate \epto dissemination algorithm.
+This algorithm would use a Push-Pull approach to further reduce the needed number of balls in the system and thus reduce the network load.

### 2016-12-29 report/abstract.tex
pass on thesis

@@
+\hm{Sorry for the delay.
+I am in dependency hell!!!
+Why on earth did you use tikz pictures?
+I gave up for now and commented all such figures.
+Can you compile them separately and include the pdf in the document?}
+
 \begin{abstract}
-\epto is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
+  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
+In this paper, we verify this claim by implementing EpTO, and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.
@@
 Our results shows \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
+  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
 \end{abstract}

### 2016-12-29 report/definitions.tex
pass on thesis

@@
 A process or peer is defined as an actor in our system running the application that needs total order.
-Each process will communicate with other processes in the distributed system, exchange events, and order them together.
+Each process communicates with other processes in the distributed system, exchange events, and order them together.
+\hm{I changed it, but you switched from the present to the future tense in the same sentence.
+You might as well check the rest of the document for other instances, if any.}
 \par
@@
 The moment where we publish data combined with the data is called an event.
+\hm{Not clear.
+Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
 \par
-A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic.
+A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic. \hm{Not clear. A ball is an abstraction. We reduce traffic by bundling events.}
 
@@
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well.
+in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well. \hm{This paragraph is confusing. Is it necessary?}
 \par
-Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a very low arbitrary probability a peer does not receive an event.
-In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
+Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a nonzero probability that a peer does not receive an event.
+In this instance there will be a hole in the sequence of delivered events, but the order of the delivered events is protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
+The probability that a peer does not receive an event is controlled by the \epto and depends on various parameters such as the fan-out, the number of rounds (time to live), the number of nodes, and the network conditions.
+In particular, this probability can be made smaller than the probability of catastrophic network failure.
+\hm{I rewrote this paragraph.
+Please reread.}
 \par

### 2016-12-29 report/evaluation.tex
pass on thesis

@@
 \label{sec:evaluation}
-\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
-\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
-\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
-\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
-\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
+%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
+%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
+%\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
+%\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
+%\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
 In this section we present our results.
@@
 When a node starts, it received an initial view and then quickly run the PSS four times to obtain a random enough view.
-\begin{figure}[hpt]
-	\centering
-	\input{figures/no-churn/bandwidth-figure.tex}
-	\vspace{-2mm} 
-	\caption{Bandwidth percentiles of a node during an experiment}
-	\vspace{-2mm} 
-	\label{fig:bandwidth}
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/no-churn/bandwidth-figure.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Bandwidth percentiles of a node during an experiment}
+% 	\vspace{-2mm} 
+% 	\label{fig:bandwidth}
+% \end{figure}
 In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups.
@@
 
-\begin{figure}[hpt]
-	\centering
-	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
-	\vspace{-2mm} 
-	\caption{Bandwidth percentiles of a node during an experiment with churn}
-	\vspace{-2mm} 
-	\label{fig:bandwidth-churn}
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Bandwidth percentiles of a node during an experiment with churn}
+% 	\vspace{-2mm} 
+% 	\label{fig:bandwidth-churn}
+% \end{figure}
 
@@
 \label{sub:local-times}
-\begin{figure*}[hpt]
-	\centering
-	\input{figures/no-churn/local-times.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination times}
-	\vspace{-2mm}
-	\label{fig:local-times} 
-\end{figure*}
+% \begin{figure*}[hpt]
+% 	\centering
+% 	\input{figures/no-churn/local-times.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination times}
+% 	\vspace{-2mm}
+% 	\label{fig:local-times} 
+% \end{figure*}
 
-\begin{figure}[hpt]
-	\centering
-	\input{figures/synthetic-churn/local-times-churn.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination times with churn}
-	\vspace{-2mm} 
-	\label{fig:local-times-churn} 
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/synthetic-churn/local-times-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination times with churn}
+% 	\vspace{-2mm} 
+% 	\label{fig:local-times-churn} 
+% \end{figure}
 In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}.
@@
 \subsection{Global Times}
-\begin{figure*}[hpt]
-	\centering
-	\input{figures/no-churn/global-times.tex}
-	\vspace{-2mm} 
-	\caption{Global dissemination times}
-	\vspace{-2mm}
-	\label{fig:global-times}  
-\end{figure*}
+% \begin{figure*}[hpt]
+% 	\centering
+% 	\input{figures/no-churn/global-times.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Global dissemination times}
+% 	\vspace{-2mm}
+% 	\label{fig:global-timexss}  
+% \end{figure*}
 
-\begin{figure}[hpt]
-	\centering
-	\input{figures/synthetic-churn/global-times-churn.tex}
-	\vspace{-2mm} 
-	\caption{Global dissemination times with churn}
-	\vspace{-2mm} 
-	\label{fig:global-times-churn} 
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/synthetic-churn/global-times-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Global dissemination times with churn}
+% 	\vspace{-2mm} 
+% 	\label{fig:global-times-churn} 
+% \end{figure}
 We computed global times as well.
@@
 \subsection{Local Dissemination stretch}
-\begin{figure}[hpt]
-	\centering
-	\input{figures/no-churn/local-delta.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination stretch}
-	\vspace{-2mm}
-	\label{fig:local-delta}  
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/no-churn/local-delta.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination stretch}
+% 	\vspace{-2mm}
+% 	\label{fig:local-delta}  
+% \end{figure}
 In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch.
@@
 
-\begin{figure}[hpt]
-	\centering
-	\input{figures/synthetic-churn/local-delta-churn.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination stretch with churn}
-	\vspace{-2mm}
-	\label{fig:local-delta-churn}   
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/synthetic-churn/local-delta-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination stretch with churn}
+% 	\vspace{-2mm}
+% 	\label{fig:local-delta-churn}   
+% \end{figure}
 In \autoref{fig:local-delta-churn} We see a completely different picture.
@@
 \subsection{Real Trace}
-\begin{figure}[hpt]
-	\centering
-	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
-	\vspace{-2mm} 
-	\caption{Bandwidth percentiles of a node during an experiment with a real churn}
-	\vspace{-2mm} 
-	\label{fig:bandwidth-real-churn}
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Bandwidth percentiles of a node during an experiment with a real churn}
+% 	\vspace{-2mm} 
+% 	\label{fig:bandwidth-real-churn}
+% \end{figure}
 \autoref{fig:bandwidth-real-churn} confirms the results from the synthetic churn.
@@
 %	\label{fig:total-bandwidth-real-churn}
-%\end{figure}
-\begin{figure}[hpt]
-	\centering
-	\input{figures/real-churn/local-times-real-churn.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination times}
-	\vspace{-2mm}
-	\label{fig:local-times-real-churn} 
-\end{figure}
-\begin{figure}[hpt]
-	\centering
-	\input{figures/real-churn/global-times-real-churn.tex}
-	\vspace{-2mm} 
-	\caption{Global dissemination times with a real churn}
-	\vspace{-2mm} 
-	\label{fig:global-times-real-churn} 
-\end{figure}
+% %\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/real-churn/local-times-real-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination times}
+% 	\vspace{-2mm}
+% 	\label{fig:local-times-real-churn} 
+% \end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/real-churn/global-times-real-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Global dissemination times with a real churn}
+% 	\vspace{-2mm} 
+% 	\label{fig:global-times-real-churn} 
+% \end{figure}
 The local and global dissemination times shown in respectively \autoref{fig:local-times-real-churn} and \autoref{fig:global-times-real-churn} still show \jgroups to outperform \epto in this scenario.
@@
 \par
-\begin{figure}[hpt]
-	\centering
-	\input{figures/real-churn/local-delta-real-churn.tex}
-	\vspace{-2mm} 
-	\caption{Local dissemination stretch with a real churn}
-	\vspace{-2mm}
-	\label{fig:local-delta-real-churn}   
-\end{figure}
+% \begin{figure}[hpt]
+% 	\centering
+% 	\input{figures/real-churn/local-delta-real-churn.tex}
+% 	\vspace{-2mm} 
+% 	\caption{Local dissemination stretch with a real churn}
+% 	\vspace{-2mm}
+% 	\label{fig:local-delta-real-churn}   
+% \end{figure}
 In \autoref{fig:local-delta-real-churn}, \epto as worse outliers than before and so does \jgroups-coord.

### 2016-12-29 report/framework.tex
WIP report
pass on thesis

@@
 EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}.
-Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
+Although the implementation in written with benchmarking in mind, the code could be adapted for real applications with only minimal changes to the sources.
+\hm{Write that the code is publicly available at some point, and where to find it.}
 \subsection{Architecture}
@@
 We implement our own PSS CYCLON operating on its own port.
-We use \jgroups 3.6.11.
+\hm{jgroups appears from nowhere.
+Rewrite that is used to compare with EpTO.} We use \jgroups 3.6.11.
 When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
-Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
+Finally, EpToTester \hm{The name is not capitalized properly (o instead of O).
+Write as a latex definition like epto, and change everywhere in the document.} is open-sourced on Github \hm{There it is.
+This should have been said earlier, when EpTOTester is first introduced.}\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
 \subsection{Deployment}
-\begin{figure}[htp]
-	\centering
-	\input{figures/architecture.tex}
-	\vspace{-2mm} 
-	\caption{EpTOTester architecture}
-	\vspace{-2mm} 
-	\label{fig:complete-architecture}
-\end{figure}
-To deploy \epto effectively we need two different Docker Swarm services.
-One containing the tracker and one containing all \epto replicas.
+% \begin{figure}[htp]
+% 	\centering
+% 	\input{figures/architecture.tex}
+% 	\vspace{-2mm} 
+% 	\caption{EpTOTester architecture}
+% 	\vspace{-2mm} 
+% 	\label{fig:complete-architecture}
+% \end{figure}
+To deploy \epto effectively we need two different Docker Swarm services: one containing the tracker and one containing all \epto replicas.
 Both of these services use the same network overlay to communicate.
@@
 This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
-Every benchmark is started through a Python script available on the master mode.
-This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
+Every benchmark is started through a Python script available on the master mode.\hm{What is the master mode?
+Node maybe?} This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
 All \epto parameters are customizable by using options provided through this script.
-Gradle is used to automate the image building in the python scripts.
+Gradle \hm{Never heard of this.
+Please add reference or url in a footnote.} is used to automate the image building in the python scripts.
 Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node.
@@
 The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}.
-\jt{The figure is inspired from the one in ErasureBench.
-How should I indicate this?}
+\jt{\autoref{fig:complete-architecture} is inspired from the one in ErasureBench.
+How should I indicate this?}\hm{Discussed during our call.
+A citation to Sebastien's paper in a footnote will do.}
 \subsection{Fault Injection}

### 2016-12-29 report/introduction.tex
pass on thesis

@@
 
-Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
+Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
 One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}.
@@
 There are some deterministic algorithms, which guarantee total order, agreement or other strong properties.
-Unfortunately, these types of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.
+Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.
 \par
@@
 Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability.
+\hm{The first three paragraphs use the word system a lot.
+You could rework this a little bit.}
 \par
-As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
+As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \hm{Write [6-13] instead.} \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
 These algorithms guarantee the dissemination of events in the system with high probability.
@@
 \par
-This is where \epto shines. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
+This is where \epto shines.
+\hm{What does ``where'' refer to?
+Not clear.} \epto, by mixing these two types of algorithms, \hm{Not sure I agree here.
+What two types?
+Not clear.
+Epto is not probabilistic and deterministic.
+Please rephrase.} provides total order along with scalability, validity and integrity.
 \epto consists of two distinct parts.
@@
 Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events.
+\hm{Write component instead of part.
+Fix before and after.}
 \par
@@
 With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Thanks to these approaches, \epto becomes highly scalable and resilient, while still providing total order.
+Thanks to these approaches, \epto is highly resilient with  imperfect networks and highly scalable as the networks grow, while always providing total order.
 \par
@@
 \par
-The next sections are dedicated to present our benchmarking solution and to present the results obtained.
+The next sections present our benchmarking solution and the results obtained.
 \autoref{sec:ordering} presents the different kinds of ordering.
@@
 We finally conclude in \autoref{sec:conclusion}.
+\hm{Write Section X instead of section X.
+Scan all document.}

### 2016-12-29 report/main.tex
Continue work on thesis
WIP report
pass on thesis

@@
 \DeclareSIUnit\mbps{MB/s}
+\DeclareSIUnit\gbps{Gb/s}
 
@@
 \newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}
+\newcommand{\hm}[1]{\mynote{Hugues}{#1}{orange}}
 
@@
 \pgfplotsset{
-	compat=1.14,
+	% compat=1.14,
 	tick label style={font=\small},
@@
 			\end{figure}
-			% === // Logos ===========================================
+			% === // Logos ==========thun=================================
 
@@
 \input{conclusion.tex}
-
 \printbibliography

### 2016-12-29 report/ordering.tex
pass on thesis

@@
 \label{sec:ordering}
-\jt{Do we still want this section?}
+\jt{Do we still want this section?}\hm{Absolutely!}
 Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.
@@
 We then have to use total order, otherwise we can end up in an inconsistent state.
+\hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}

### 2016-12-30 report/definitions.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

@@
 \par
-An event is defined as data sent at a given time by a peer.
-For example, we could imagine a system where each process publishes some data to other peers.
-The moment where we publish data combined with the data is called an event.
-\hm{Not clear.
+An event $j$ is defined as \\$e_j=(broadcasterID,timestamp,payload)$, where the timestamp is the local time at the process when the event is created and is used with the broadcasterID to order events.
+The payload represents the data sent.
+%\hm{Not clear.
 Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
 \par
-A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic. \hm{Not clear. A ball is an abstraction. We reduce traffic by bundling events.}
+A ball is an abstraction taken from the balls-and-bins problem.
+In practice, we bundle sets of events together and send them as one packet to reduce the network traffic overhead.
+\hm{Not clear.
+A ball is an abstraction.
+We reduce traffic by bundling events.} \jt{Is this better?}
 
@@
 of messages transmitted per process per round is logarithmic
-in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well. \hm{This paragraph is confusing. Is it necessary?}
+in the number of processes, ...''.
+The number of rounds needed to deliver is logarithmic as well.
+\hm{This paragraph is confusing.
+Is it necessary?} \jt{I remember having questions about what scaling well meant during the workshop that is why I added this definition.
+If you feel it is not needed we can remove it}
 \par
 Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a nonzero probability that a peer does not receive an event.
-In this instance there will be a hole in the sequence of delivered events, but the order of the delivered events is protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
-The probability that a peer does not receive an event is controlled by the \epto and depends on various parameters such as the fan-out, the number of rounds (time to live), the number of nodes, and the network conditions.
+In this instance there will be a hole in the sequence of delivered events, but the order of delivered events is protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
+The probability that a peer does not receive an event is controlled by \epto and depends on various parameters such as the fan-out, the number of rounds (time to live), the number of nodes, and the network conditions.
 In particular, this probability can be made smaller than the probability of catastrophic network failure.
-\hm{I rewrote this paragraph.
-Please reread.}
+%\hm{I rewrote this paragraph.
+Please reread.}\jt{Done.}
 \par

### 2016-12-30 report/evaluation.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

--- 2016-12-28 report/evaluation.tex
@@
 Each VM uses Debian as its O/S.
-EpTOTester is packaged as docker images.
+\eptotester is packaged as docker images.
 We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services.
@@
  	\centering
-	\input{figures/no-churn/bandwidth-figure.tex}
+ 	\includegraphics[width=\linewidth]{figures/bandwidth-nochurn.pdf}
  	\vspace{-2mm} 
@@
 Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum bandwidth.
-
  \begin{figure}[hpt]
  	\centering
-	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
+ 	\includegraphics[width=\linewidth]{figures/bandwidth-synth-churn.pdf}
  	\vspace{-2mm} 
@@
  \end{figure}
-
+\par
 In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns.
@@
 \end{table}
-
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/no-churn/total-bytes.tex}
-%	\vspace{-2mm} 
-%	\caption{Total bytes sent/received during an average experiment}
-%	\vspace{-2mm} 
-%	\label{fig:total-bandwidth}
-%\end{figure}
 In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups.
@@
 \end{table}
-
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/synthetic-churn/total-bytes-churn.tex}
-%	\vspace{-2mm} 
-%	\caption{Total bytes sent/received during an average experiment with churn}
-%	\vspace{-2mm} 
-%	\label{fig:total-bandwidth-churn}
-%\end{figure}
 In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn.
@@
  	\centering
-	\input{figures/no-churn/local-times.tex}
+ 	\includegraphics[width=\linewidth]{figures/local-times-nochurn.pdf}
  	\vspace{-2mm} 
@@
  	\centering
-	\input{figures/synthetic-churn/local-times-churn.tex}
+ 	\includegraphics[width=\linewidth]{figures/local-times-synth-churn.pdf}
  	\vspace{-2mm} 
@@
  	\centering
-	\input{figures/no-churn/global-times.tex}
+ 	\includegraphics[width=\linewidth]{figures/global-times-nochurn.pdf}
  	\vspace{-2mm} 
@@
  	\vspace{-2mm}
-	\label{fig:global-times}  
+ 	\label{fig:global-timexss}  
  \end{figure*}
@@
  	\centering
-	\input{figures/synthetic-churn/global-times-churn.tex}
+ 	\includegraphics[width=\linewidth]{figures/global-times-synth-churn.pdf}
  	\vspace{-2mm} 
@@
  	\centering
-	\input{figures/no-churn/local-delta.tex}
+ 	\includegraphics[width=\linewidth]{figures/local-diss-stretch-nochurn.pdf}
  	\vspace{-2mm} 
@@
  	\centering
-	\input{figures/synthetic-churn/local-delta-churn.tex}
+ 	\includegraphics[width=\linewidth]{figures/local-diss-stretch-synth-churn.pdf}
  	\vspace{-2mm} 
@@
 \end{table}
-
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/no-churn/total-events-sent.tex}
-%	\vspace{-2mm} 
-%	\caption{Total events sent per experiment on average}
-%	\vspace{-2mm}
-%	\label{fig:total-events}   
-%\end{figure}
 In \autoref{table:total-events} we see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
@@
 \end{table}
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/synthetic-churn/total-events-sent-churn.tex}
-%	\vspace{-2mm} 
-%	\caption{Total events sent per experiment during churn on average}
-%	\vspace{-2mm}
-%	\label{fig:total-events-churn}  
-%\end{figure}
-
 In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
@@
  	\centering
-	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
+ 	\includegraphics[width=\linewidth]{figures/bandwidth-real-churn.pdf}
  	\vspace{-2mm} 
@@
 \par
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/real-churn/total-bytes-real-churn.tex}
-%	\vspace{-2mm} 
-%	\caption{Total bytes sent/received during an average experiment with churn}
-%	\vspace{-2mm} 
-%	\label{fig:total-bandwidth-real-churn}
-%\end{figure}
 \begin{figure}[hpt]
 	\centering
-	\input{figures/real-churn/local-times-real-churn.tex}
+	\includegraphics[width=\linewidth]{figures/local-times-real-churn.pdf}
 	\vspace{-2mm} 
@@
 	\centering
-	\input{figures/real-churn/global-times-real-churn.tex}
+	\includegraphics[width=\linewidth]{figures/global-times-real-churn.pdf}
 	\vspace{-2mm} 
@@
 	\centering
-	\input{figures/real-churn/local-delta-real-churn.tex}
+	\includegraphics[width=\linewidth]{figures/local-diss-stretch-real-churn.pdf}
 	\vspace{-2mm} 
@@
 We see that on average \jgroups seems to deliver more events than \epto, but its standard deviation is way higher than that of \epto, suggesting \epto to be more stable and consistent.
-%\begin{figure}[h]
-%	\centering
-%	\input{figures/real-churn/total-events-sent-real-churn.tex}
-%	\vspace{-2mm} 
-%	\caption{Total events sent per experiment during churn on average}
-%	\vspace{-2mm}
-%	\label{fig:total-events-real-churn}  
-%\end{figure}
 \subsection{Problems encountered}

### 2016-12-30 report/framework.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

--- 2016-12-27 report/framework.tex
@@
 \end{figure}
-EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}.
-Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
+\eptotester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}.
+Although the implementation in written with benchmarking in mind, the code could be adapted for real applications with only minimal changes to the sources.
+\eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
 \subsection{Architecture}
@@
 \subsection{Implementation}
-We implement the data sent as randomly generated UUIDs.
+We implement the payload as randomly generated UUIDs.
 We implement our own PSS CYCLON operating on its own port.
+We compare \epto to the deterministic total order algorithm provided by \jgroups.
 We use \jgroups 3.6.11.
@@
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
-Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
 \subsection{Deployment}
@@
  	\centering
-	\input{figures/architecture.tex}
+ 	\includegraphics[width=\linewidth]{figures/complete-architecture.pdf}
  	\vspace{-2mm} 
-	\caption{EpTOTester architecture}
+ 	\caption[Caption]{\eptotester architecture\footnotemark}
  	\vspace{-2mm} 
@@
  \end{figure}
-To deploy \epto effectively we need two different Docker Swarm services.
-One containing the tracker and one containing all \epto replicas.
+\footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
+To deploy \epto effectively we need two different Docker Swarm services: one containing the tracker and one containing all \epto replicas.
 Both of these services use the same network overlay to communicate.
@@
 This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
-Every benchmark is started through a Python script available on the master mode.
+Every benchmark is started through a Python script available on the master node.
 This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
 All \epto parameters are customizable by using options provided through this script.
-Gradle is used to automate the image building in the python scripts.
+Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.
 Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node.
 \jgroups deployment is much the same.
-The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}.
-\jt{The figure is inspired from the one in ErasureBench.
-How should I indicate this?}
+The entire \eptotester architecture is shown in \autoref{fig:complete-architecture}.
 \subsection{Fault Injection}

### 2016-12-30 report/introduction.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

@@
 
-Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years.
+Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed network has been one the hot topics in distributed computing research for many years.
 One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}.
 \epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
-In addition, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.
+Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.
 \par 
-There are many other algorithms for disseminating and ordering events in a distributed system.
-There are some deterministic algorithms, which guarantee total order, agreement or other strong properties.
+There are many other algorithms for disseminating and ordering events in a distributed network.
+There are deterministic algorithms, which guarantee total order, agreement or other strong properties.
 Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.
 \par
-The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the system.
+The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the network.
 This causes a massive amount of network traffic and overhead on the system.
@@
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability.
+Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the algorithm's scalability.
 \hm{The first three paragraphs use the word system a lot.
-You could rework this a little bit.}
+You could rework this a little bit.} \jt{I tried replacing some 'system' by synonyms such as 'network' or 'structure'.}
 \par
-As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \hm{Write [6-13] instead.} \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}.
+As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach  \autocite{birman1999bimodal,carvalho2007emergent,demers1987epidemic,eugster2003lightweight,felber2002probabilistic,hayden1996probabilistic,kim2004gossip,Koldehofe02simplegossiping}.
 These algorithms guarantee the dissemination of events in the system with high probability.
@@
 \par
-This is where \epto shines.
-\hm{What does ``where'' refer to?
-Not clear.} \epto, by mixing these two types of algorithms, \hm{Not sure I agree here.
-What two types?
-Not clear.
-Epto is not probabilistic and deterministic.
-Please rephrase.} provides total order along with scalability, validity and integrity.
-\epto consists of two distinct parts.
-The first part is probabilistic dissemination.
-\epto guarantees that all peers will receive an event with arbitrarily high probability.
-The second part of \epto is deterministic ordering.
-Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events.
-\hm{Write component instead of part.
-Fix before and after.}
+\epto claims to solve these seemingly irreconcilable differences.
+By using a probabilistic dissemination together with deterministic ordering, \epto provides total order along with scalability, validity and integrity.
+Thus, \epto consists of two distinct components.
+\epto's dissemination component guarantees that all peers will receive an event with arbitrarily high probability.
+Once peers have received all events, \epto's ordering component orders them using the events timestamp, and in case of a tie use the broadcaster ID of the events.
 \par
-To model the first part, \epto is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
+To model the first component, \epto is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}.
 The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls.
@@
 The next sections present our benchmarking solution and the results obtained.
-\autoref{sec:ordering} presents the different kinds of ordering.
-\autoref{sec:definitions} defines the terms used in the paper.
-\autoref{sec:epto} presents \epto and the architecture of our implementation.
-\autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced.
-We then present possible future work in \autoref{sec:future-work}.
-We finally conclude in \autoref{sec:conclusion}.
-\hm{Write Section X instead of section X.
-Scan all document.}
+Section \ref{sec:ordering} presents the different kinds of ordering.
+Section \ref{sec:definitions} defines the terms used in the paper.
+Section \ref{sec:epto} presents \epto and the architecture of our implementation.
+Section \ref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced.
+We then present possible future work in Section \ref{sec:future-work}.
+We finally conclude in Section \ref{sec:conclusion}.

### 2016-12-30 report/ordering.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

@@
 \label{sec:ordering}
-\jt{Do we still want this section?}\hm{Absolutely!}
 Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.

### 2016-12-30 report/abstract.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues
Fix more stuff pointed by Hugues

@@
-\hm{Sorry for the delay.
-I am in dependency hell!!!
-Why on earth did you use tikz pictures?
-I gave up for now and commented all such figures.
-Can you compile them separately and include the pdf in the document?}
-
 \begin{abstract}
   \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO, and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
+In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.
 We then compare them using a synthetic churn trace and a real one using the same metrics.
-Our results shows \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
+Our results show \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
   \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}

### 2016-12-30 report/evaluation.tex
Fix more stuff pointed by Hugues

@@
 \textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well.
-Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\gbps} switched network.
+Is this fine as is?}Our cluster consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
 Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM.
 We deploy 12 virtual machines on these hosts.
-Each virtual machine has acces to 4 CPUs and 4 VCPUs as well as \SI{7}{\giga\byte} of RAM.
+Each virtual machine has access 4 VCPUs as well as \SI{7}{\giga\byte} of RAM.
 We use KVM as our hypervisor.
@@
 Every benchmark run with churn is run with the $(100,50)$ parameters.
-When we have a global even throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput.
+When we have a global event throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput.
 For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
 \par
-We compared \epto and \jgroups based on their bandwidth, the time to deliver every expected events, their local dissemination stretch and finally the number of events sent.
+We compare \epto and \jgroups based on their bandwidth, the local and global time to deliver every expected events, their local dissemination stretch and finally the number of events sent on average.
 \subsection{Bandwidth}
 The initial bandwidth peaks observed for \epto are due to the PSS initialization.
-When a node starts, it received an initial view and then quickly run the PSS four times to obtain a random enough view.
+When a node starts, it receives an initial view and then quickly runs the PSS active thread four times to jump-start itself.
  \begin{figure}[hpt]
@@
  	\vspace{-2mm} 
- 	\caption{Bandwidth percentiles of a node during an experiment}
+ 	\caption{Bandwidth percentiles of nodes during a stable experiment}
  	\vspace{-2mm} 
@@
 In lower peers scenario such as the ones presented in \autoref{fig:bandwidth} \jgroups is clearly at an advantage.
-Since \epto has a worse baseline we will reach the maximum bandwidth possible much quicker when increasing the event throughput.
+Since \epto has a worse baseline we will reach the maximum bandwidth available much quicker when increasing the event throughput.
 
@@
 Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}.
-Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum bandwidth.
+Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.
  \begin{figure}[hpt]
@@
  	\vspace{-2mm} 
- 	\caption{Bandwidth percentiles of a node during an experiment with churn}
+ 	\caption{Bandwidth percentiles of nodes during an unstable experiment with synthetic churn}
  	\vspace{-2mm} 
@@
 In the second one, we still kill one node per minute, but we immediately create a new one.
-For \jgroups we ran the benchmarks once without killing the coordinator and once killing it.
+For \jgroups we run the benchmarks once without killing the coordinator and once killing it.
 
-We can see that the churn doesn't affect  \epto at all when there are only nodes leaving.
+We can see that the churn does not affect \epto at all when there are only nodes leaving.
 We have small peaks when adding a node to \SI{3}{\mbps} or less.
@@
 In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn.
-One hypothesis for this is that a JGroups replica takes a longer time to start up compared to stopping a replica.
+One hypothesis for this is that a \jgroups replica takes a longer time to start up compared to stopping a replica.
 Therefore the overall benchmark has a longer time with less than 100 replicas.
 We also do not see a difference whether we kill the coordinator or not.
-This can be explained by the fact that before the detection of the faulty coordinator \jgroups is forced to a stop for time up to \SI{20}{\second}.
+This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}.
 The big spike afterwards compensates for this hole.
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
-We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system.
+We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.
 \subsection{Global Times}
@@
  	\vspace{-2mm}
- 	\label{fig:global-timexss}  
+ 	\label{fig:global-times}  
  \end{figure*}
@@
  \end{figure}
-We computed global times as well.
-They are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}.
+Global times are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}.
 These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.
@@
 When under churn, the 95th percentile of \jgroups is at \SI{31}{\milli\second} compared to \SI{14}{\milli\second} when there is no churn.
-The highest percentiles are at more than \SI{10}{\second}.
+The highest percentiles are at more than \SI{20}{\second}.
 This effect is due to the coordinator dying as we clearly see that it does not happen when we do not kill it.
@@
 On the contrary, \epto performs very well under churn.
-The median degrade a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
+The median degrades a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
 \subsection{Events sent}
@@
 \end{table}
+
 In \autoref{table:total-events} we see that both \epto and \jgroups deliver the same amount of events. This is expected in a perfect environment.
@@
 In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
-When killing and adding nodes, JGroups seems to deliver a smaller amount of nodes, however it does not look significant enough to draw any conclusion.
+When killing and adding nodes, \jgroups seems to deliver a smaller amount of nodes, however it does not look significant enough to draw any conclusion.
 \jt{I didn't run any statistical analysis}
@@
 \end{table}
-\autoref{table:total-bandwidth-real-churn} also confirms our earlier results.
-We run the experiment for approximately 3 times the synthetic period and we get approximately 3 times more \SI{}{\giga\byte} used.
+\autoref{table:total-bandwidth-real-churn} also confirms our earlier results. \epto is still within the $c*log_2(n)$ ratio fixed earlier.
 \par
@@
 The local and global dissemination times shown in respectively \autoref{fig:local-times-real-churn} and \autoref{fig:global-times-real-churn} still show \jgroups to outperform \epto in this scenario.
-Again we want to emphasize that \epto would more than probably outperform \jgroups in a bigger size cluster.
+Again, we want to emphasize that it is expected for \jgroups to perform better than \epto with the cluster size we use.
 \par

### 2016-12-30 report/framework.tex
Fix more stuff pointed by Hugues

@@
 \eptotester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}.
-Although the implementation in written with benchmarking in mind, the code could be adapted for real applications with only minimal changes to the sources.
+Although the implementation is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
 \eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
@@
 An application extending our Application class can \epto-broadcast and \epto-deliver events.
-Events broadcasted to the Dissemination component are sent over the network every $\delta$ period.
-Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
+Events broadcasted to the Dissemination component are sent over the network every $\delta$ period, where $\delta$ is a unit of time.
+Every ball received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
 They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
@@
 The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers.
-To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system.
+To obtain an initial random view, we contact an independent tracker implemented as a simple Python web server that keeps track of dead and alive nodes in the system.
 We want to emphasize that the tracker is not required.
@@
 We use \jgroups 3.6.11.
-When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers.
+When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker does not support MULTICAST.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.

### 2016-12-30 report/introduction.tex
Fix more stuff pointed by Hugues

@@
 By using a probabilistic dissemination together with deterministic ordering, \epto provides total order along with scalability, validity and integrity.
-Thus, \epto consists of two distinct components.
+\epto consists of two distinct components.
 \epto's dissemination component guarantees that all peers will receive an event with arbitrarily high probability.
@@
 	\includegraphics[width=\linewidth]{figures/BnB.jpeg}
-	\caption{Balls-and-Bins \autocite{bnb}.}
+	\caption[Caption]{Balls-and-Bins\footnotemark}
 	\label{fig:balls-and-bins}
 \end{figure}
+\footnotetext{Figure inspired from \autocite{bnb}}
 \par
@@
 Until now, the creators of \epto have only tested this algorithm in a simulated environment.
-In this work, we implement \epto in pure Kotlin and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems.
+In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems.
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.

### 2016-12-30 report/ordering.tex
Add total order use-case

@@
 Partial order can be useful since it is less costly to implement.
-However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order.
-We then have to use total order, otherwise we can end up in an inconsistent state.
-\hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}
+However, in some cases the order of all events is important.
+For example, imagine we have multiple databases around the world.
+We want them to appear as if it is only one database.
+To achieve this, every operation done a particular database would have to be replicated on all the others, we would then have to make sure that every operation is executed in the same order on every database to end up in the same state.
+\hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}\jt{I came up with a possible total order application.
+However I am not sure if it is the best use case.
+Maybe Miguel has a better idea.}

### 2016-12-31 report/definitions.tex
Fix more stuff pointed by Hugues
Fix small typos

@@
 \hm{I changed it, but you switched from the present to the future tense in the same sentence.
-You might as well check the rest of the document for other instances, if any.}
+You might as well check the rest of the document for other instances, if any.} \jt{I checked the whole document again I think it is fixed now}
 \par
-An event $j$ is defined as \\$e_j=(broadcasterID,timestamp,payload)$, where the timestamp is the local time at the process when the event is created and is used with the broadcasterID to order events.
+An event $j$ is defined as: $$e_j \triangleq (broadcasterID,timestamp,payload)$$ where the timestamp is the local time at the process when the event is created and is used with the broadcasterID to order events.
 The payload represents the data sent.
@@
 \par
-A ball is an abstraction taken from the balls-and-bins problem.
-In practice, we bundle sets of events together and send them as one packet to reduce the network traffic overhead.
-\hm{Not clear.
-A ball is an abstraction.
-We reduce traffic by bundling events.} \jt{Is this better?}
+A ball is an abstraction taken from the balls-and-bins problem. In practice, we bundle sets of events together and send them as one packet to reduce the network traffic overhead.
 
-We define \epto scaling well as it is defined in  \autocite{matos2015epto}:  ``The number
-of messages transmitted per process per round is logarithmic
-in the number of processes, ...''.
-The number of rounds needed to deliver is logarithmic as well.
-\hm{This paragraph is confusing.
-Is it necessary?} \jt{I remember having questions about what scaling well meant during the workshop that is why I added this definition.
-If you feel it is not needed we can remove it}
+The number of rounds needed to deliver an event, the fan-out and the number of messages transmitted per process per round are all logarithmic in the number of processes. \autocite{matos2015epto}
 \par

### 2016-12-31 report/evaluation.tex
Fix small typos

@@
 \par
-\textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well.
-Is this fine as is?}Our cluster consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
+\textbf{Testbed.} Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
+It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
 Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM.

### 2016-12-31 report/framework.tex
Fix small typos

@@
 They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
-The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie.
+The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.
 The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers.
@@
 We implement our own PSS CYCLON operating on its own port.
-We compare \epto to the deterministic total order algorithm provided by \jgroups.
+We compare \epto to the deterministic total order algorithm SEQUENCER provided by \jgroups.
 We use \jgroups 3.6.11.

### 2017-01-12 report/definitions.tex
General comments.

@@
 \label{sec:definitions}
+\mm{merge part about event with background section. The part about the conventions/settings used in the experiments should be moved to the evaluation.}
+
 A process or peer is defined as an actor in our system running the application that needs total order.

### 2017-01-12 report/evaluation.tex
General comments.

@@
 \par 
-We used specific settings recommended by \jgroups for a big cluster.
+We used specific settings recommended by \jgroups for a big cluster.\mm{space allowing, put the configuration used in the paper, otherwise point to the git file.}
 \par
@@
 \par
-\textbf{Testbed.} Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
+\textbf{Testbed.} \mm{this sentence is not necessary}Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
 It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
@@
 Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}.
-Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.
+Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.\mm{it would be perfect if we could experimentally show this.}
  \begin{figure}[hpt]
@@
 \subsection{Total GigaBytes sent/received}
-In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.
+In all cases, \jgroups receives more bytes than it sends.
+This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.\mm{we need to explain why this is the case}
 \par
@@
 \label{sub:local-times}
+\mm{explain what this is measuring, same for the other metrics}
  \begin{figure*}[hpt]
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
-We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.
+We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.}
 \subsection{Global Times}
@@
 \subsection{Problems encountered}
+\mm{the part regarding  could probably be expanded and moved to the previous section, where we discuss the implementation.
+the part regarding Jgroups could also be expanded.
+do we have a rough idea of the frequency of this ?
+it's also important do discuss how we measure holes so it's clear that this is not a problem on our side }
 The \epto simulation and theory is assuming we can have balls of infinite size.

### 2017-01-12 report/framework.tex
General comments.

@@
 \label{sec:epto}
+\mm{along the lines of presenting the tool as a contribution, could this be more elaborated ?
+The part that briefly describes EpTO could be moved to background.
+The part about the use of UDP, tracker et all belongs to the implementation.
+}
 \begin{figure}[htp]
@@
 When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker does not support MULTICAST.
-This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
+This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.\mm{cite to the jgroups documentation}
 \subsection{Deployment}

### 2017-01-12 report/introduction.tex
General comments.

@@
 \section{Introduction}
+
+\mm{Reorder a little bit such that the motivation and existing approaches to total order are introduced first and then epto is presented as an alternative approach.
+Also stress the motivation is to run a TO algorithm in a very large scale and draw insights from it, similarly to what was done in the papers "paxos made live" and "worldwide consensus.".
+Point out that there are few works evaluating such protocols in a really large scale.}\mm{We also need to make the contribution clear: In this case it is an implementation of a total order algorithm and making it available as open source.
+It would be good if we could put this on a more generic light, along the lines of "a set of tools to evaluate large scale algorithms" or at least present the tool itself, eptotester as a contribution, wdyt ?
+this would make the paper stronger }
 
@@
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
-\epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.
-These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
-Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases} in the system.
+\epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.\mm{this is imprecise and can lead to confusion, rephrase} These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
+Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases}\mm{link} in the system.
 \par

### 2017-01-12 report/ordering.tex
General comments.

@@
 \label{sec:ordering}
+\mm{rename to background. we also need to include some related work on TO (see epto paper) and on practical evaluation (e.g.: the papers pointed above plus splay}
 Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.
@@
 However I am not sure if it is the best use case.
-Maybe Miguel has a better idea.}
+Maybe Miguel has a better idea.}\mm{we can point that that TO is equivalent to consensus and used in approaches such as distirbuted state machines.
+Actually it is the fact that TO is not scalable that lead to the emergence of all the work on eventual consistency.}

### 2017-01-12 report/abstract.tex
Add Related work and other based on Miguel comments

@@
   \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
+In this paper, we verify this claim by implementing a set of tools to evaluate large scale algorithms, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.

### 2017-01-12 report/definitions.tex
Add Related work and other based on Miguel comments

--- 2016-12-31 report/definitions.tex
@@
 Each process communicates with other processes in the distributed system, exchange events, and order them together.
-\hm{I changed it, but you switched from the present to the future tense in the same sentence.
-You might as well check the rest of the document for other instances, if any.} \jt{I checked the whole document again I think it is fixed now}
 \par
@@
 The payload represents the data sent.
-%\hm{Not clear.
-Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
 \par

### 2017-01-12 report/introduction.tex
Add Related work and other based on Miguel comments

@@
 Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the algorithm's scalability.
-\hm{The first three paragraphs use the word system a lot.
-You could rework this a little bit.} \jt{I tried replacing some 'system' by synonyms such as 'network' or 'structure'.}
 \par
@@
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
-\epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.\mm{this is imprecise and can lead to confusion, rephrase} These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
-Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases}\mm{link} in the system.
+\epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view.
+Peers use a tracker that keeps track of dead and alive peers when they first join the network.
+These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
+Furthermore, we can inject synthetic or real churn using the Failure Trace Archive databases\footnote{\href{http://fta.scem.uws.edu.au/}{http://fta.scem.uws.edu.au/}} in the system.
 \par
 The next sections present our benchmarking solution and the results obtained.
-Section \ref{sec:ordering} presents the different kinds of ordering.
+Section \ref{sec:background} exposes the background of \eptotester and presents the different kind of ordering.
 Section \ref{sec:definitions} defines the terms used in the paper.

### 2017-01-12 report/ordering.tex
Add Related work and other based on Miguel comments

--- 2016-12-30 report/ordering.tex
@@
-\section{Ordering Algorithms}
-\label{sec:ordering}
+\section{Background}
+\label{sec:background}
 Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.
@@
 To achieve this, every operation done a particular database would have to be replicated on all the others, we would then have to make sure that every operation is executed in the same order on every database to end up in the same state.
-\hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}\jt{I came up with a possible total order application.
-However I am not sure if it is the best use case.
-Maybe Miguel has a better idea.}
+However, Total Order is not usually scalable.
+This lead to the emergence of most of the work on eventual consistency.
+\subsection{Related Work}
+Related work concerning Total Order is already extensively covered in \autocite{matos2015epto}.
+
+\autocites[]{Chandra2007}[]{Maia2011} each test extensively a different large scale distributed system.
+However, they do not offer a framework to benchmark different systems than theirs.
+\autocite{Leonini2009} presents a framework named SPlay to deploy such applications.
+Unfortunately, SPlay only allows for LUA application.
+\eptotester offers a convenient framework to deploy distributed applications provided they run in a docker container.

### 2017-01-12 report/abstract.tex
Take into account all remaining comments from Miguel

--- 2016-12-30 report/abstract.tex
@@
   \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
+In this paper, we verify this claim by designing an architecture to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.

### 2017-01-12 report/conclusion.tex
Take into account all remaining comments from Miguel

@@
 \label{sec:conclusion}
-\jt{Do later}
+Benchmarking large scale distributed systems is a hard task.
+One needs to have access to a rather large cluster, possibly on different sites to reproduce real world latencies.
+These benchmarks can cost a lot forcing researchers to rely on simulations.
+Unfortunately, simulations often disregard small but important details that only happen in a real world scenario.
+
+In this paper, we implement \eptotester an architecture designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network.
+Furthermore, we implement \epto and evaluate the claims made in \autocite{matos2015epto} against \jgroups SEQUENCER, a total order algorithm relying on a coordinator.
+Unfortunately, our cluster cannot handle more than 100 nodes at the same time.
+Due to this limitation, \epto performs worse than \jgroups.
+We want to reiterate that this is expected as \epto is designed for large distributed networks.
+\jt{Let's write it this way for the moment if we are able to run something on GCE or AWS we can always change it later.}

### 2017-01-12 report/evaluation.tex
Take into account all remaining comments from Miguel

@@
 \par 
-We used specific settings recommended by \jgroups for a big cluster.\mm{space allowing, put the configuration used in the paper, otherwise point to the git file.}
+We used specific settings recommended by \jgroups for a big cluster \autocite{udp-largecluster}.
 \par
@@
 \par
-\textbf{Testbed.} \mm{this sentence is not necessary}Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
+\textbf{Testbed.} \mm{this sentence is not necessary} \jt{I specified it as the following text is really close to the one in Sebastien work as we use the same cluster and I don't want any problem being accused of copy}Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
 It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
@@
 \subsection{Bandwidth}
+The bandwidth metric shows the instantaneous bandwidth, where the bandwidth is the addition of the upload and the download speed.
 The initial bandwidth peaks observed for \epto are due to the PSS initialization.
@@
 \subsection{Total GigaBytes sent/received}
+This metric shows the Total number of bytes sent and received on average during an experiment.
 In all cases, \jgroups receives more bytes than it sends.
-This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.\mm{we need to explain why this is the case}
+This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.
+Thus, when a new peer joins the cluster, the TCPGOSSIP will send a new view to every peer already in the cluster.
 \par
@@
 \label{sub:local-times}
-\mm{explain what this is measuring, same for the other metrics}
+The local time represents the total time it takes for a peer to deliver every events sent.
  \begin{figure*}[hpt]
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
-We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.}
+We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.} \jt{Yes I though about it too.
+Unfortunately we can't make it work for the moment :/}
 \subsection{Global Times}
+The global time represents the total time it takes for all peers to deliver every events sent.
  \begin{figure*}[hpt]
@@
 \subsection{Local Dissemination stretch}
+The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
  \begin{figure}[hpt]
@@
 In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch.
-The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
 \par
@@
 \subsection{Problems encountered}
-\mm{the part regarding  could probably be expanded and moved to the previous section, where we discuss the implementation.
-the part regarding Jgroups could also be expanded.
-do we have a rough idea of the frequency of this ?
-it's also important do discuss how we measure holes so it's clear that this is not a problem on our side }
-The \epto simulation and theory is assuming we can have balls of infinite size.
-In practice, packet loss happens.
-We must therefore limit the ball size to a practical number.
-We use balls with a maximum size of \SI{1432}{\byte}.
-We choose this number as the default MTU on Ethernet is \SI{1500}{\byte}.
-We then subtract the bytes required for the IPv4 and UDP headers.
-\par
 When running \jgroups we sometimes get failed runs, where either some peers will experience holes or peer will deliver out of order events.
-This only happens when there is churn happening and when the coordinator dies.
+We measure  These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies.
+These problems appeared 3 times out of 13 runs for \jgroups $1\{kill,add\}/minute$ and 2 times out of 7 for \jgroups with real churn.
 We have not investigated these problems further.
 This points to either \jgroups failing when there is too much churn or a bug present in the SEQUENCER implementation.
+
+We measure holes as one or more events that were delivered by at least one peer in the cluster but not on the peer containing the hole.
+As stated earlier, we specifically choose a $c$ high enough to never experience holes in \epto as \jgroups should never suffer from holes.
+If either a \jgroups or \epto fails to deliver every event sent during the run, the run is considered as having a problem.

### 2017-01-12 report/framework.tex
Take into account all remaining comments from Miguel

--- 2016-12-31 report/framework.tex
@@
 \label{sec:epto}
-\begin{figure}[htp]
-	\includegraphics[width=\linewidth]{figures/architecture.pdf}
-	\caption{\epto architecture \autocite{matos2015epto}.}
-	\label{fig:epto-architecture}
-\end{figure}
-\eptotester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}.
-Although the implementation is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
-\eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
-\subsection{Architecture}
-\autoref{fig:epto-architecture} illustrates the architecture of a single replica.
-An application extending our Application class can \epto-broadcast and \epto-deliver events.
-Events broadcasted to the Dissemination component are sent over the network every $\delta$ period, where $\delta$ is a unit of time.
-Every ball received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
-They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
-The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.
-The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers.
+\eptotester is a framework using Docker and Docker Swarm.
+It is designed to ease the deployment of large scale distributed systems, more particularly \jgroups and \epto.
+We use \eptotester to assess the claims made in \autocite{matos2015epto}.
+Although \epto's implementation is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
+\subsection{Implementation}
+We implement the network layer using UDP and a PSS CYCLON to obtain a random view of peers.
 To obtain an initial random view, we contact an independent tracker implemented as a simple Python web server that keeps track of dead and alive nodes in the system.
@@
 In practice it works well, but using a DHT is certainly a possibility.
-\subsection{Implementation}
+
 We implement the payload as randomly generated UUIDs.
@@
 We use \jgroups 3.6.11.
-When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker does not support MULTICAST.
+When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker services do not yet support MULTICAST.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
+
+The \epto simulation and theory is assuming we can have balls of infinite size.
+In practice, packet loss happens.
+We must therefore limit the ball size to a practical number.
+We use balls with a maximum size of \SI{1432}{\byte}.
+We choose this number as the default MTU on Ethernet is \SI{1500}{\byte}.
+We then subtract the bytes required for the IPv4 and UDP headers.
 \subsection{Deployment}

### 2017-01-12 report/introduction.tex
Take into account all remaining comments from Miguel

@@
 \section{Introduction}
-
-\mm{Reorder a little bit such that the motivation and existing approaches to total order are introduced first and then epto is presented as an alternative approach.
-Also stress the motivation is to run a TO algorithm in a very large scale and draw insights from it, similarly to what was done in the papers "paxos made live" and "worldwide consensus.".
-Point out that there are few works evaluating such protocols in a really large scale.}\mm{We also need to make the contribution clear: In this case it is an implementation of a total order algorithm and making it available as open source.
-It would be good if we could put this on a more generic light, along the lines of "a set of tools to evaluate large scale algorithms" or at least present the tool itself, eptotester as a contribution, wdyt ?
-this would make the paper stronger }
+\jt{If we present it as a framework we need to find a new name}
 
 Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed network has been one the hot topics in distributed computing research for many years.
-One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}.
-\epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system.
-Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.
-\par 
-There are many other algorithms for disseminating and ordering events in a distributed network.
+There are many algorithms for disseminating and ordering events in a distributed network.
 There are deterministic algorithms, which guarantee total order, agreement or other strong properties.
@@
 As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
+
+One of the recently designed algorithms in the field is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to solve these seemingly irreconcilableŝ
+differences. Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.
 \par
-\epto claims to solve these seemingly irreconcilable differences.
 By using a probabilistic dissemination together with deterministic ordering, \epto provides total order along with scalability, validity and integrity.
@@
 Thanks to these approaches, \epto is highly resilient with  imperfect networks and highly scalable as the networks grow, while always providing total order.
-\par
+\subsection{Contributions}
 Until now, the creators of \epto have only tested this algorithm in a simulated environment.
-In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems.
+In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking framework called \eptotester to show that \epto is suitable for real-world large-scale distributed systems.
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
@@
 Furthermore, we can inject synthetic or real churn using the Failure Trace Archive databases\footnote{\href{http://fta.scem.uws.edu.au/}{http://fta.scem.uws.edu.au/}} in the system.
+\eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
+Finally we want to emphasize that few works evaluate such protocols in a really large scale \autocites[]{Chandra2007}[]{Maia2011}.
 \par

### 2017-01-12 report/ordering.tex
Take into account all remaining comments from Miguel

@@
 This lead to the emergence of most of the work on eventual consistency.
+
+\subsection{\epto Architecture}
+\begin{figure}[htp]
+	\includegraphics[width=\linewidth]{figures/architecture.pdf}
+	\caption{\epto architecture \autocite{matos2015epto}.}
+	\label{fig:epto-architecture}
+\end{figure}
+\autoref{fig:epto-architecture} illustrates the architecture of a single \epto replica.
+An application extending our Application class can \epto-broadcast and \epto-deliver events.
+Events broadcasted to the Dissemination component are sent over the network every $\delta$ period, where $\delta$ is a unit of time.
+Every ball received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
+They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
+The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.
+The network layer is managed by \eptotester.
 \subsection{Related Work}
@@
 
-\autocites[]{Chandra2007}[]{Maia2011} each test extensively a different large scale distributed system.
-However, they do not offer a framework to benchmark different systems than theirs.
+Both \autocite{Chandra2007} and \autocite{Maia2011} extensively test a particular large scale distributed system.
+However, they do not offer a framework to benchmark different systems.
 \autocite{Leonini2009} presents a framework named SPlay to deploy such applications.
-Unfortunately, SPlay only allows for LUA application.
-\eptotester offers a convenient framework to deploy distributed applications provided they run in a docker container.
+Unfortunately, SPlay only allows for LUA applications to run.
+\eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container .

### 2017-01-12 report/introduction.tex
Fix small typos

@@
 
-One of the recently designed algorithms in the field is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to solve these seemingly irreconcilableŝ
-differences. Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.
+One of the recently designed algorithms in the field is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to solve these seemingly irreconcilable
+differences between deterministic and probabilistic algorithms.
+Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.
 \par
@@
 The next sections present our benchmarking solution and the results obtained.
-Section \ref{sec:background} exposes the background of \eptotester and presents the different kind of ordering.
+Section \ref{sec:background} presents the different kind of ordering, \epto architecture and related work.
 Section \ref{sec:definitions} defines the terms used in the paper.

### 2017-01-12 report/ordering.tex
Fix small typos

@@
 This lead to the emergence of most of the work on eventual consistency.
+\epto claims to solve this issue.
 
@@
 \subsection{Related Work}
-Related work concerning Total Order is already extensively covered in \autocite{matos2015epto}.
+Related work concerning Total Order is already extensively covered in \autocite{matos2015epto}. \jt{Is this sufficient or should I recap everything presented in \autocite{matos2015epto}?}
 
@@
 Unfortunately, SPlay only allows for LUA applications to run.
-\eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container .
+\eptotester offers a convenient architecture to deploy distributed applications regardless of the programming language used, provided they run in a docker container.

### 2017-01-13 projects/EpTOTester/cluster/run_benchmarks.py
Initial work for a framework

@@ -23,7 +23,6 @@ from datetime import datetime
 from docker import errors
 from docker import types
 from docker import utils
-from logging import config
 from nodes_trace import NodesTrace
 
 
@@ -31,11 +30,7 @@ cli = docker.Client(base_url='unix://var/run/docker.sock')
 MANAGER_IP = '172.16.2.119'
 LOCAL_DATA = '/home/jocelyn/tmp/data'
 CLUSTER_DATA = '/home/debian/data'
-DISTANT_REPOSITORY = 'swarm-m:5000/'
-SERVICE_NAME = 'epto'
-TRACKER_NAME = 'epto-tracker'
-NETWORK_NAME = 'epto_network'
-SUBNET = '172.130.0.0/16'
+CONFIG = None
 
 
 def create_service(service_name, image, env=None, mounts=None, placement=None, replicas=1):
@@ -47,7 +42,7 @@ def create_service(service_name, image, env=None, mounts=None, placement=None, r
                                    placement=placement)
     logger.debug(task_tmpl)
     cli.create_service(task_tmpl, name=service_name, mode={'Replicated': {'Replicas': replicas}},
-                       networks=[{'Target': NETWORK_NAME}])
+                       networks=[{'Target': CONFIG['service']['network']['name']}])
 
 
 def run_churn(time_to_start):
@@ -66,10 +61,10 @@ def run_churn(time_to_start):
         repository = ''
     else:
         hosts_fname = 'hosts'
-        repository = DISTANT_REPOSITORY
+        repository = CONFIG['repository']['name']
 
     delta = args.period
-    churn = Churn(hosts_filename=hosts_fname, service_name=SERVICE_NAME, repository=repository)
+    churn = Churn(hosts_filename=hosts_fname, service_name=CONFIG['service']['name'], repository=repository)
     churn.set_logger_level(log_level)
 
     # Add initial cluster
@@ -136,16 +131,12 @@ if __name__ == '__main__':
     parser = argparse.ArgumentParser(description='Run benchmarks',
                                      formatter_class=argparse.ArgumentDefaultsHelpFormatter)
     parser.add_argument('peer_number', type=int, help='With how many peer should it be ran')
-    parser.add_argument('delta', type=int, help='Period for an EpTO round in ms')
     parser.add_argument('time_add', type=int, help='Delay experiments start in seconds')
     parser.add_argument('time_to_run', type=int, help='For how long should the experiment run in seconds')
-    parser.add_argument('rate', type=int, help='At which frequency should a peer send an event in ms')
[…]

### 2017-01-13 projects/EpTOTester/cluster/run_benchmarks.py
WIP

@@ -51,10 +51,15 @@ def run_churn(time_to_start):
         logger.info(args.synthetic)
         nodes_trace = NodesTrace(synthetic=args.synthetic)
     else:
-        # Website02.db epoch starts 1st January 201. Exact formula obtained from Sebastien Vaucher
-        websites_epoch = 730753 + 1 + 86400. / (16 * 3600 + 11 * 60 + 10)
-        nodes_trace = NodesTrace(database='websites02.db', min_time=websites_epoch + 22200,
-                                 max_time=websites_epoch + 22200 + 7200, time_factor=2)
+        real_churn_params = CONFIG['real_churn']
+        # Website02.db epoch starts 1st January 2001. Exact formula obtained from Sebastien Vaucher
+        # websites_epoch = 730753 + 1 + 86400. / (16 * 3600 + 11 * 60 + 10)
+        nodes_trace = NodesTrace(database='websites02.db', min_time=real_churn_params['epoch'] +
+                                                                    real_churn_params['start_time'],
+                                 max_time=real_churn_params['epoch'] +
+                                          real_churn_params['start_time'] +
+                                          real_churn_params['duration'],
+                                 time_factor=real_churn_params['time_factor'])
 
     if args.local:
         hosts_fname = None
@@ -168,8 +173,8 @@ if __name__ == '__main__':
     def signal_handler(signal, frame):
         logger.info('Stopping Benchmarks')
         try:
-            cli.remove_service(TRACKER_NAME)
-            cli.remove_service(SERVICE_NAME)
+            cli.remove_service(CONFIG['tracker']['name'])
+            cli.remove_service(CONFIG['service']['name'])
             if not args.local:
                 time.sleep(15)
                 with open('hosts', 'r') as file:
@@ -222,11 +227,11 @@ if __name__ == '__main__':
         wait_on_service(CONFIG['tracker']['name'], 1)
         time_to_start = int((time.time() * 1000) + args.time_add)
         logger.debug(datetime.utcfromtimestamp(time_to_start / 1000).isoformat())
-        environment_vars = {'PEER_NUMBER': args.peer_number, 'DELTA': args.delta,
-                            'TIME': time_to_start, 'TIME_TO_RUN': args.time_to_run,
-                            'RATE': args.rate, 'FIXED_RATE': args.fixed_rate,
-                            'CONSTANT': args.constant, 'CHURN_RATE': args.churn_rate,
-                            'MESSAGE_LOSS': args.message_loss}
[…]

### 2017-01-16 report/framework.tex

@@
 \subsection{Implementation}
-We implement the network layer using UDP and a PSS CYCLON to obtain a random view of peers.
+We implement the network layer using UDP and our own PSS CYCLON operating on its own port to obtain a random view of peers.
 To obtain an initial random view, we contact an independent tracker implemented as a simple Python web server that keeps track of dead and alive nodes in the system.
@@
 We implement the payload as randomly generated UUIDs.
-We implement our own PSS CYCLON operating on its own port.
 We compare \epto to the deterministic total order algorithm SEQUENCER provided by \jgroups.
 We use \jgroups 3.6.11.
-When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker services do not yet support MULTICAST.
+When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers, as Docker services do not yet support MULTICAST.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.

### 2017-01-18 projects/EpTOTester/scripts/container-start-script.sh
Lay out base for framework

@@ -3,21 +3,14 @@
 
 MY_IP_ADDR=$(/bin/hostname -i)
 
-echo 'Starting epto peer'
+echo 'Starting Peer'
 MY_IP_ADDR=($MY_IP_ADDR)
 echo "${MY_IP_ADDR[0]}"
-echo "${PEER_NUMBER}"
-echo "$DELTA"
-echo "$TIME"
-echo "${TIME_TO_RUN}"
-echo "$RATE"
-echo "$CONSTANT"
-echo "$FIXED_RATE"
-echo "$CHURN_RATE"
-echo "$MESSAGE_LOSS"
 
 dstat_pid=0
-java_pid=0
+app_pid=0
+
+mkdir -p /data/capture
 
 signal_handler() {
     curl "http://epto-tracker:4321/terminate"
@@ -26,8 +19,8 @@ signal_handler() {
         kill $dstat_pid
     fi
 
-    if [ $java_pid -ne 0 ]; then
-        kill $java_pid
+    if [ $app_pid -ne 0 ]; then
+        kill $app_pid
     fi
     echo "KILLED PROCESSES"
 
@@ -45,7 +38,7 @@ dstat_pid=${!}
 java -Xms100m -Xmx210m -cp ./epto-1.0-SNAPSHOT-all.jar -Dlogfile.name="${MY_IP_ADDR[0]}" -Djava.net.preferIPv4Stack=true utilities.Main --delta "$DELTA" \
 --rate "$RATE" -c "$CONSTANT" --fixed-rate "$FIXED_RATE" --churn-rate "$CHURN_RATE" --message-loss "$MESSAGE_LOSS" \
 "${MY_IP_ADDR[0]}" "http://epto-tracker:4321" "${PEER_NUMBER}" "$TIME" "$TIME_TO_RUN" &
-java_pid=${!}
+app_pid=${!}
 
-wait ${java_pid}
+wait ${app_pid}
 kill ${dstat_pid}

### 2017-01-18 projects/JGroupsTester/cluster/churn.py

@@ -1,46 +1,21 @@
 #!/usr/bin/env python3
-import glob
 import logging
 import random
-import re
 import subprocess
-
-
-def churn_tuple(s):
-    try:
-        _to_kill, _to_create = map(int, s.split(','))
-        return _to_kill, _to_create
-    except:
-        raise TypeError("Tuples must be (int, int)")
-
-
-def get_peer_list(path='../data/*.txt'):
-    with open(glob.glob(path)[0], 'r') as f:
-        a_list = []
-        for line in f.readlines():
-            match = re.match(r'\d+ - View: (.+)', line)
-            if match:
-                a_list = match.group(1).split(',')
-                break
-        if not a_list:
-            raise LookupError('No view found in file {}'.format(f.name))
-        return a_list
 
 
 class Churn:
     """
     Author: Jocelyn Thode
 
-    A class in charge of adding/suspending nodes to create churn in a JGroups SEQUENCER cluster
+    A class in charge of adding/suspending nodes to create churn in a EpTO cluster
     """
 
-    def __init__(self, hosts_filename=None, kill_coordinator_round=None, service_name='jgroups', repository=''):
+    def __init__(self, hosts_filename=None, service_name='', repository=''):
         self.containers = {}
-        self.coordinator = None
         self.peer_list = []
-        self.periods = 0
-        self.suspended_containers = []
         self.logger = logging.getLogger('churn')
+        self.suspended_containers = []
 
         self.service_name = service_name
         self.repository = repository
@@ -48,11 +23,6 @@ class Churn:
         if hosts_filename is not None:
             with open(hosts_filename, 'r') as file:
                 self.hosts += list(line.rstrip() for line in file)
-        if kill_coordinator_round is None:
-            self.kill_coordinator_round = []
-        else:
-            self.kill_coordinator_round = kill_coordinator_round
-        self.kill_index = 0
         self.cluster_size = 0
 
     def suspend_processes(self, to_suspend_nb):
@@ -69,48 +39,14 @@ class Churn:
 
         for i in range(to_suspend_nb):
             command_suspend = ["docker", "kill", '--signal=SIGUSR1']
-            self.logger.info("Variables: {} {}".format(self.periods, self.kill_coordinator_round))
-            # If we missed kill period kill the coord at the next chance
-            if self.periods in self.kill_coordinator_round or \
-                    (self.kill_index < len(self.kill_coordinator_round) and
-                     self.periods > self.kill_coordinator_round[self.kill_index]):
-                self.logger.info("Killing coordinator")
[…]

### 2017-01-18 projects/JGroupsTester/cluster/jgroups-churn.py

@@ -1,21 +1,40 @@
 #!/usr/bin/env python3
+import glob
 import logging
 import random
+import re
 import subprocess
 
+from churn import Churn
 
-class Churn:
+
+def get_peer_list(path='../data/*.txt'):
+    with open(glob.glob(path)[0], 'r') as f:
+        a_list = []
+        for line in f.readlines():
+            match = re.match(r'\d+ - View: (.+)', line)
+            if match:
+                a_list = match.group(1).split(',')
+                break
+        if not a_list:
+            raise LookupError('No view found in file {}'.format(f.name))
+        return a_list
+
+
+class JGroupsChurn(Churn):
     """
     Author: Jocelyn Thode
 
-    A class in charge of adding/suspending nodes to create churn in a EpTO cluster
+    A class in charge of adding/suspending nodes to create churn in a JGroups SEQUENCER cluster
     """
 
-    def __init__(self, hosts_filename=None, service_name='', repository=''):
+    def __init__(self, hosts_filename=None, service_name='', repository='', kill_coordinator_round=None):
         self.containers = {}
+        self.coordinator = None
         self.peer_list = []
-        self.logger = logging.getLogger('churn')
+        self.periods = 0
         self.suspended_containers = []
+        self.logger = logging.getLogger('churn')
 
         self.service_name = service_name
         self.repository = repository
@@ -23,6 +42,11 @@ class Churn:
         if hosts_filename is not None:
             with open(hosts_filename, 'r') as file:
                 self.hosts += list(line.rstrip() for line in file)
+        if kill_coordinator_round is None:
+            self.kill_coordinator_round = []
+        else:
+            self.kill_coordinator_round = kill_coordinator_round
+        self.kill_index = 0
         self.cluster_size = 0
 
     def suspend_processes(self, to_suspend_nb):
@@ -39,14 +63,48 @@ class Churn:
 
         for i in range(to_suspend_nb):
             command_suspend = ["docker", "kill", '--signal=SIGUSR1']
+            self.logger.info("Variables: {} {}".format(self.periods, self.kill_coordinator_round))
+            # If we missed kill period kill the coord at the next chance
+            if self.periods in self.kill_coordinator_round or \
+                    (self.kill_index < len(self.kill_coordinator_round) and
+                     self.periods > self.kill_coordinator_round[self.kill_index]):
+                self.logger.info("Killing coordinator")
+                command_suspend += [self.coordinator]
+                for host in self.hosts:
+                    self._refresh_host_containers(host)
[…]

### 2017-01-18 projects/JGroupsTester/cluster/run_benchmarks.py

@@ -18,13 +18,11 @@ import time
 import yaml
 
 from churn import Churn
-from churn import churn_tuple
 from churn import get_peer_list
 from datetime import datetime
 from docker import errors
 from docker import types
 from docker import utils
-from logging import config
 from nodes_trace import NodesTrace
 
 
@@ -72,8 +70,8 @@ def run_churn(time_to_start):
 
     delta = args.delta
     logging.debug("Kill_coordinator = {}".format(args.kill_coordinator))
-    churn = Churn(hosts_filename=hosts_fname, kill_coordinator_round=args.kill_coordinator, service_name=SERVICE_NAME,
-                  repository=repository)
+    churn = Churn(hosts_filename=hosts_fname, service_name=SERVICE_NAME,
+                  repository=repository, kill_coordinator_round=args.kill_coordinator)
     churn.set_logger_level(log_level)
     logger.debug("Kill coordinator rounds: {}".format(churn.kill_coordinator_round))
 
@@ -139,6 +137,14 @@ def create_logger():
     with open('logger.yaml') as f:
         conf = yaml.load(f)
         logging.config.dictConfig(conf)
+
+
+def churn_tuple(s):
+    try:
+        _to_kill, _to_create = map(int, s.split(','))
+        return _to_kill, _to_create
+    except:
+        raise TypeError("Tuples must be (int, int)")
 
 if __name__ == '__main__':
     parser = argparse.ArgumentParser(description='Run benchmarks',

### 2017-01-18 projects/EpTOTester/cluster/run_benchmarks.py
Remove EpTO reference

@@ -250,8 +250,7 @@ if __name__ == '__main__':
                        mounts=[types.Mount(target='/data', source=log_storage, type='bind')],
                        replicas=service_replicas, mem_limit=APP_CONFIG['service']['mem_limit'])
 
-        # TODO clean text
-        logger.info('Running EpTO tester -> Experiment: {:d}/{:d}'.format(run_nb, args.runs))
+        logger.info('Running Benchmark -> Experiment: {:d}/{:d}'.format(run_nb, args.runs))
         if args.churn:
             thread = threading.Thread(target=run_churn, args=[time_to_start + args.delay], daemon=True)
             thread.start()

### 2017-01-18 projects/EpTOTester/setup_cluster.sh
Add username variable

@@ -2,6 +2,7 @@
 # This scripts  will build the images, push them to the repository run the tests
 # Credits : https://github.com/sebyx31/ErasureBench/blob/master/projects/erasure-tester/benchmark_on_cluster.sh
 
+USERNAME="debian"
 MANAGER_IP=172.16.2.119
 
 trap "kill ${ssh_pid}" TERM INT
@@ -21,4 +22,4 @@ docker push localhost:5000/epto-tracker:latest
 
 kill ${ssh_pid}
 
-rsync -av --copy-links cluster/ debian@${MANAGER_IP}:~/epto
+rsync -av --copy-links cluster/ ${USERNAME}@${MANAGER_IP}:~/epto

### 2017-01-18 projects/EpTOTester/cluster/benchmark.py
Work on generalizing Benchmark

@@ -1,0 +1,173 @@
+#!/usr/bin/env python3
+"""
+Author: Jocelyn Thode
+
+This script is in charge of running the benchmarks either locally or on the cluster.
+
+It creates the network, services and the churn if we need it.
+
+"""
+import argparse
+import docker
+import logging
+import re
+import signal
+import subprocess
+import threading
+import time
+import yaml
+
+from churn import Churn
+from datetime import datetime
+from docker import errors
+from docker import types
+from docker import utils
+from logging import config
+from nodes_trace import NodesTrace
+
+
+class Benchmark:
+    def __init__(self, config, local, use_tracker, churn=None):
+        self.cli = docker.Client(base_url='unix://var/run/docker.sock')
+        self.config = config
+        self.local = local
+        self.logger = logging.getLogger('benchmarks')
+        self.churn = churn
+        self.use_tracker = use_tracker
+
+    def run(self):
+        if args.local:
+            service_image = self.config['service']['name']
+            if args.tracker:
+                tracker_image = self.config['tracker']['name']
+            with subprocess.Popen(['../gradlew', '-p', '..', 'docker'],
+                                  stdin=subprocess.PIPE, stdout=subprocess.PIPE, stderr=subprocess.PIPE,
+                                  universal_newlines=True) as p:
+                for line in p.stdout:
+                    print(line, end='')
+        else:
+            service_image = self.config['repository'] + self.config['service']['name']
+            for line in self.cli.pull(service_image, stream=True, decode=True):
+                print(line)
+            if self.use_tracker:
+                tracker_image = self.config['repository'] + self.config['tracker']['name']
+                for line in self.cli.pull(tracker_image, stream=True):
+                    print(line)
+            try:
+                self.cli.init_swarm()
+                if not self.local:
+                    self.logger.info('Joining Swarm on every hosts:')
+                    token = self.cli.inspect_swarm()['JoinTokens']['Worker']
+                    subprocess.call(['parallel-ssh', '-t', '0', '-h', 'hosts', 'docker', 'swarm',
+                                     'join', '--token', token, '{:s}:2377'.format(MANAGER_IP)])
+                ipam_pool = utils.create_ipam_pool(subnet=self.config['service']['network']['subnet'])
+                ipam_config = utils.create_ipam_config(pool_configs=[ipam_pool])
+                self.cli.create_network(self.config['service']['network']['name'], 'overlay', ipam=ipam_config)
+            except errors.APIError:
[…]

### 2017-01-18 projects/EpTOTester/cluster/run_benchmarks.py

@@ -18,6 +18,7 @@ import time
 import yaml
 
 from churn import Churn
+from benchmark import Benchmark
 from datetime import datetime
 from docker import errors
 from docker import types
@@ -26,98 +27,12 @@ from logging import config
 from nodes_trace import NodesTrace
 
 
-cli = docker.Client(base_url='unix://var/run/docker.sock')
 with open('config.yaml', 'r') as f:
     config = yaml.load(f)
     MANAGER_IP = config['manager_ip']
     LOCAL_MANAGER_IP = config['local_manager_ip']
     LOCAL_DATA = config['local_data']
     CLUSTER_DATA = config['cluster_data']
-
-
-def create_service(service_name, image, env=None, mounts=None, placement=None, replicas=1, mem_limit=314572800):
-    container_spec = types.ContainerSpec(image=image, env=env, mounts=mounts)
-    logger.debug(container_spec)
-    task_tmpl = types.TaskTemplate(container_spec,
-                                   resources=types.Resources(mem_limit=mem_limit),
-                                   restart_policy=types.RestartPolicy(),
-                                   placement=placement)
-    logger.debug(task_tmpl)
-    cli.create_service(task_tmpl, name=service_name, mode={'Replicated': {'Replicas': replicas}},
-                       networks=[{'Target': APP_CONFIG['service']['network']['name']}])
-
-
-def run_churn(time_to_start):
-    logger.debug('Time to start churn: {:d}'.format(time_to_start))
-    if args.synthetic:
-        logger.info(args.synthetic)
-        nodes_trace = NodesTrace(synthetic=args.synthetic)
-    else:
-        real_churn_params = APP_CONFIG['real_churn']
-        # Website02.db epoch starts 1st January 2001. Exact formula obtained from Sebastien Vaucher
-        # websites_epoch = 730753 + 1 + 86400. / (16 * 3600 + 11 * 60 + 10)
-        nodes_trace = NodesTrace(database=real_churn_params['database'], min_time=real_churn_params['epoch'] +
-                                                                    real_churn_params['start_time'],
-                                 max_time=real_churn_params['epoch'] +
-                                          real_churn_params['start_time'] +
-                                          real_churn_params['duration'],
-                                 time_factor=real_churn_params['time_factor'])
-
-    if args.local:
-        hosts_fname = None
-        repository = ''
-    else:
-        hosts_fname = 'hosts'
[…]

### 2017-01-18 projects/JGroupsTester/cluster/benchmark.py

@@ -1,0 +1,294 @@
+#!/usr/bin/env python3
+"""
+Author: Jocelyn Thode
+
+This script is in charge of running the benchmarks either locally or on the cluster.
+
+It creates the network, services and the churn if we need it.
+
+"""
+import argparse
+import docker
+import logging
+import re
+import signal
+import subprocess
+import threading
+import time
+import yaml
+
+from jgroups_churn import JGroupsChurn
+from datetime import datetime
+from docker import errors
+from docker import types
+from docker import utils
+from nodes_trace import NodesTrace
+
+
+cli = docker.Client(base_url='unix://var/run/docker.sock')
+MANAGER_IP = '172.16.2.119'
+LOCAL_DATA = '/home/jocelyn/tmp/data'
+CLUSTER_DATA = '/home/debian/data'
+LOCAL_DATA_FILES = '{:s}/*.txt'.format(LOCAL_DATA)
+CLUSTER_DATA_FILES = '{:s}/*.txt'.format(CLUSTER_DATA)
+DISTANT_REPOSITORY = 'swarm-m:5000/'
+SERVICE_NAME = 'jgroups'
+TRACKER_NAME = 'jgroups-tracker'
+NETWORK_NAME = 'jgroups_network'
+SUBNET = '172.113.0.0/16'
+
+
+def create_service(service_name, image, env=None, mounts=None, placement=None, replicas=1):
+    container_spec = types.ContainerSpec(image=image, env=env, mounts=mounts)
+    logger.debug(container_spec)
+    task_tmpl = types.TaskTemplate(container_spec,
+                                   resources=types.Resources(mem_limit=314572800),
+                                   restart_policy=types.RestartPolicy(),
+                                   placement=placement)
+    logger.debug(task_tmpl)
+    cli.create_service(task_tmpl, name=service_name, mode={'Replicated': {'Replicas': replicas}},
+                       networks=[{'Target': NETWORK_NAME}])
+
+
+def run_churn(time_to_start):
+    logger.debug('Time to start churn: {:d}'.format(time_to_start))
+    if args.synthetic:
+        logger.info(args.synthetic)
+        nodes_trace = NodesTrace(synthetic=args.synthetic)
+    else:
+        # Website02.db epoch starts 1st January 201. Exact formula obtained from Sebastien Vaucher
+        websites_epoch = 730753 + 1 + 86400. / (16 * 3600 + 11 * 60 + 10)
[…]

### 2017-01-18 projects/JGroupsTester/cluster/jgroups_churn.py

@@ -4,21 +4,9 @@ import logging
 import random
 import re
 import subprocess
+import threading
 
 from churn import Churn
-
-
-def get_peer_list(path='../data/*.txt'):
-    with open(glob.glob(path)[0], 'r') as f:
-        a_list = []
-        for line in f.readlines():
-            match = re.match(r'\d+ - View: (.+)', line)
-            if match:
-                a_list = match.group(1).split(',')
-                break
-        if not a_list:
-            raise LookupError('No view found in file {}'.format(f.name))
-        return a_list
 
 
 class JGroupsChurn(Churn):
@@ -28,7 +16,8 @@ class JGroupsChurn(Churn):
     A class in charge of adding/suspending nodes to create churn in a JGroups SEQUENCER cluster
     """
 
-    def __init__(self, hosts_filename=None, service_name='', repository='', kill_coordinator_round=None):
+    def __init__(self, hosts_filename=None, service_name='', repository='', kill_coordinator_round=None,
+                 delay=0, file_path=''):
         self.containers = {}
         self.coordinator = None
         self.peer_list = []
@@ -48,6 +37,7 @@ class JGroupsChurn(Churn):
             self.kill_coordinator_round = kill_coordinator_round
         self.kill_index = 0
         self.cluster_size = 0
+        threading.Timer(delay, self._get_peer_list, args=(file_path,)).start()
 
     def suspend_processes(self, to_suspend_nb):
         """
@@ -202,3 +192,17 @@ class JGroupsChurn(Churn):
 
         self.containers[host] = subprocess.check_output(command_ps,
                                                         universal_newlines=True).splitlines()
+
+    # TODO check where Error is caught
+    def _get_peer_list(self, path='../data/*.txt'):
+        with open(glob.glob(path)[0], 'r') as f:
+            a_list = []
+            for line in f.readlines():
+                match = re.match(r'\d+ - View: (.+)', line)
+                if match:
+                    a_list = match.group(1).split(',')
+                    break
+            if not a_list:
+                raise LookupError('No view found in file {}'.format(f.name))
+            self.peer_list = a_list
+            self.coordinator = a_list[0]

### 2017-01-19 projects/EpTOTester/cluster/benchmark.py
Continue work on framework

@@ -36,9 +36,11 @@ class Benchmark:
         self.use_tracker = use_tracker
 
     def run(self):
-        if args.local:
+        args.time_add *= 1000
+        args.time_to_run *= 1000
+        if self.local:
             service_image = self.config['service']['name']
-            if args.tracker:
+            if self.use_tracker:
                 tracker_image = self.config['tracker']['name']
             with subprocess.Popen(['../gradlew', '-p', '..', 'docker'],
                                   stdin=subprocess.PIPE, stdout=subprocess.PIPE, stderr=subprocess.PIPE,
@@ -68,6 +70,70 @@ class Benchmark:
                 if not self.cli.networks(names=[self.config['service']['network']['name']]):
                     self.logger.error('Network  doesn\'t exist!')
                     exit(1)
+
+        for run_nb, _ in enumerate(range(args.runs), 1):
+            if args.tracker:
+                create_service(APP_CONFIG['tracker']['name'], tracker_image,
+                               placement={'Constraints': ['node.role == manager']},
+                               mem_limit=APP_CONFIG['service']['mem_limit'])
+                wait_on_service(APP_CONFIG['tracker']['name'], 1)
+            time_to_start = int((time.time() * 1000) + args.time_add)
+            logger.debug(datetime.utcfromtimestamp(time_to_start / 1000).isoformat())
+
+            environment_vars = {**APP_CONFIG['service']['parameters'],
+                                **{'PEER_NUMBER': args.peer_number,
+                                   'TIME': time_to_start,
+                                   'TIME_TO_RUN': args.time_to_run}}
+            environment_vars = ['{:s}={}'.format(k, v) for k, v in environment_vars.items()]
+            logger.debug(environment_vars)
+
+            service_replicas = 0 if args.churn else args.peer_number
+            log_storage = LOCAL_DATA if args.local else CLUSTER_DATA
+            create_service(APP_CONFIG['service']['name'], service_image, env=environment_vars,
+                           mounts=[types.Mount(target='/data', source=log_storage, type='bind')],
+                           replicas=service_replicas, mem_limit=APP_CONFIG['service']['mem_limit'])
+
+            logger.info('Running Benchmark -> Experiment: {:d}/{:d}'.format(run_nb, args.runs))
+            if args.churn:
+                thread = threading.Thread(target=run_churn, args=[time_to_start + args.delay], daemon=True)
+                thread.start()
[…]

### 2017-01-19 projects/EpTOTester/cluster/run_benchmarks.py

@@ -96,8 +96,6 @@ if __name__ == '__main__':
     benchmark = Benchmark(APP_CONFIG, args.local, log_level, churn)
     benchmark.set_logger_level(log_level)
     benchmark.run()
-    args.time_add *= 1000
-    args.time_to_run *= 1000
 
     def signal_handler(signal, frame):
         logger.info('Stopping Benchmarks')
@@ -106,100 +104,6 @@ if __name__ == '__main__':
     signal.signal(signal.SIGINT, signal_handler)
     signal.signal(signal.SIGTERM, signal_handler)
 
-    if args.local:
-        service_image = APP_CONFIG['service']['name']
-        if args.tracker:
-            tracker_image = APP_CONFIG['tracker']['name']
-        with subprocess.Popen(['../gradlew', '-p', '..', 'docker'],
-                              stdin=subprocess.PIPE, stdout=subprocess.PIPE, stderr=subprocess.PIPE,
-                              universal_newlines=True) as p:
-            for line in p.stdout:
-                print(line, end='')
-    else:
-        service_image = APP_CONFIG['repository'] + APP_CONFIG['service']['name']
-        for line in cli.pull(service_image, stream=True, decode=True):
-            print(line)
-        if args.tracker:
-            tracker_image = APP_CONFIG['repository'] + APP_CONFIG['tracker']['name']
-            for line in cli.pull(tracker_image, stream=True):
-                print(line)
-    try:
-        cli.init_swarm()
-        if not args.local:
-            logger.info('Joining Swarm on every hosts:')
-            token = cli.inspect_swarm()['JoinTokens']['Worker']
-            subprocess.call(['parallel-ssh', '-t', '0', '-h', 'hosts', 'docker', 'swarm',
-                             'join', '--token', token, '{:s}:2377'.format(MANAGER_IP)])
-        ipam_pool = utils.create_ipam_pool(subnet=APP_CONFIG['service']['network']['subnet'])
-        ipam_config = utils.create_ipam_config(pool_configs=[ipam_pool])
-        cli.create_network(APP_CONFIG['service']['network']['name'], 'overlay', ipam=ipam_config)
-    except errors.APIError:
-        logger.info('Host is already part of a swarm')
-        if not cli.networks(names=[APP_CONFIG['service']['network']['name']]):
-            logger.error('Network  doesn\'t exist!')
-            exit(1)
 
-    for run_nb, _ in enumerate(range(args.runs), 1):
-        if args.tracker:
[…]

### 2017-01-19 projects/JGroupsTester/cluster/run_benchmarks.py

@@ -17,8 +17,7 @@ import threading
 import time
 import yaml
 
-from churn import Churn
-from churn import get_peer_list
+from jgroups_churn import JGroupsChurn
 from datetime import datetime
 from docker import errors
 from docker import types
@@ -31,6 +30,7 @@ MANAGER_IP = '172.16.2.119'
 LOCAL_DATA = '/home/jocelyn/tmp/data'
 CLUSTER_DATA = '/home/debian/data'
 LOCAL_DATA_FILES = '{:s}/*.txt'.format(LOCAL_DATA)
+CLUSTER_DATA_FILES = '{:s}/*.txt'.format(CLUSTER_DATA)
 DISTANT_REPOSITORY = 'swarm-m:5000/'
 SERVICE_NAME = 'jgroups'
 TRACKER_NAME = 'jgroups-tracker'
@@ -69,32 +69,26 @@ def run_churn(time_to_start):
         repository = DISTANT_REPOSITORY
 
     delta = args.delta
-    logging.debug("Kill_coordinator = {}".format(args.kill_coordinator))
-    churn = Churn(hosts_filename=hosts_fname, service_name=SERVICE_NAME,
-                  repository=repository, kill_coordinator_round=args.kill_coordinator)
+    if args.local:
+        the_path = LOCAL_DATA_FILES
+    else:
+        the_path = CLUSTER_DATA_FILES
+    delay = int((time_to_start - (time.time() * 1000)) / 1000)
+    # TODO be able to use the right class here
+    churn = JGroupsChurn(hosts_filename=hosts_fname, service_name=SERVICE_NAME,
+                  repository=repository, kill_coordinator_round=args.kill_coordinator,
+                  delay=delay, file_path=the_path)
+
     churn.set_logger_level(log_level)
-    logger.debug("Kill coordinator rounds: {}".format(churn.kill_coordinator_round))
 
     # Add initial cluster
     logger.debug('Initial size: {}'.format(nodes_trace.initial_size()))
     churn.add_processes(nodes_trace.initial_size())
-    delay = int((time_to_start - (time.time() * 1000)) / 1000)
     logger.debug('Delay: {:d}'.format(delay))
     logger.info('Starting churn at {:s} UTC'
                 .format(datetime.utcfromtimestamp(time_to_start // 1000).isoformat()))
     time.sleep(delay)
     logger.info('Starting churn')
-    try:
-        if args.local:
-            churn.peer_list = get_peer_list(LOCAL_DATA_FILES)
-        else:
[…]

### 2017-01-19 report/definitions.tex
Change cluster parameters to evaluation section

@@
 \label{sec:definitions}
+\jt{I think event should stay in this section what do you think?}
 A process or peer is defined as an actor in our system running the application that needs total order.
@@
 Please reread.}\jt{Done.}
-\par
-We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second.
-We tested \epto and \jgroups with three different cluster parameters:
-\begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
-	\item[\textbf{$(50,50)$}:] 50 peers with a global event throughput of 50 events per second.
-	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
-	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
-\end{description}

### 2017-01-19 report/evaluation.tex
Change cluster parameters to evaluation section

@@
 We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
+\par
+We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second.
+We tested \epto and \jgroups with three different cluster parameters:
+\begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
+	\item[\textbf{$(50,50)$}:] 50 peers with a global event throughput of 50 events per second.
+	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
+	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
+\end{description}
 \par 

### 2017-01-24 projects/EpTOTester/lsdssuite/lsdssuite/benchmark.py
Upgrade to new LSDSSuite version

@@ -1,3 +1,4 @@
+# coding: utf-8
 import logging
 import re
 import subprocess
@@ -21,8 +22,9 @@ class Benchmark:
 
     def __init__(self, app_config, cluster_config, local, use_tracker,
                  churn=None,
-                 cli=docker.Client(base_url='unix://var/run/docker.sock')):
-        self.cli = cli
+                 client=docker.DockerClient(
+                     base_url='unix://var/run/docker.sock')):
+        self.client = client
         self.app_config = app_config
         self.cluster_config = cluster_config
         self.local = local
@@ -55,18 +57,20 @@ class Benchmark:
         else:
             service_image = (self.app_config['repository']['name']
                              + self.app_config['service']['name'])
-            for line in self.cli.pull(service_image, stream=True, decode=True):
+            for line in self.client.images.pull(service_image, stream=True,
+                                                decode=True):
                 print(line)
             if self.use_tracker:
                 tracker_image = (self.app_config['repository']['name']
                                  + self.app_config['tracker']['name'])
-                for line in self.cli.pull(tracker_image, stream=True):
+                for line in self.client.images.pull(tracker_image,
+                                                    stream=True):
                     print(line)
             try:
-                self.cli.init_swarm()
+                self.client.swarm.init()
                 if not self.local:
                     self.logger.info('Joining Swarm on every hosts:')
-                    token = self.cli.inspect_swarm()['JoinTokens']['Worker']
+                    token = self.client.swarm.attrs['JoinTokens']['Worker']
                     subprocess.call(
                         [
                             'parallel-ssh',
@@ -85,12 +89,12 @@ class Benchmark:
                     subnet=self.app_config['service']['network']['subnet'])
                 ipam_config = utils.create_ipam_config(
                     pool_configs=[ipam_pool])
-                self.cli.create_network(
+                self.client.networks.create(
                     self.app_config['service']['network']['name'],
-                    'overlay', ipam=ipam_config)
+                    driver='overlay', ipam=ipam_config)
             except errors.APIError:
                 self.logger.info('Host is already part of a swarm')
-                if not self.cli.networks(
+                if not self.client.networks.list(
                         names=[self.app_config['service']['network']['name']]):
                     self.logger.error('Network  doesn\'t exist!')
                     exit(1)
@@ -100,8 +104,7 @@ class Benchmark:
                 self._create_service(
                     self.app_config['tracker']['name'],
                     tracker_image,
[…]

### 2017-01-24 projects/EpTOTester/lsdssuite/lsdssuite/churn.py

@@ -1,7 +1,9 @@
+# coding: utf-8
 import logging
 
 import random
 import subprocess
+import yaml
 
 
 class Churn:
@@ -12,7 +14,7 @@ class Churn:
     """
 
     def __init__(self, hosts_filename=None, service_name='', repository='',
-                 period=5000, delay=0, synthetic=None):
+                 period=5000, delay=0, synthetic=False):
         self.containers = {}
         self.peer_list = []
         self.logger = logging.getLogger('churn')
@@ -23,6 +25,8 @@ class Churn:
         self.period = period
         self.delay = delay
         self.synthetic = synthetic
+        with open('config/churn.yaml', 'r') as f:
+            self.churn_params = yaml.load(f)
         self.hosts = ['localhost']
         if hosts_filename is not None:
             with open(hosts_filename, 'r') as file:

### 2017-01-24 projects/EpTOTester/lsdssuite/run_benchmarks.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python3
+# coding: utf-8
 """
 Author: Jocelyn Thode
 
@@ -27,14 +28,6 @@ def create_logger():
         logging.config.dictConfig(conf)
 
 
-def churn_tuple(s):
-    try:
-        _to_kill, _to_create = map(int, s.split(','))
-        return _to_kill, _to_create
-    except:
-        raise TypeError("Tuples must be (int, int)")
-
-
 if __name__ == '__main__':
     parser = argparse.ArgumentParser(
         description='Run benchmarks',
@@ -58,11 +51,6 @@ if __name__ == '__main__':
     parser.add_argument('-l', '--local', action='store_true',
                         help='Run locally')
     parser.add_argument(
-        '-t',
-        '--tracker',
-        action='store_true',
-        help='Specify whether the app uses a tracker')
-    parser.add_argument(
         '-n',
         '--runs',
         type=int,
@@ -85,11 +73,8 @@ if __name__ == '__main__':
     churn_parser.add_argument(
         '--synthetic',
         '-s',
-        metavar='N',
-        type=churn_tuple,
-        nargs='+',
-        help='Pass the synthetic list (to_kill,to_create)'
-             '(example: 0,100 0,1 1,0)')
+        action='store_true',
+        help='Whether to use synthetic churn or not')
     churn_parser.add_argument(
         '--delay',
         '-d',
@@ -99,7 +84,8 @@ if __name__ == '__main__':
              'should the tester start in ms')
 
     args = parser.parse_args()
-    APP_CONFIG = yaml.load(args.config)
+    APP_CONFIG = yaml.load(args.app_config)
+    use_tracker = True if 'tracker' in APP_CONFIG else False
 
     if args.verbose:
         log_level = logging.DEBUG
@@ -132,13 +118,13 @@ if __name__ == '__main__':
         APP_CONFIG,
         CLUSTER_PARAMETERS,
         args.local,
-        args.tracker,
+        use_tracker,
         churn)
     benchmark.set_logger_level(log_level)
 
     def signal_handler(signal, frame):
         logger.info('Stopping Benchmarks')
-        benchmark.stop()
+        benchmark.stop(is_signal=True)
         exit(0)
 
     signal.signal(signal.SIGINT, signal_handler)

### 2017-01-24 projects/JGroupsTester/lsdssuite/jgroups_churn.py

@@ -16,7 +16,7 @@ class JGroupsChurn(Churn):
 
     def __init__(self, hosts_filename=None, service_name='', repository='',
                  period=5000, delay=0, synthetic=None,
-                 file_path='', kill_coordinator_round=None):
+                 file_path='', kill_coordinator_rounds=None):
         super().__init__(hosts_filename, service_name, repository, period,
                          delay, synthetic)
         self.coordinator = None
@@ -24,10 +24,10 @@ class JGroupsChurn(Churn):
         self.periods = 0
         self._file_path = file_path
 
-        if kill_coordinator_round is None:
-            self.kill_coordinator_round = []
+        if kill_coordinator_rounds is None:
+            self.kill_coordinator_rounds = []
         else:
-            self.kill_coordinator_round = kill_coordinator_round
+            self.kill_coordinator_rounds = kill_coordinator_rounds
         self.kill_index = 0
 
     def suspend_processes(self, to_suspend_nb):
@@ -48,11 +48,11 @@ class JGroupsChurn(Churn):
             self.logger.debug(
                 "Variables: {} {}".format(
                     self.periods,
-                    self.kill_coordinator_round))
+                    self.kill_coordinator_rounds))
             # If we missed kill period kill the coord at the next chance
-            if (self.periods in self.kill_coordinator_round
-                or (self.kill_index < len(self.kill_coordinator_round)
-                    and self.periods > self.kill_coordinator_round[self.kill_index])):
+            if (self.periods in self.kill_coordinator_rounds
+                or (self.kill_index < len(self.kill_coordinator_rounds)
+                    and self.periods > self.kill_coordinator_rounds[self.kill_index])):
                 self.logger.info("Killing coordinator")
                 command_suspend += [self.coordinator]
                 for host in self.hosts:

### 2017-01-24 projects/JGroupsTester/lsdssuite/lsdssuite/benchmark.py

@@ -1,3 +1,4 @@
+# coding: utf-8
 import logging
 import re
 import subprocess
@@ -21,8 +22,9 @@ class Benchmark:
 
     def __init__(self, app_config, cluster_config, local, use_tracker,
                  churn=None,
-                 cli=docker.Client(base_url='unix://var/run/docker.sock')):
-        self.cli = cli
+                 client=docker.DockerClient(
+                     base_url='unix://var/run/docker.sock')):
+        self.client = client
         self.app_config = app_config
         self.cluster_config = cluster_config
         self.local = local
@@ -55,18 +57,20 @@ class Benchmark:
         else:
             service_image = (self.app_config['repository']['name']
                              + self.app_config['service']['name'])
-            for line in self.cli.pull(service_image, stream=True, decode=True):
+            for line in self.client.images.pull(service_image, stream=True,
+                                                decode=True):
                 print(line)
             if self.use_tracker:
                 tracker_image = (self.app_config['repository']['name']
                                  + self.app_config['tracker']['name'])
-                for line in self.cli.pull(tracker_image, stream=True):
+                for line in self.client.images.pull(tracker_image,
+                                                    stream=True):
                     print(line)
             try:
-                self.cli.init_swarm()
+                self.client.swarm.init()
                 if not self.local:
                     self.logger.info('Joining Swarm on every hosts:')
-                    token = self.cli.inspect_swarm()['JoinTokens']['Worker']
+                    token = self.client.swarm.attrs['JoinTokens']['Worker']
                     subprocess.call(
                         [
                             'parallel-ssh',
@@ -85,12 +89,12 @@ class Benchmark:
                     subnet=self.app_config['service']['network']['subnet'])
                 ipam_config = utils.create_ipam_config(
                     pool_configs=[ipam_pool])
-                self.cli.create_network(
+                self.client.networks.create(
                     self.app_config['service']['network']['name'],
-                    'overlay', ipam=ipam_config)
+                    driver='overlay', ipam=ipam_config)
             except errors.APIError:
                 self.logger.info('Host is already part of a swarm')
-                if not self.cli.networks(
+                if not self.client.networks.list(
                         names=[self.app_config['service']['network']['name']]):
                     self.logger.error('Network  doesn\'t exist!')
                     exit(1)
@@ -100,8 +104,7 @@ class Benchmark:
                 self._create_service(
                     self.app_config['tracker']['name'],
                     tracker_image,
-                    placement={
[…]

### 2017-01-24 projects/JGroupsTester/lsdssuite/lsdssuite/nodes_trace.py

@@ -1,5 +1,4 @@
-#!/usr/bin/env python3
-
+# coding: utf-8
 import sqlite3
 import time
 

### 2017-01-24 projects/JGroupsTester/lsdssuite/run_benchmarks.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python3
+# coding: utf-8
 """
 Author: Jocelyn Thode
 
@@ -27,14 +28,6 @@ def create_logger():
         logging.config.dictConfig(conf)
 
 
-def churn_tuple(s):
-    try:
-        _to_kill, _to_create = map(int, s.split(','))
-        return _to_kill, _to_create
-    except:
-        raise TypeError("Tuples must be (int, int)")
-
-
 if __name__ == '__main__':
     parser = argparse.ArgumentParser(
         description='Run benchmarks',
@@ -58,11 +51,6 @@ if __name__ == '__main__':
     parser.add_argument('-l', '--local', action='store_true',
                         help='Run locally')
     parser.add_argument(
-        '-t',
-        '--tracker',
-        action='store_true',
-        help='Specify whether the app uses a tracker')
-    parser.add_argument(
         '-n',
         '--runs',
         type=int,
@@ -85,11 +73,8 @@ if __name__ == '__main__':
     churn_parser.add_argument(
         '--synthetic',
         '-s',
-        metavar='N',
-        type=churn_tuple,
-        nargs='+',
-        help='Pass the synthetic list (to_kill,to_create)'
-             '(example: 0,100 0,1 1,0)')
+        action='store_true',
+        help='Whether to use synthetic churn or not')
     churn_parser.add_argument(
         '--delay',
         '-d',
@@ -100,6 +85,7 @@ if __name__ == '__main__':
 
     args = parser.parse_args()
     APP_CONFIG = yaml.load(args.app_config)
+    use_tracker = True if 'tracker' in APP_CONFIG else False
 
     if args.verbose:
         log_level = logging.DEBUG
@@ -128,8 +114,8 @@ if __name__ == '__main__':
             delay=args.delay,
             synthetic=args.synthetic,
             file_path=file_path,
-            kill_coordinator_round=APP_CONFIG['service']
-            ['special_parameters']['kill_coordinator_round'])
+            kill_coordinator_rounds=APP_CONFIG['service']
+            ['special_parameters']['kill_coordinator_rounds'])
         churn.set_logger_level(log_level)
     else:
         churn = None
@@ -137,13 +123,13 @@ if __name__ == '__main__':
         APP_CONFIG,
         CLUSTER_PARAMETERS,
         args.local,
-        args.tracker,
+        use_tracker,
         churn)
     benchmark.set_logger_level(log_level)
 
[…]

### 2017-01-24 projects/EpTOTester/lsdssuite/lsdssuite/benchmark.py
Fix bugs

@@ -57,15 +57,13 @@ class Benchmark:
         else:
             service_image = (self.app_config['repository']['name']
                              + self.app_config['service']['name'])
-            for line in self.client.images.pull(service_image, stream=True,
-                                                decode=True):
-                print(line)
+            self.logger.info(self.client.images.pull(service_image))
+
             if self.use_tracker:
                 tracker_image = (self.app_config['repository']['name']
                                  + self.app_config['tracker']['name'])
-                for line in self.client.images.pull(tracker_image,
-                                                    stream=True):
-                    print(line)
+                self.logger.info(self.client.images.pull(tracker_image))
+
             try:
                 self.client.swarm.init()
                 if not self.local:
@@ -77,7 +75,7 @@ class Benchmark:
                             '-t',
                             '0',
                             '-h',
-                            'hosts',
+                            'config/hosts',
                             'docker',
                             'swarm',
                             'join',
@@ -190,7 +188,8 @@ class Benchmark:
 
             if not self.local:
                 subprocess.call(
-                    'parallel-ssh -t 0 -h hosts "mkdir -p {path}/test-{nb}/capture &&'
+                    'parallel-ssh -t 0 -h config/hosts'
+                    ' "mkdir -p {path}/test-{nb}/capture &&'
                     ' mv {path}/*.txt {path}/test-{nb}/ &&'
                     ' mv {path}/capture/*.csv {path}/test-{nb}/capture/"'
                     .format(path=self.cluster_config['cluster_data'],
@@ -223,7 +222,7 @@ class Benchmark:
             self.client.api.remove_service(self.app_config['service']['name'])
             if not self.local and is_signal:
                 time.sleep(15)
-                with open('hosts', 'r') as file:
+                with open('config/hosts', 'r') as file:
                     for host in file.read().splitlines():
                         subprocess.call('rsync --remove-source-files '
                                         '-av {:s}:{:s}/*.txt ../data'

### 2017-01-24 projects/EpTOTester/setup_cluster.sh

@@ -22,4 +22,4 @@ docker push localhost:5000/epto-tracker:latest
 
 kill ${ssh_pid}
 
-rsync -av --copy-links cluster/ ${USERNAME}@${MANAGER_IP}:~/epto
+rsync -av --copy-links lsdssuite/ ${USERNAME}@${MANAGER_IP}:~/epto

### 2017-01-24 projects/JGroupsTester/lsdssuite/lsdssuite/benchmark.py

@@ -57,15 +57,13 @@ class Benchmark:
         else:
             service_image = (self.app_config['repository']['name']
                              + self.app_config['service']['name'])
-            for line in self.client.images.pull(service_image, stream=True,
-                                                decode=True):
-                print(line)
+            self.logger.info(self.client.images.pull(service_image))
+
             if self.use_tracker:
                 tracker_image = (self.app_config['repository']['name']
                                  + self.app_config['tracker']['name'])
-                for line in self.client.images.pull(tracker_image,
-                                                    stream=True):
-                    print(line)
+                self.logger.info(self.client.images.pull(tracker_image))
+
             try:
                 self.client.swarm.init()
                 if not self.local:
@@ -77,7 +75,7 @@ class Benchmark:
                             '-t',
                             '0',
                             '-h',
-                            'hosts',
+                            'config/hosts',
                             'docker',
                             'swarm',
                             'join',
@@ -190,7 +188,8 @@ class Benchmark:
 
             if not self.local:
                 subprocess.call(
-                    'parallel-ssh -t 0 -h hosts "mkdir -p {path}/test-{nb}/capture &&'
+                    'parallel-ssh -t 0 -h config/hosts'
+                    ' "mkdir -p {path}/test-{nb}/capture &&'
                     ' mv {path}/*.txt {path}/test-{nb}/ &&'
                     ' mv {path}/capture/*.csv {path}/test-{nb}/capture/"'
                         .format(path=self.cluster_config['cluster_data'],
@@ -223,7 +222,7 @@ class Benchmark:
             self.client.api.remove_service(self.app_config['service']['name'])
             if not self.local and is_signal:
                 time.sleep(15)
-                with open('hosts', 'r') as file:
+                with open('config/hosts', 'r') as file:
                     for host in file.read().splitlines():
                         subprocess.call('rsync --remove-source-files '
                                         '-av {:s}:{:s}/*.txt ../data'

### 2017-01-24 projects/JGroupsTester/setup_cluster.sh

@@ -21,4 +21,4 @@ docker push localhost:5000/jgroups-tracker:latest
 
 kill ${ssh_pid}
 
-rsync -av --copy-links cluster/ ${USERNAME}@${MANAGER_IP}:~/jgroups
+rsync -av --copy-links lsdssuite/ ${USERNAME}@${MANAGER_IP}:~/jgroups

### 2017-01-24 report/framework.tex
Add more related work and improve framework part

@@
 We use \eptotester to assess the claims made in \autocite{matos2015epto}.
-Although \epto's implementation is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
+Although the implementation of \epto is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
+We  use \eptotester instead of other solutions such as SPlay, which only supports LUA, as we need to test existing protocols written in different programming languages.
 \subsection{Implementation}
 We implement the network layer using UDP and our own PSS CYCLON operating on its own port to obtain a random view of peers.
-To obtain an initial random view, we contact an independent tracker implemented as a simple Python web server that keeps track of dead and alive nodes in the system.
+To obtain an initial random view, we contact an independent tracker implemented as a simple Python web server that keeps track of dead and alive nodes in the system using GET requests.
 We want to emphasize that the tracker is not required.
@@
 We use \jgroups 3.6.11.
-When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers, as Docker services do not yet support MULTICAST.
+When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library as a tracker instead of the traditional MULTICAST option to coordinate peers, as Docker services do not yet support MULTICAST.
 This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
@@
 \footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
-To deploy \epto effectively we need two different Docker Swarm services: one containing the tracker and one containing all \epto replicas.
+To deploy a protocol in our framework we need two different Docker Swarm services: one containing the tracker and one containing all replicas.
 Both of these services use the same network overlay to communicate.
@@
 This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
-All \epto parameters are customizable by using options provided through this script.
+Application parameters are customizable through a YAML configuration file.
+Tracker and replica communication is up to the user.
 Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.
 Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node.
-\jgroups deployment is much the same.
-The entire \eptotester architecture is shown in \autoref{fig:complete-architecture}.
+\jgroups uses the same deployment method.
+The \eptotester architecture is shown in \autoref{fig:complete-architecture}.
 \subsection{Fault Injection}
@@
 The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
+
+Synthetic churn follows strict periods.
+Every period of time, it modifies the cluster size according to the instructions, until there are no more steps.
+These instructions can either, do nothing, increase the cluster size, reduce the cluster size or both at the same time.
+
+Real traces are based on FTA traces as stated earlier.
+Every time we query the database and obtain every cluster change that happened since the last query according to the trace.
+Thus the smaller the period, the more accurate it is regarding to the real trace.
+\subsection{Extensions}
+\eptotester is a general framework and should satisfy most decentralized applications needs with almost no tweaking.
+However, some protocols are centralized.
+This is the case for \jgroups.
+\jgroups relies on MULTICAST and has a coordinator that order every events sent.
+\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need.
+In our experiments we wrote an extension of the Churn module, so that it can deal with the coordinator, killing it only when we want to.
+This provides us with fine grain control over the benchmarks.
+
+\eptotester supports the use of a tracker as a separate service.
+If the application using the framework needs a tracker it only has to provide the docker image name for the tracker in the application configuration YAML file.

### 2017-01-24 report/ordering.tex
Add more related work and improve framework part

--- 2017-01-12 report/ordering.tex
@@
 \subsection{Related Work}
-Related work concerning Total Order is already extensively covered in \autocite{matos2015epto}.
+Deterministic total order algorithms are analyzed extensively in \autocite{defago2004total}.
+
+One probabilistic total order algorithm is \textit{pbcast} \autocite{pbcast1996}.
+It disseminates events in the same manner as \epto.
+However, it uses a synchronous model and has weaker dissemination guarantees than \epto.
+These shortcomings are addressed in \autocite{birman1999bimodal} at the cost of delivering events in FIFO order, thus losing its total ordering properties.
+
+Another protocol named \textit{PABCast} \autocite{felber2002} provides almost the same properties as \epto.
+However \textit{PABCast} only offers probabilistic order whereas \epto offers deterministic order.
+Furthermore more, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages.
+This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases.
 
@@
 Unfortunately, SPlay only allows for LUA applications to run.
-\eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container .
+\eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container.
+This makes it easier for existing protocol implementations to be benchmarked using our framework.

### 2017-01-25 report/abstract.tex
Rework framework and background sections

--- 2016-12-30 report/abstract.tex
@@
   \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
-In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
+In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
 We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.

### 2017-01-25 report/background.tex
Rework framework and background sections

--- 2017-01-24 report/ordering.tex
@@
 However \textit{PABCast} only offers probabilistic order whereas \epto offers deterministic order.
-Furthermore more, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages.
+Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages.
 This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases.

### 2017-01-25 report/evaluation.tex
Rework framework and background sections

@@
 We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second.
-We tested \epto and \jgroups with three different cluster parameters:
+We test \epto and \jgroups with three different cluster parameters:
 \begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
@@
 \par 
-We used specific settings recommended by \jgroups for a big cluster \autocite{udp-largecluster}.
+We use specific settings recommended by \jgroups for a big cluster \autocite{udp-largecluster}.
 \par

### 2017-01-25 report/framework.tex
Change a sentence
Rework framework and background sections

@@
 We  use \eptotester instead of other solutions such as SPlay, which only supports LUA, as we need to test existing protocols written in different programming languages.
-\subsection{Implementation}
+\subsection{Deployment}
+\begin{figure}[htp]
+	\centering
+	\includegraphics[width=\linewidth]{figures/complete-architecture.pdf}
+	\vspace{-2mm} 
+	\caption[Caption]{\epto using the \eptotester framework\footnotemark}
+	\vspace{-2mm} 
+	\label{fig:complete-architecture}
+\end{figure}
+\footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
+To deploy a protocol in our framework we need two different Docker Swarm services: one containing the tracker and one containing all replicas.
+Both of these services use the same network overlay to communicate.
+This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
+This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
+Every benchmark is started through a Python script available on the master node.
+This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results through the framework python classes.
+Application, churn and cluster  parameters are customizable through YAML configuration files.
+Tracker and replica communication is up to the user.
+For example, \epto uses HTTP, while \jgroups uses TCP directly.
+Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.
+Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking framework on the master node.
+\jgroups uses the same deployment method.
+Our framework applied to \epto can be seen in \autoref{fig:complete-architecture}.
+\subsection{Fault Injection}
+Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}.
+The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
+
+Synthetic churn follows strict periods.
+Every period of time, it modifies the cluster size according to the instructions, until there are no more steps.
+These instructions can either, do nothing, increase the cluster size, reduce the cluster size or both at the same time.
+
+Real traces are based on FTA traces as stated earlier.
+Every time we query the database, we obtain every cluster change that happened since the last query according to the trace.
+Thus the smaller the period, the more accurate it is regarding to the real trace.
+\subsection{Extensions}
+\eptotester is a general framework and should satisfy most decentralized applications needs with almost no tweaking.
+However, some protocols are centralized.
+This is the case for \jgroups.
+\jgroups relies on MULTICAST and has a coordinator that order every events sent.
+\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need.
+In our experiments we wrote an extension of the Churn module, so that it can deal with the coordinator, killing it only when we want to.
+This provides us with fine-grained control over the benchmarks.
+
+\eptotester supports the use of a tracker as a separate service.
+If the application using the framework needs a tracker it only has to provide the docker image name for the tracker in the application configuration YAML file for the tracker to be created alongside the application service.
+\subsection{Protocols Implementation}
 We implement the network layer using UDP and our own PSS CYCLON operating on its own port to obtain a random view of peers.
@@
 We then subtract the bytes required for the IPv4 and UDP headers.
-\subsection{Deployment}
- \begin{figure}[htp]
- 	\centering
- 	\includegraphics[width=\linewidth]{figures/complete-architecture.pdf}
- 	\vspace{-2mm} 
- 	\caption[Caption]{\eptotester architecture\footnotemark}
- 	\vspace{-2mm} 
- 	\label{fig:complete-architecture}
- \end{figure}
-\footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
-To deploy a protocol in our framework we need two different Docker Swarm services: one containing the tracker and one containing all replicas.
-Both of these services use the same network overlay to communicate.
-This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
-This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
-Every benchmark is started through a Python script available on the master node.
-This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results.
-Application parameters are customizable through a YAML configuration file.
-Tracker and replica communication is up to the user.
-Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.
-Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node.
-\jgroups uses the same deployment method.
-The \eptotester architecture is shown in \autoref{fig:complete-architecture}.
-\subsection{Fault Injection}
-Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}.
-The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
 
-Synthetic churn follows strict periods.
-Every period of time, it modifies the cluster size according to the instructions, until there are no more steps.
-These instructions can either, do nothing, increase the cluster size, reduce the cluster size or both at the same time.
-
-Real traces are based on FTA traces as stated earlier.
-Every time we query the database and obtain every cluster change that happened since the last query according to the trace.
-Thus the smaller the period, the more accurate it is regarding to the real trace.
-\subsection{Extensions}
-\eptotester is a general framework and should satisfy most decentralized applications needs with almost no tweaking.
-However, some protocols are centralized.
-This is the case for \jgroups.
-\jgroups relies on MULTICAST and has a coordinator that order every events sent.
-\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need.
-In our experiments we wrote an extension of the Churn module, so that it can deal with the coordinator, killing it only when we want to.
-This provides us with fine grain control over the benchmarks.
-
-\eptotester supports the use of a tracker as a separate service.
-If the application using the framework needs a tracker it only has to provide the docker image name for the tracker in the application configuration YAML file.
+When benchmarking \jgroups we have to send a SIGSTOP to the containers instead of killing them, as otherwise the kernel closes the TCP connection to the TCPGOSSIP nicely, making the leave uneventful.
+SIGSTOP lets us simulate a harsh connection drop.

### 2017-02-02 projects/JGroupsTester/scripts/container-start-script.sh
Do not resolve domain name

@@ -1,11 +1,13 @@
 #!/usr/bin/env bash
 # This script needs to run in the container
 MY_IP_ADDR=$(/bin/hostname -i)
-TRACKER_IP=$(dig +short jgroups-tracker)
+TRACKER_IP=jgroups-tracker #$(dig +short jgroups-tracker)
 
 echo 'Starting Peer'
 MY_IP_ADDR=($MY_IP_ADDR)
 echo "${MY_IP_ADDR[0]}"
+
+mkdir -p /data/capture
 
 dstat_pid=0
 app_pid=0

### 2017-02-02 presentation/slides.tex
Add Initial Presentation (WIP)

\setbeamercovered{transparent}

\begin{document}

% -------------------------------------------------------------------------------
% -------------------------------------------------------------------------------

\begin{frame}[plain]

\titlepage

\end{frame}

% -------------------------------------------------------------------------------
% -------------------------------------------------------------------------------

\subtitle[Outline]{Outline}

\begin{frame}{Outline}
    \begin{itemize}
        \item Motivation
        \item \epto{} explained
        \item \sys
        \item Evaluation
        \item Conclusion
    \end{itemize}
\end{frame}

\subtitle[Introduction]{Introduction}

\begin{frame}{Motivation}
    \epto{} was only evaluated using a \textbf{simulation}.

    We need an evaluation with real peers to expose possible \textbf{limitations}.
\end{frame}

\begin{frame}{Motivation}
  \begin{itemize}
  \item Comparing \epto{} meant testing it against other algorithms
  \item No framework to easily deploy algorithms without having to rewrite them
  \end{itemize}

\end{frame}

\subtitle[Description]{Description}

\begin{frame}{What is \epto{}?}
	\textbf{Epidemic Total Order Algorithm:}
	\begin{itemize}
		\item Probabilistic dissemination algorithm
		\item Provides deterministic total order
		\item Scales in the number of peers and events
		\item Churn resistant
	\end{itemize}
\end{frame}

%\begin{frame}{What is \epto{}? (2)}
%	\textbf{Properties satisfied:}
%\begin{itemize}
%	\item Integrity
%	\item Validity
%	\item Total Order
%	\item Probabilistic Agreement
%\end{itemize}
%\end{frame}

%\begin{frame}{Properties satisfied}
%	\begin{itemize}
%		\item Integrity
%		\item Validity
%		\item Total Order
%		\item Probabilistic Agreement
%	\end{itemize}
%\end{frame}
\begin{frame}{\epto{} architecture}
    \begin{figure}
        \includegraphics[scale=0.5]{figures/architecture}
    \end{figure}
\tiny{M. Matos, H. Mercier, P. Felber, R. Oliveira and J. Pereira, “EpTO: An epidemic total order algorithm for Large-scale distributed systems”, in Proceedings of the 16th Annual Middleware Conference, ACM, 2015, pp. 100–111.
}
\end{frame}

\begin{frame}{\sys{} features}
    \begin{itemize}
    	\item Compatible with any distributed algorithm provided it runs on Docker
    	\item support for a user-provided tracker
    	\item Automated benchmarking execution
    	\item Containers allow for more than 1 peer per physical node
    	\item Can simulate churn or follow real traces
    \end{itemize}
\end{frame}

\begin{frame}{\sys{} architecture}
	\begin{figure}
		\includegraphics[scale=1.2]{complete-architecture}
	\end{figure}
\end{frame}

\begin{frame}{\sys{} Configuration and Logging}
	The protocol, churn and framework configuration is done through \textsc{YAML} files

	The protocols logs must be written in a file to be extracted to the host
\end{frame}

\subtitle[Evaluation]{Evaluation}

\begin{frame}{Evaluation}
    We evaluate \epto{} against \jgroups{} according scaling peers and global event throughput per second.

    We write $(n,e)$ where $n$ is the number of peers and $e$ is the global event throughput per second.
\end{frame}

\begin{frame}{Bandwidth}
	\begin{figure}
		\includegraphics[scale=0.6]{bandwidth-nochurn}
	\end{figure}
\end{frame}

\begin{frame}{Bandwidth Synthetic Churn}
	\begin{figure}
		\includegraphics[scale=0.68]{bandwidth-synth-churn}
	\end{figure}
\end{frame}

\begin{frame}{Bandwidth Real Churn}
	\begin{figure}
		\includegraphics[scale=0.65]{bandwidth-real-churn}
	\end{figure}
\end{frame}

\begin{frame}{Local Dissemination Stretch}
	\begin{figure}
		\includegraphics[scale=1]{local-diss-stretch-nochurn}
	\end{figure}
\end{frame}

\begin{frame}{Local Dissemination Stretch Synthetic Churn}
	\begin{figure}
		\includegraphics[scale=1]{local-diss-stretch-synth-churn}
	\end{figure}
\end{frame}

\begin{frame}{Local Dissemination Stretch Real Churn}
	\begin{figure}
		\includegraphics[scale=1]{local-diss-stretch-real-churn}
	\end{figure}
\end{frame}

\begin{frame}{Local Times}
	\begin{figure}
		\includegraphics[scale=0.70]{local-times-nochurn}
	\end{figure}
\end{frame}

\begin{frame}{Local Times Synthetic Churn}
	\begin{figure}
		\includegraphics[scale=1]{local-times-synth-churn}
	\end{figure}
\end{frame}

\begin{frame}{Local Times Real Churn}
	\begin{figure}
		\includegraphics[scale=1]{local-times-real-churn}
	\end{figure}
\end{frame}

\begin{frame}{Global Times}
	\begin{figure}
		\includegraphics[scale=0.60]{global-times-nochurn}
	\end{figure}
\end{frame}

\begin{frame}{Global Times Synthetic Churn}
	\begin{figure}
		\includegraphics[scale=1]{global-times-synth-churn}
	\end{figure}
\end{frame}

\begin{frame}{Global Times Real Churn}
	\begin{figure}
		\includegraphics[scale=1]{global-times-real-churn}
	\end{figure}
\end{frame}

\subtitle[Conclusion]{Conclusion}

\begin{frame}{Future Work}
    \begin{itemize}
        \item Obtain more resources to have stronger results
        \item Implement a Push-Pull \epto{} version
        \item Use Kubernetes instead of Docker + Docker swarm
        \item Refine Framework Architecture
    \end{itemize}
\end{frame}

\begin{frame}[standout]
	Questions?
\end{frame}

% -------------------------------------------------------------------------------
% -------------------------------------------------------------------------------

\end{document}

% -------------------------------------------------------------------------------

### 2017-02-03 presentation/slides.tex
Add slides and rework graphics

@@
 
+\section{Introduction}
 \subtitle[Introduction]{Introduction}
@@
 \begin{frame}{Motivation}
-    \epto{} was only evaluated using a \textbf{simulation}.
-
-    We need an evaluation with real peers to expose possible \textbf{limitations}.
+    \epto{} was only evaluated using a \textbf{simulation}. We need an evaluation with real peers to:
+    \begin{itemize}
+    	\item Expose possible \textbf{limitations}
+    	\item Confirm simulation \textbf{results}
+    \end{itemize} 
 \end{frame}
@@
   \item Comparing \epto{} meant testing it against other algorithms
-  \item No framework to easily deploy algorithms without having to rewrite them
+  \item No framework to easily benchmark algorithms without having to rewrite them
   \end{itemize}
@@
 		\item Probabilistic dissemination algorithm
+			\begin{itemize}
+				\item Using balls-and-bins
+			\end{itemize}
 		\item Provides deterministic total order
 		\item Scales in the number of peers and events
+			\begin{itemize}
+				\item Parameters increase logarithmically
+			\end{itemize}
 		\item Churn resistant
@@
 
+\section{Evaluation}
 \subtitle[Evaluation]{Evaluation}
@@
 	\begin{figure}
-		\includegraphics[scale=0.6]{bandwidth-nochurn}
+		\includegraphics[scale=0.9]{bandwidth-nochurn}
 	\end{figure}
@@
 
+\begin{frame}{Total \si{\giga\byte} sent/received}
+\begin{table}
+	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\begin{tabular}{llSSS}
+		\toprule
+		&& \multicolumn{3}{c}{Cluster parameters} \\
+		\cmidrule{3-5}
+		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
+		\midrule
+		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
+		&{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
+		\midrule
+		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
+		&{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
+		\bottomrule
+	\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Total \si{\giga\byte} sent/received}
+	\begin{table}
+		\centering
+		\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+			\begin{tabular}{lSSS}
+				\toprule
+				&& \multicolumn{2}{c}{Churn parameters} \\
+				\cmidrule{3-4}
+				{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
+				\midrule
+				\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
+				&{Sending}& 21.21(025) & 26.57(032)\\
+				\midrule
+				\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
+				\midrule
+				\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
+				\bottomrule
+			\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}{Total \si{\giga\byte} sent/received}
+\begin{table}[hpt]
+	\centering
+	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+	\begin{tabular}{lSS}
+		\toprule
+		&& \multicolumn{1}{c}{Churn parameters} \\
+		\cmidrule{3-3}
+		{Protocol}&& {Real Trace} \\
+		\midrule
+		\multirow{2}{*}{\epto}&{Receive}& 81.41(108)\\
+		&{Sending}& 82.67(108)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.56(008)\\
+		&{Sending}& 5.40(008)\\
+		\midrule
+		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.58(005)\\
+		&{Sending}& 5.43(005)\\
+		\bottomrule
+	\end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Total events sent in a stable environment}
+	\begin{table}
+		\centering
+		\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+		\begin{tabular}{lSSS}
+			\toprule
+			& \multicolumn{3}{c}{Cluster parameters} \\
+			\cmidrule{2-4}
+			Protocol & {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
+			\midrule
+			\epto & 59993.8(33) & 119898.2(97) & 59913.0(1643) \\
+			\jgroups & 59961.9(109) & 119885.7(50) & 60023.1(2871) \\
+			\bottomrule
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}{Total events sent with a synthetic churn}
+	\begin{table}
+		\centering
+		\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+		\begin{tabular}{lSSS}
+			\toprule
+			& \multicolumn{2}{c}{Cluster parameters} \\
+			\cmidrule{2-3}
+			Protocol & {1 kill/minute} & {1\{kill,add\}/minute} \\
+			\midrule
+			\epto & 53898.5(1339) & 59798.6(1401) \\
+			\jgroups-coord & 53834.7(1755) & 59507.9(2409) \\
+			\jgroups-nocoord & 53830.5(2003) & 59450.5(1751) \\
+			\bottomrule
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}{Total events sent during a real trace}
+	\begin{table}
+		\centering
+		\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 6, table-align-uncertainty}
+		\begin{tabular}{lS}
+			\toprule
+			Protocol & {Events sent}\\
+			\midrule
+			\epto & 165844.2(2102)\\
+			\jgroups-coord & 166183.0(13681)\\
+			\jgroups-nocoord & 166585.8(8249)\\
+			\bottomrule
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\section{Conclusion}
 \subtitle[Conclusion]{Conclusion}

### 2017-02-03 presentation/slides.tex
Move Global Times to Appendix
Use real appendix
Move other slides to appendix

@@
 
-\begin{frame}{Bandwidth Synthetic Churn}
+\begin{frame}{Bandwidth Synthetic Churn $(100,50)$}
 	\begin{figure}
@@
 
-\begin{frame}{Local Dissemination Stretch Synthetic Churn}
+\begin{frame}{Local Dissemination Stretch Synthetic Churn $(100,50)$}
 	\begin{figure}
@@
 
-\begin{frame}{Local Times Synthetic Churn}
+\begin{frame}{Local Times Synthetic Churn $(100,50)$}
 	\begin{figure}
@@
 		\includegraphics[scale=1]{local-times-real-churn}
-	\end{figure}
-\end{frame}
-
-\begin{frame}{Global Times}
-	\begin{figure}
-		\includegraphics[scale=0.60]{global-times-nochurn}
-	\end{figure}
-\end{frame}
-
-\begin{frame}{Global Times Synthetic Churn}
-	\begin{figure}
-		\includegraphics[scale=1]{global-times-synth-churn}
-	\end{figure}
-\end{frame}
-
-\begin{frame}{Global Times Real Churn}
-	\begin{figure}
-		\includegraphics[scale=1]{global-times-real-churn}
 	\end{figure}
@@
 
-\begin{frame}{Total \si{\giga\byte} sent/received}
-	\begin{table}
-		\centering
-		\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
-			\begin{tabular}{lSSS}
-				\toprule
-				&& \multicolumn{2}{c}{Churn parameters} \\
-				\cmidrule{3-4}
-				{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
-				\midrule
-				\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
-				&{Sending}& 21.21(025) & 26.57(032)\\
-				\midrule
-				\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
-				\midrule
-				\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
-				\bottomrule
-			\end{tabular}
-	\end{table}
-\end{frame}
-
-\begin{frame}{Total \si{\giga\byte} sent/received}
-\begin{table}[hpt]
-	\centering
-	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
-	\begin{tabular}{lSS}
-		\toprule
-		&& \multicolumn{1}{c}{Churn parameters} \\
-		\cmidrule{3-3}
-		{Protocol}&& {Real Trace} \\
-		\midrule
-		\multirow{2}{*}{\epto}&{Receive}& 81.41(108)\\
-		&{Sending}& 82.67(108)\\
-		\midrule
-		\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.56(008)\\
-		&{Sending}& 5.40(008)\\
-		\midrule
-		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.58(005)\\
-		&{Sending}& 5.43(005)\\
-		\bottomrule
-	\end{tabular}
-\end{table}
-\end{frame}
-
 \begin{frame}{Total events sent in a stable environment}
@@
 
-\begin{frame}{Total events sent with a synthetic churn}
+\begin{frame}{Total events sent with a synthetic churn $(100,50)$}
 	\begin{table}
@@
 
+\begin{frame}{Limitations}
+	\begin{itemize}
+		\item Difference not strong enough at 100 peers scale
+		\item High CPU usage
+		\item Docker problems on AWS/GCE
+	\end{itemize}
+\end{frame}
+
 \begin{frame}{Future Work}
@@
 
+\appendix
+
+\begin{frame}{Total \si{\giga\byte} sent/received Synthetic Churn $(100,50)$}
+	\begin{table}
+		\centering
+		\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+		\begin{tabular}{lSSS}
+			\toprule
+			&& \multicolumn{2}{c}{Churn parameters} \\
+			\cmidrule{3-4}
+			{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
+			\midrule
+			\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
+			&{Sending}& 21.21(025) & 26.57(032)\\
+			\midrule
+			\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
+			\midrule
+			\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
+			\bottomrule
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}{Total \si{\giga\byte} sent/received Real Churn}
+	\begin{table}[hpt]
+		\centering
+		\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
+		\begin{tabular}{lSS}
+			\toprule
+			&& \multicolumn{1}{c}{Churn parameters} \\
+			\cmidrule{3-3}
+			{Protocol}&& {Real Trace} \\
+			\midrule
+			\multirow{2}{*}{\epto}&{Receive}& 81.41(108)\\
+			&{Sending}& 82.67(108)\\
+			\midrule
+			\multirow{2}{*}{\jgroups-coord}&{Receive}& 5.56(008)\\
+			&{Sending}& 5.40(008)\\
+			\midrule
+			\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 5.58(005)\\
+			&{Sending}& 5.43(005)\\
+			\bottomrule
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}{Global Times}
+	\begin{figure}
+		\includegraphics[scale=0.65]{global-times-nochurn}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Global Times Synthetic Churn}
+	\begin{figure}
+		\includegraphics[scale=1]{global-times-synth-churn}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Global Times Real Churn}
+	\begin{figure}
+		\includegraphics[scale=1]{global-times-real-churn}
+	\end{figure}
+\end{frame}
+
 % -------------------------------------------------------------------------------

### 2017-02-03 presentation/slides.tex
Fix errors

@@
 		\item Provides deterministic total order
-		\item Scales in the number of peers and events
+		\item Scales wells with the number of peers
 			\begin{itemize}
@@
 \begin{frame}{Evaluation}
-    We evaluate \epto{} against \jgroups{} according scaling peers and global event throughput per second.
+    We evaluate \epto{} against \jgroups{}, scaling peers and global event throughput per second.
 

### 2017-02-06 report/abstract.tex
Correct typos in report

@@
 We then compare them using a synthetic churn trace and a real one using the same metrics.
-Our results show \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
+Our results show \epto is performing as expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
   \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}

### 2017-02-06 report/background.tex
Correct typos in report

@@
 \label{sec:background}
-Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system.
+Distributed systems and centralized systems alike need to preserve the temporal order of events produced by concurrent processes in the system.
 When there are separated processes that can only communicate through messages, you cannot easily order these messages.
@@
 Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages.
-This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases.
+This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases, making it less desirable than \epto.
 
 Both \autocite{Chandra2007} and \autocite{Maia2011} extensively test a particular large scale distributed system.
-However, they do not offer a framework to benchmark different systems.
-\autocite{Leonini2009} presents a framework named SPlay to deploy such applications.
+However, they do not offer a framework to benchmark different systems or use SPlay \autocite{Leonini2009}.
+SPlay is a framework used to deploy distributed applications.
 Unfortunately, SPlay only allows for LUA applications to run.
-\eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container.
+\eptotester offers a convenient framework to deploy distributed softare regardless of the programming language used, provided they run in a docker container.
 This makes it easier for existing protocol implementations to be benchmarked using our framework.

### 2017-02-06 report/conclusion.tex
Correct typos in report

@@
 
-In this paper, we implement \eptotester an architecture designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network.
+In this paper, we implement \eptotester a framework designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network.
 Furthermore, we implement \epto and evaluate the claims made in \autocite{matos2015epto} against \jgroups SEQUENCER, a total order algorithm relying on a coordinator.
@@
 We want to reiterate that this is expected as \epto is designed for large distributed networks.
-\jt{Let's write it this way for the moment if we are able to run something on GCE or AWS we can always change it later.}

### 2017-02-06 report/evaluation.tex
Correct typos in report

@@
 	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
-	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
+	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters are also used for all synthetic churns.
 \end{description}
@@
 Each VM uses Debian as its O/S.
-\eptotester is packaged as docker images.
+Protocols are packaged as docker images.
 We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services.
-Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
+Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be certain we will not have out of memory problems.
 \par
-\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}.
+\textbf{Experiment parameters.} Every benchmark except the ones following a real trace are run 10 times, each during \SI{20}{\minute}.
 When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}.
 The benchmarks following a real trace are run 5 times.
-The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}.
+The trace is sped-up 2 times, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}.
 Every benchmark run with churn is run with the $(100,50)$ parameters.
@@
 Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups.
-Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}.
+While \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally.
 Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.\mm{it would be perfect if we could experimentally show this.}
@@
 We have small peaks when adding a node to \SI{3}{\mbps} or less.
-Probably due to running the PSS initialization method on top of having one more node spreading rumors in the system.
+Probably due to running the PSS initialization method.
 This is confirmed at the end of the plot where \epto goes back to a normal Bandwidth after stabilization.
@@
 \label{sub:local-times}
-The local time represents the total time it takes for a peer to deliver every events sent.
+The local time represents the total time it takes for a peer to deliver every .
  \begin{figure*}[hpt]
@@
 \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}.
-We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.} \jt{Yes I though about it too.
+We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.
+Furthermore, in our test environment \jgroups is not limited by the bandwidth.
+If the coordinator would have a lower limit of bandwidth, the time to send all events when the coordinator dies would be longer.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.} \jt{Yes I though about it too.
 Unfortunately we can't make it work for the moment :/}
@@
 \subsection{Events sent}
-The variation observed in the different tests are due to the fact that we run an experiment for a period of time and that in some configurations a randomness decides whether an event is sent or not.
+The variation observed in the different tests are due to the fact that we use \textit{Thread.sleep()} between each send and that in some configurations a randomness decides whether an event is sent or not.
 \begin{table}[hpt]
@@
 \end{figure}
-In \autoref{fig:local-delta-real-churn}, \epto as worse outliers than before and so does \jgroups-coord.
+In \autoref{fig:local-delta-real-churn}, \epto has worse outliers than before and so does \jgroups-coord.
 \begin{table}[hpt]
@@
 	\toprule
-	Protocol &\\
+	Protocol & {Events sent}\\
 	\midrule
@@
 \subsection{Problems encountered}
-When running \jgroups we sometimes get failed runs, where either some peers will experience holes or peer will deliver out of order events.
-We measure  These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies.
+When running \jgroups we sometimes get failed runs, where either some peers will experience holes or a peer will deliver out of order events.
+These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies.
 These problems appeared 3 times out of 13 runs for \jgroups $1\{kill,add\}/minute$ and 2 times out of 7 for \jgroups with real churn.
@@
 As stated earlier, we specifically choose a $c$ high enough to never experience holes in \epto as \jgroups should never suffer from holes.
-If either a \jgroups or \epto fails to deliver every event sent during the run, the run is considered as having a problem.
+If either \jgroups or \epto fails to deliver every event sent during the run, the run is considered as having a problem.

### 2017-02-06 report/figures/no-churn/bandwidth-figure.tex
WIP report
Correct typos in report

@@
 legend cell align=left,
-legend style={at={(1.9,1.5)},anchor=north, font=\small, draw=none},]
+legend style={at={(1.8,1.5)},anchor=north, font=\small, draw=none, fill=none},]
 \nextgroupplot[ymax=4,ytick={0,1,2,3,4},
@@
 \node[anchor=south] at (group c3r1.north) {$(100,50)$};
-\node[anchor=south, rotate=-90] at (group c3r1.east){EpTO};
-\node[anchor=south, rotate=-90] at (group c3r2.east){JGroups};
+\node[anchor=south, rotate=-90] at (group c3r1.east){\epto};
+\node[anchor=south, rotate=-90] at (group c3r2.east){\jgroups};
 \end{tikzpicture}

### 2017-02-06 report/figures/no-churn/global-times.tex
WIP report
Correct typos in report

@@
 legend cell align=left,
-legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none},
+legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none, fill=none},
 cycle list name=my colors,
@@
 \addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tableglobaltime};
-\legend{EpTO, JGroups}
+\legend{\epto, \jgroups}
 %

### 2017-02-06 report/figures/no-churn/local-times.tex
WIP report
Correct typos in report

@@
 \addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaltime};
-\legend{EpTO, JGroups}
+\legend{\epto, \jgroups}
 %

### 2017-02-06 report/figures/no-churn/total-bytes.tex
Start working on report
Correct typos in report

--- 2016-12-01 report/figures/total-bytes.tex
@@
 	draw=none,
+	fill=none,
 },
@@
 	% Receive
-	(EpTO, 10.8956669416) +- (0, 0.112678096321221)
-	(JGroups, 0.6420599937) +- (0, 0.0164341768113215)
+	(EpTO, 10.8454953841) +- (0, 0.158556056880199)
+	(JGroups, 0.8112389025) +- (0, 0.0275185471028981)
 };
@@
 	% Send
-	(EpTO, 10.8952655703) +- (0, 0.112677849128705)
-	(JGroups, 0.628793349) +- (0, 0.0164560780074577)
+	(EpTO, 10.84508943053) +- (0, 0.158553532839913)
+	(JGroups, 0.7979795057) +- (0, 0.0275124778152395)
 };
@@
 	% Receive
-	(EpTO, 22.1900560357) +- (0, 0.513548491076014)
-	(JGroups, 1.1888473068) +- (0, 0.0236261886904278)
+	(EpTO, 22.3122950339) +- (0, 0.387040306037242)
+	(JGroups, 1.481260734) +- (0, 0.0112693613185708)
 };
@@
 	% Send
-	(EpTO, 22.1896583376) +- (0, 0.513548672987696)
-	(JGroups, 1.1756330318) +- (0, 0.0236412207168411)
+	(EpTO, 22.3118855827) +- (0, 0.387040491969926)
+	(JGroups, 1.4680794813) +- (0, 0.0112627440922006)
 };
@@
 	% Receive
-	(EpTO, 25.9895900054) +- (0, 0.285368750128887)
-	(JGroups, 1.6842673321) +- (0, 0.014352534074787)
+	(EpTO, 26.0194016553) +- (0, 0.2773178289900919968)
+	(JGroups, 2.0044813206) +- (0, 0.0103933483299930001)
 };
@@
 	% Send
-	(EpTO, 25.9864516777) +- (0, 0.285365521649361)
-	(JGroups, 1.6291726389) +- (0, 0.0124299318567327)
+	(EpTO, 26.0161708501) +- (0, 0.277318407393540023)
+	(JGroups, 1.9504773287) +- (0, 0.0104754176417391)
 };

### 2017-02-06 report/figures/no-churn/total-events-sent.tex
Start working on report
Correct typos in report

--- 2016-12-01 report/figures/total-events-sent.tex
@@
 	% Receive
-	(EpTO, 59994.9) +- (0, 2.76686746259348)
-	(JGroups, 59993) +- (0, 15.1217282962848)
+	(EpTO, 59993.8) +- (0, 3.32665998663345)
+	(JGroups, 59961.9) +- (0, 10.9285558667798)
 };
@@
 	% Receive
-	(EpTO, 119908.5) +- (0, 8.04501225742594)
-	(JGroups, 119924.7) +- (0, 5.69697775628018)
+	(EpTO, 119898.2) +- (0, 9.75021367287237)
+	(JGroups, 119885.7) +- (0, 5.01220732035492)
 };
@@
 	% Receive
-	(EpTO, 59993) +- (0, 237.827201509368)
-	(JGroups, 59974.6) +- (0, 216.060279654648)
+	(EpTO, 59913) +- (0, 164.332319131421)
+	(JGroups, 60023.1) +- (0, 287.150928181603)
 };

### 2017-02-06 report/figures/real-churn/bandwidth-figure-real-churn.tex
Correct typos in report

@@
 },
-stack plots=y,area style, enlarge x limits=false, 
 ymin=0,
@@
 legend cell align=left,
-legend style={at={(0.6,1.9)},anchor=north, font=\small, draw=none},]
+legend style={at={(0.5,1.9)},anchor=north, font=\small, draw=none, fill=none},]
 \nextgroupplot[height=2.3cm, 
@@
 \addplot[const plot, color=blue, mark=none,set layers,
-axis lines = middle,] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
+axis lines = middle, axis on top=false] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
 \nextgroupplot[ymax=4,ytick={0,1,2,3,4},
-yticklabels={0,1,2,3,4}]
+yticklabels={0,1,2,3,4},stack plots=y,area style, enlarge x limits=false, ]
 \addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
@@
 %
-\nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
+\nextgroupplot[stack plots=y,area style, enlarge x limits=false, ]
 \addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
@@
 %
-\nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
+\nextgroupplot[ymax=4,ytick={0,1,2,3,4},stack plots=y,area style, enlarge x limits=false, ]
 \addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;

### 2017-02-06 report/figures/real-churn/global-times-real-churn.tex
Correct typos in report

@@
 legend cell align=left,
-legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none},
+legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none, fill=none},
 cycle list name=my colors,

### 2017-02-06 report/figures/real-churn/local-delta-real-churn.tex
WIP report
Correct typos in report

--- 2016-12-22 report/figures/real-churn/local-delta-real-churn.tex
@@
 legend cell align=left,
-legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
+legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none, fill=none},
 cycle list name=my colors,
@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaldeltas};
-\legend{EpTO, JGroups-coord, JGroups-nocoord}
+\legend{\epto, \jgroups-coord, \jgroups-nocoord}
 \end{axis}

### 2017-02-06 report/figures/real-churn/total-events-sent-real-churn.tex
Correct typos in report

@@
 	draw=none,
+	fill=none
 },

### 2017-02-06 report/figures/synthetic-churn/global-times-churn.tex
WIP report
Correct typos in report

@@
 legend cell align=left,
-legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
+legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none, fill=none},
 cycle list name=my colors,
@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tableglobaltime};
-\legend{EpTO, JGroups-nocoord, JGroups-coord}
+\legend{\epto, \jgroups-nocoord, \jgroups-coord}
 %

### 2017-02-06 report/figures/synthetic-churn/local-delta-churn.tex
WIP report
Correct typos in report

@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tablelocaldeltas};
-\legend{EpTO, JGroups-coord, JGroups-nocoord}
+\legend{\epto, \jgroups-coord, \jgroups-nocoord}
 %

### 2017-02-06 report/figures/synthetic-churn/local-times-churn.tex
WIP report
Correct typos in report

@@
 \addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {\tablelocaltime};
-\legend{EpTO, JGroups-nocoord, JGroups-coord}
+\legend{\epto, \jgroups-nocoord, \jgroups-coord}
 %

### 2017-02-06 report/figures/synthetic-churn/total-bytes-churn.tex
Start working on report
Correct typos in report

--- 2016-12-01 report/figures/total-bytes-churn.tex
@@
 	% Receive
-	(EpTO, 21.3437665258) +- (0, 0.302297622101786)
-	(JGroups-nocoord, 1.3338886813) +- (0, 0.00853400183682598)
-	(JGroups-coord, 1.3440551501) +- (0, 0.0164341768113215)
+	(EpTO, 21.001445292) +- (0, 0.248116313343146)
+	(JGroups-nocoord, 1.5381357651) +- (0, 0.0115143286137117)
+	(JGroups-coord, 1.5510862161) +- (0, 0.0181726968637206)
 };
@@
 	% Send
-	(EpTO, 21.4504884001) +- (0, 0.0157182231373203)
-	(JGroups-nocoord, 1.285075604) +- (0, 0.00872334968295833)
-	(JGroups-coord, 1.2945710642) +- (0, 0.0158581561021704)
+	(EpTO, 21.216776898) +- (0, 0.255744716454951)
+	(JGroups-nocoord, 1.4935079775) +- (0, 0.0119366849923253)
+	(JGroups-coord, 1.5064716889) +- (0, 0.0186384151121228)
 };
@@
 	% Receive
-	(EpTO, 27.0797668262) +- (0, 0.300652054138594)
-	(JGroups-nocoord, 1.6071496365) +- (0, 0.00972747186106245)
-	(JGroups-coord, 1.6039433255) +- (0, 0.0084962148465832)
+	(EpTO, 26.325256447) +- (0, 0.320079175714576)
+	(JGroups-nocoord, 1.8338254834) +- (0, 0.0187827622384514)
+	(JGroups-coord, 1.8511940589) +- (0, 0.0183913437417294)
 };
@@
 	% Send
-	(EpTO, 26.8856375946) +- (0, 0.0157182231373203)
-	(JGroups-nocoord, 1.5459290852) +- (0, 0.00937891633596378)
-	(JGroups-coord, 1.5426003209) +- (0, 0.00829044483237941)
+	(EpTO, 26.5733934455) +- (0, 0.320312180261336)
+	(JGroups-nocoord, 1.7777051021) +- (0, 0.0190481557873808)
+	(JGroups-coord, 1.7959073697) +- (0, 0.0189017779847593)
 };

### 2017-02-06 report/figures/synthetic-churn/total-events-sent-churn.tex
Start working on report
Correct typos in report

--- 2016-12-01 report/figures/total-events-sent-churn.tex
@@
 	% Receive
-	(EpTO, 54169.6) +- (0, 158.689214084218)
-	(JGroups-nocoord, 53981.9) +- (0, 102.155056436554)
-	(JGroups-coord, 53953.4) +- (0, 180.331300050151)
+	(EpTO, 53898.5) +- (0, 133.952935350029)
+	(JGroups-nocoord, 53830.5) +- (0, 200.355933278752)
+	(JGroups-coord, 53834.7) +- (0, 175.463861426411)
 };
@@
 	% Receive
-	(EpTO, 59853.9) +- (0, 228.339148344445)
-	(JGroups-nocoord, 57813.8) +- (0, 128.414087146925)
-	(JGroups-coord, 57873.3) +- (0, 150.893671172785)
+	(EpTO, 59798.6) +- (0, 140.107260498678)
+	(JGroups-nocoord, 59450.5) +- (0, 175.171820412607)
+	(JGroups-coord, 59507.9) +- (0, 240.914669079693)
 };

### 2017-02-06 report/framework.tex
Correct typos in report

@@
 Every benchmark is started through a Python script available on the master node.
-This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results through the framework python classes.
+This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results through the framework Python classes.
 Application, churn and cluster  parameters are customizable through YAML configuration files.
@@
 For example, \epto uses HTTP, while \jgroups uses TCP directly.
+Tracker implementation is also left to the user, although a REST Tracker implemented in Python is provided by default.
+The user can also opt to not use a tracker at all.
 Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.
@@
 Synthetic churn follows strict periods.
-Every period of time, it modifies the cluster size according to the instructions, until there are no more steps.
-These instructions can either, do nothing, increase the cluster size, reduce the cluster size or both at the same time.
+Every period of time, it modifies the cluster size according to the instructions until there are no more steps.
+These instructions can either: do nothing, increase the cluster size, reduce the cluster size or both at the same time.
 
 Real traces are based on FTA traces as stated earlier.
-Every time we query the database, we obtain every cluster change that happened since the last query according to the trace.
-Thus the smaller the period, the more accurate it is regarding to the real trace.
+Every time we query the database, we obtain every cluster change that happened since the last query according to the trace and the time elapsed between both queries.
+Thus, the smaller the period, the more accurate it is regarding to the real trace.
 \subsection{Extensions}
@@
 \jgroups relies on MULTICAST and has a coordinator that order every events sent.
-\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need.
+\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need, thanks to class inheritance.
 In our experiments we wrote an extension of the Churn module, so that it can deal with the coordinator, killing it only when we want to.

### 2017-02-06 report/future-work.tex
Correct typos in report

@@
 \label{sec:future-work}
-\jt{Do we have Push-Pull references?}
 Future work should focus on testing \epto on a bigger cluster, preferably at a point where \jgroups can no longer function, as \epto is designed to work on very large number of peers.
 As it is, \epto does indeed work as intended in our infrastructure and under churn, but a larger test would showcase \epto's ability to handle large peer numbers.
+Some preliminary tests were done on AWS EC2\footnote{https://aws.amazon.com/}.
+Unfortunately we encountered a Docker bug that would not let us run our benchmarks.
+Some preliminary work have been done on the framework to use Kubernetes\footnote{https://kubernetes.io/} with some success.
 \par
@@
 This algorithm would use a Push-Pull approach to further reduce the needed number of balls in the system and thus reduce the network load.
+An early prototype has been designed and implemented, but not yet tested.

### 2017-02-06 report/introduction.tex
Change cluster parameters to evaluation section
Correct typos in report

@@
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the algorithm's scalability.
+Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the scalability of the algorithm.
 \par
@@
 \epto's dissemination component guarantees that all peers will receive an event with arbitrarily high probability.
-Once peers have received all events, \epto's ordering component orders them using the events timestamp, and in case of a tie use the broadcaster ID of the events.
+\epto's ordering component works alongside the dissemination component and orders the received events using their timestamp, and in case of a tie use the broadcaster ID attached to them.
 \par
@@
 With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes.
-Thanks to these approaches, \epto is highly resilient with  imperfect networks and highly scalable as the networks grow, while always providing total order.
+Thanks to these approaches, \epto is highly resilient with  imperfect networks and highly scalable as the network grow, while always providing total order.
 \subsection{Contributions}

### 2017-02-07 report/abstract.tex
general pass.

@@
 \begin{abstract}
-  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
-In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups.
-We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers.
+  \epto is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
+In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm provided by the \jgroups toolkit.
+We first compare them in a perfect environment assessing the scalability in terms of events throughput and number of peers.
 We then compare them using a synthetic churn trace and a real one using the same metrics.
-Our results show \epto is performing as expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
+Our results show \epto is performing as expected, even though in the system sizes we tested \jgroups is still more efficient in a stable environment.
+By extrapolating our results for larger system sizes we can clearly see that \jgroups starts to degrade with the system size. 
+Still, further testing on a bigger cluster is required to evaluate how \epto performs when \jgroups is no longer able to keep up with scale.
   \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}

### 2017-02-07 report/background.tex
general pass.

@@
 We want them to appear as if it is only one database.
-To achieve this, every operation done a particular database would have to be replicated on all the others, we would then have to make sure that every operation is executed in the same order on every database to end up in the same state.
-However, Total Order is not usually scalable.
-This lead to the emergence of most of the work on eventual consistency.
-\epto claims to solve this issue.
+To achieve this, every operation done in a particular database would have to be replicated on all the others, thus we would have to make sure that every operation is executed in the same order on every database to end up in the same state.
+Total Order algorithms address this problem.
+However, Total Order is not usually scalable. This lead to the emergence of most of the work on eventual consistency. \epto claims to solve this issue.
 
@@
 \end{figure}
-\autoref{fig:epto-architecture} illustrates the architecture of a single \epto replica.
-An application extending our Application class can \epto-broadcast and \epto-deliver events.
+\autoref{fig:epto-architecture} illustrates the architecture of a single \epto process.
+An application using \epto can \epto-broadcast and \epto-deliver events.
 Events broadcasted to the Dissemination component are sent over the network every $\delta$ period, where $\delta$ is a unit of time.
-Every ball received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
+Every ball\mm{explain what a ball is} received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
 They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
-The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.
-The network layer is managed by \eptotester.
+The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.\mm{this terms are only introduced later in the definitions section, reorganize} The network layer is managed by \eptotester.
 \subsection{Related Work}
@@
 However \textit{PABCast} only offers probabilistic order whereas \epto offers deterministic order.
-Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages.
+Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto does not have such restriction.
 This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases, making it less desirable than \epto.
@@
 Both \autocite{Chandra2007} and \autocite{Maia2011} extensively test a particular large scale distributed system.
-However, they do not offer a framework to benchmark different systems or use SPlay \autocite{Leonini2009}.
+However, they do not offer a framework to benchmark different systems such as SPlay \autocite{Leonini2009}.
 SPlay is a framework used to deploy distributed applications.
 Unfortunately, SPlay only allows for LUA applications to run.
-\eptotester offers a convenient framework to deploy distributed softare regardless of the programming language used, provided they run in a docker container.
+\eptotester offers a convenient framework to deploy distributed software regardless of the programming language used, provided they run in a docker container.
 This makes it easier for existing protocol implementations to be benchmarked using our framework.

### 2017-02-07 report/conclusion.tex
general pass.

@@
 Due to this limitation, \epto performs worse than \jgroups.
-We want to reiterate that this is expected as \epto is designed for large distributed networks.
+Still, by extrapolating the observed behavior of \epto and \jgroups to a larger number of peers than we tested we can clearly see that \jgroups, albeit being more efficient, scales poorly than \epto.
+Further experiments, in larger settings are needed to confirm this.

### 2017-02-07 report/definitions.tex
Correct typos in report
general pass.

@@
 \label{sec:definitions}
-\jt{I think event should stay in this section what do you think?}
+\mm{several of these terms appear before, this needs to be moved upwards.}
 A process or peer is defined as an actor in our system running the application that needs total order.

### 2017-02-07 report/evaluation.tex
general pass.

@@
 \textbf{Experiment parameters.} Every benchmark except the ones following a real trace are run 10 times, each during \SI{20}{\minute}.
-When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}.
+When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and runs for \SI{17}{\minute}.
 The benchmarks following a real trace are run 5 times.
@@
 \par
-In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns.
+In \autoref{fig:bandwidth-churn} we analyze two different synthetic churns.
 In the first one we kill one node per minute.
@@
 We can see that the churn does not affect \epto at all when there are only nodes leaving.
-We have small peaks when adding a node to \SI{3}{\mbps} or less.
-Probably due to running the PSS initialization method.
+We have small peaks when adding a node to \SI{3}{\mbps} or less which can be explained by the PSS initialization.
 This is confirmed at the end of the plot where \epto goes back to a normal Bandwidth after stabilization.
@@
 On the other hand, when killing the coordinator in \jgroups we can see a huge spike in bandwidth, going from \SI{1.2}{\mbps} to more than \SI{15}{\mbps}.
-This is due to how \jgroups operates when selecting a new coordinator.
+This is due to how \jgroups operates when selecting a new coordinator which requires detecting the coordinator failure, electing a new coordinator and propagating these changes.
 
@@
 This is due to \jgroups having to update the view and propagate it to every peer.
+
+These results are quite interesting as they confirm two common held beliefs in the community: deterministic protocols behave pooly under churn, and probabilistic protocols are largely unaffected by churn.
 
@@
 We also do not see a difference whether we kill the coordinator or not.
-This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}.
-The big spike afterwards compensates for this hole.
+This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second} making the system unavailable.
 \subsection{Local Times}
 \label{sub:local-times}
-The local time represents the total time it takes for a peer to deliver every .
+The local time represents the total time it takes for a peer to deliver every \mm{missing text}.
  \begin{figure*}[hpt]
@@
 \subsection{Global Times}
-The global time represents the total time it takes for all peers to deliver every events sent.
+The global time represents the total time it takes for all peers to deliver every event sent.\mm{this is imprecise, review. same for local time}
  \begin{figure*}[hpt]
@@
 These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.
+\mm{explain why the fact that the system system stops for about 20 seconds when the coordinator dies is not observable in these measurements. }
 
@@
  \end{figure}
-In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch.
+In \autoref{fig:local-delta}, we see the percentiles of the local dissemination stretch.
 \par
@@
 \subsection{Events sent}
+\mm{explain first what this is before pointing out what was observed.}
 The variation observed in the different tests are due to the fact that we use \textit{Thread.sleep()} between each send and that in some configurations a randomness decides whether an event is sent or not.
@@
 \end{table}
-In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events.
-When killing and adding nodes, \jgroups seems to deliver a smaller amount of nodes, however it does not look significant enough to draw any conclusion.
-\jt{I didn't run any statistical analysis}
+In \autoref{table:total-events-churn} When only killing nodes, \epto and \jgroups again deliver the same amount of events. When killing and adding nodes, \jgroups delivers a smaller amount of events.
+This is explained by the fact that \jgroups nodes take longer to start but the differences do not look significant enough to draw any conclusion. \jt{I didn't run any statistical analysis}
 \subsection{Real Trace}

### 2017-02-07 report/framework.tex
general pass.

@@
 \eptotester is a framework using Docker and Docker Swarm.
-It is designed to ease the deployment of large scale distributed systems, more particularly \jgroups and \epto.
+It is designed to ease the deployment of large scale distributed systems and in this particular case \jgroups and \epto.
 We use \eptotester to assess the claims made in \autocite{matos2015epto}.
@@
 	\vspace{-2mm} 
-	\caption[Caption]{\epto using the \eptotester framework\footnotemark}
+	\caption[Caption]{\epto using the \eptotester\mm{this needs to be explained in more detail in the text in particular the concept of different services, the benchmark, etc.
+the churn modules are also missing} framework\footnotemark}
 	\vspace{-2mm} 
@@
 Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}.
-The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
+The synthetic churn provided by it was improved to support adding and removing nodes at the same time\mm{point out that when this is simultaneous it applies to different sets of nodes.}.
 
@@
 This is the case for \jgroups.
-\jgroups relies on MULTICAST and has a coordinator that order every events sent.
+\jgroups relies on MULTICAST and has a coordinator that orders every events sent.
 \eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need, thanks to class inheritance.
@@
 When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library as a tracker instead of the traditional MULTICAST option to coordinate peers, as Docker services do not yet support MULTICAST.
-This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
+This is not a problem as in a real WAN \jgroups could not rely on MULTICAST\mm{reference the jgroups page where they mention this}.
 
-The \epto simulation and theory is assuming we can have balls of infinite size.
-In practice, packet loss happens.
+The \epto simulation and theory is assuming we can have balls of infinite size which of course is not possible in practice.
 We must therefore limit the ball size to a practical number.
@@
 
-When benchmarking \jgroups we have to send a SIGSTOP to the containers instead of killing them, as otherwise the kernel closes the TCP connection to the TCPGOSSIP nicely, making the leave uneventful.
+When benchmarking \jgroups we have to send a SIGSTOP to the containers instead of killing them, as otherwise the kernel closes the TCP connection to the TCPGOSSIP nicely, making it appear as a regular leave instead of a failure.
 SIGSTOP lets us simulate a harsh connection drop.

### 2017-02-07 report/introduction.tex
general pass.

@@
 There are deterministic algorithms, which guarantee total order, agreement or other strong properties.
-Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.
+Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}\mm{change the citation styles to numbers}.
 \par
@@
 This causes a massive amount of network traffic and overhead on the system.
-Moreover, an agreement feature for an asynchronous system requires to
+Moreover, an agreement property for an asynchronous system requires to
 explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}.
-Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the scalability of the algorithm.
+Due to faults and churn in large-scale distributed systems, the failure detector uncertainty turns into a bottleneck thus limiting the scalability of the algorithm.
 \par
 As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach  \autocite{birman1999bimodal,carvalho2007emergent,demers1987epidemic,eugster2003lightweight,felber2002probabilistic,hayden1996probabilistic,kim2004gossip,Koldehofe02simplegossiping}.
-These algorithms guarantee the dissemination of events in the system with high probability.
-This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable.
+These algorithms guarantee the dissemination of events in the system with high probability and due to their inherent redundancy work well even under churn, faults and message loss.
 As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.
@@
 	\includegraphics[width=\linewidth]{figures/BnB.jpeg}
-	\caption[Caption]{Balls-and-Bins\footnotemark}
+	\caption[Caption]{Balls-and-Bins\mm{the figure needs to be referenced and described, otherwise it's best to remove it.}\footnotemark}
 	\label{fig:balls-and-bins}
@@
 Until now, the creators of \epto have only tested this algorithm in a simulated environment.
-In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking framework called \eptotester to show that \epto is suitable for real-world large-scale distributed systems.
+In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking framework called \eptotester\mm{why not call this a generic name since it's partially used for testing JGroups as well ?} to show that \epto is suitable for real-world large-scale distributed systems.
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
 \epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view.
-Peers use a tracker that keeps track of dead and alive peers when they first join the network.
+Peers discover other peers through a tracker that keeps track of dead and alive peers when they first join the network.
+Note that this is only used for bootstrap and is not required in the regular operation of the algorithm.
 These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm.
@@
 The next sections present our benchmarking solution and the results obtained.
-Section \ref{sec:background} presents the different kind of ordering, \epto architecture and related work.
+Section \ref{sec:background} presents the \epto architecture and related work.
 Section \ref{sec:definitions} defines the terms used in the paper.
-Section \ref{sec:epto} presents \epto and the architecture of our implementation.
+Section \ref{sec:epto} presents \eptotester and the architecture of our implementation.
 Section \ref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced.

### 2017-02-07 report/background.tex
Fix report based on Miguel's feedback

@@
 Events broadcasted to the Dissemination component are sent over the network every $\delta$ period, where $\delta$ is a unit of time.
-Every ball\mm{explain what a ball is} received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
+Every ball received from the network is unwrapped and its events are analyzed by the dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL).
 They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order.
-The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.\mm{this terms are only introduced later in the definitions section, reorganize} The network layer is managed by \eptotester.
+The order is based on the \textit{timestamp} of the events given by the logical clock and their \textit{broadcasterID} in case of a tie.
+The network layer is managed by \sys.
 \subsection{Related Work}
@@
 Unfortunately, SPlay only allows for LUA applications to run.
-\eptotester offers a convenient framework to deploy distributed software regardless of the programming language used, provided they run in a docker container.
+\sys offers a convenient framework to deploy distributed software regardless of the programming language used, provided they run in a docker container.
 This makes it easier for existing protocol implementations to be benchmarked using our framework.

### 2017-02-07 report/conclusion.tex
Fix report based on Miguel's feedback

--- 2017-02-06 report/conclusion.tex
@@
 
-In this paper, we implement \eptotester a framework designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network.
+In this paper, we implement \sys a framework designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network.
 Furthermore, we implement \epto and evaluate the claims made in \autocite{matos2015epto} against \jgroups SEQUENCER, a total order algorithm relying on a coordinator.
@@
 Due to this limitation, \epto performs worse than \jgroups.
-We want to reiterate that this is expected as \epto is designed for large distributed networks.
+Still, by extrapolating the observed behavior of \epto and \jgroups to a larger number of peers than we tested we can clearly see that \jgroups, albeit being more efficient, scales worse than \epto.
+Further experiments, in larger settings are needed to confirm this.

### 2017-02-07 report/evaluation.tex
Fix report based on Miguel's feedback

@@
 \par
-\textbf{Testbed.} \mm{this sentence is not necessary} \jt{I specified it as the following text is really close to the one in Sebastien work as we use the same cluster and I don't want any problem being accused of copy}Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
+\textbf{Testbed.} Our work uses the same cluster first used in \autocite{vaucher2016erasure}.
 It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network.
@@
 \subsection{Total GigaBytes sent/received}
-This metric shows the Total number of bytes sent and received on average during an experiment.
+This metric shows the total number of bytes sent and received on average during an experiment.
 In all cases, \jgroups receives more bytes than it sends.
@@
 \label{sub:local-times}
-The local time represents the total time it takes for a peer to deliver every \mm{missing text}.
+The local time represents the total time elapsed locally for a peer, between its first event sent and the last event it has delivered to the application.
  \begin{figure*}[hpt]
@@
 \subsection{Global Times}
-The global time represents the total time it takes for all peers to deliver every event sent.\mm{this is imprecise, review. same for local time}
+The global time represents the total time elapsed, between the earliest event sent and the latest event delivered to the application across all peers.
  \begin{figure*}[hpt]
@@
 These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.
-\mm{explain why the fact that the system system stops for about 20 seconds when the coordinator dies is not observable in these measurements. }
 
-Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times}.
+Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times} even when the coordinator dies and the system grinds to a halt for more than \SI{20}{\second}.
+This is caused by the fact that on our testbed \jgroups is not limited bandwidth-wise.
+This means \jgroups can just quickly send events at a higher global event throughput per second for a moment to catch up.
+In an environment where not as much bandwidth would be available to \jgroups the global times and local times would probably be affected by this issue.
 \subsection{Local Dissemination stretch}
@@
 \subsection{Events sent}
-\mm{explain first what this is before pointing out what was observed.}
-The variation observed in the different tests are due to the fact that we use \textit{Thread.sleep()} between each send and that in some configurations a randomness decides whether an event is sent or not.
+The events sent metrics shows the average number of events sent during an experiment run.
+The variation observed in the different tests are due to the fact that we use \textit{Thread.sleep()} between each send and that in some configurations a Random Number Generator decides whether an event is sent or not during a period.
 \begin{table}[hpt]

### 2017-02-07 report/framework.tex
Fix report based on Miguel's feedback

@@
 \label{sec:epto}
-\eptotester is a framework using Docker and Docker Swarm.
+\sys is a framework using Docker and Docker Swarm.
 It is designed to ease the deployment of large scale distributed systems and in this particular case \jgroups and \epto.
-We use \eptotester to assess the claims made in \autocite{matos2015epto}.
+We use \sys to assess the claims made in \autocite{matos2015epto}.
 Although the implementation of \epto is written with benchmarking in mind, the code can be adapted for real applications with only minimal changes to the sources.
-We  use \eptotester instead of other solutions such as SPlay, which only supports LUA, as we need to test existing protocols written in different programming languages.
+We  use \sys instead of other solutions such as SPlay, which only supports LUA, as we need to test existing protocols written in different programming languages.
 \subsection{Deployment}
@@
 	\vspace{-2mm} 
-	\caption[Caption]{\epto using the \eptotester\mm{this needs to be explained in more detail in the text in particular the concept of different services, the benchmark, etc.
-the churn modules are also missing} framework\footnotemark}
+	\caption[Caption]{\epto using the \sys framework\footnotemark}
 	\vspace{-2mm} 
@@
 \end{figure}
+\mm{this needs to be explained in more detail in the text in particular the concept of different services, the benchmark, etc. the churn modules are also missing}
 \footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
 To deploy a protocol in our framework we need two different Docker Swarm services: one containing the tracker and one containing all replicas.
+We use two services to isolate the tracker from the peers.
+This lets us kill peers without having to worry about the tracker and also scale the tracker in case we need replication for load-balancing purposes.
 Both of these services use the same network overlay to communicate.
 This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
-This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications.
+This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications as well as a performant load-balancing of our nodes at no costs.
 Every benchmark is started through a Python script available on the master node.
@@
 \subsection{Extensions}
-\eptotester is a general framework and should satisfy most decentralized applications needs with almost no tweaking.
+\sys is a general framework and should satisfy most decentralized applications needs with almost no tweaking.
 However, some protocols are centralized.
 This is the case for \jgroups.
-\jgroups relies on MULTICAST and has a coordinator that orders every events sent.
-\eptotester is written such that the main Benchmark and Churn modules are extensible to suit any need, thanks to class inheritance.
+\jgroups relies on MULTICAST\footnote{http://jgroups.org/overview.html} and has a coordinator that orders every events sent.
+\sys is written such that the main Benchmark and Churn modules are extensible to suit any need, thanks to class inheritance.
 In our experiments we wrote an extension of the Churn module, so that it can deal with the coordinator, killing it only when we want to.
@@
 
-\eptotester supports the use of a tracker as a separate service.
+\sys supports the use of a tracker as a separate service.
 If the application using the framework needs a tracker it only has to provide the docker image name for the tracker in the application configuration YAML file for the tracker to be created alongside the application service.
@@
 When implementing \jgroups we use TCPGOSSIP \autocite{tcpgossip} provided by the \jgroups library as a tracker instead of the traditional MULTICAST option to coordinate peers, as Docker services do not yet support MULTICAST.
-This is not a problem as in a real WAN \jgroups could not rely on MULTICAST\mm{reference the jgroups page where they mention this}.
+This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
 

### 2017-02-07 report/introduction.tex
Fix report based on Miguel's feedback

@@
 There are deterministic algorithms, which guarantee total order, agreement or other strong properties.
-Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}\mm{change the citation styles to numbers}.
+Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.
 \par
@@
 In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?
+\autoref{fig:balls-and-bins} shows the \textit{balls-and-bins} problem.
 \begin{figure}
 	\includegraphics[width=\linewidth]{figures/BnB.jpeg}
-	\caption[Caption]{Balls-and-Bins\mm{the figure needs to be referenced and described, otherwise it's best to remove it.}\footnotemark}
+	\caption[Caption]{Balls-and-bins problem\footnotemark}
 	\label{fig:balls-and-bins}
@@
 Until now, the creators of \epto have only tested this algorithm in a simulated environment.
-In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking framework called \eptotester\mm{why not call this a generic name since it's partially used for testing JGroups as well ?} to show that \epto is suitable for real-world large-scale distributed systems.
+In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking framework called \sys to show that \epto is suitable for real-world large-scale distributed systems.
 We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments.
@@
 Furthermore, we can inject synthetic or real churn using the Failure Trace Archive databases\footnote{\href{http://fta.scem.uws.edu.au/}{http://fta.scem.uws.edu.au/}} in the system.
-\eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
+\sys is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
 Finally we want to emphasize that few works evaluate such protocols in a really large scale \autocites[]{Chandra2007}[]{Maia2011}.
@@
 The next sections present our benchmarking solution and the results obtained.
+Section \ref{sec:definitions} defines the terms used in the paper.
 Section \ref{sec:background} presents the \epto architecture and related work.
-Section \ref{sec:definitions} defines the terms used in the paper.
-Section \ref{sec:epto} presents \eptotester and the architecture of our implementation.
+Section \ref{sec:epto} presents \sys and the architecture of our implementation.
 Section \ref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced.

### 2017-02-07 report/framework.tex
Fix more

@@
 We use two services to isolate the tracker from the peers.
-This lets us kill peers without having to worry about the tracker and also scale the tracker in case we need replication for load-balancing purposes.
+This lets us kill and add peers without having to worry about the tracker and also scale the tracker in case we need replication for load-balancing purposes.
 Both of these services use the same network overlay to communicate.
 This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12.
-This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications as well as a performant load-balancing of our nodes at no costs.
+This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications as well as a performant load-balancing of our nodes at no cost.
 Every benchmark is started through a Python script available on the master node.
-This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results through the framework Python classes.
+This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results through the framework Python module.
 Application, churn and cluster  parameters are customizable through YAML configuration files.
@@
 The user can also opt to not use a tracker at all.
+\sys generates at least one log file per peer.
+Any important benchmarking information must be logged to a file to be later exploited.
 Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building.

### 2017-02-07 presentation/slides.tex
Fix typos in presentation

@@
 			\end{itemize}
-		\item Provides deterministic total order
-		\item Scales wells with the number of peers
+		\item Provides deterministic total order,  integrity and validity
+		\item Scales well with the number of peers
 			\begin{itemize}
@@
 \begin{frame}{Evaluation}
-    We evaluate \epto{} against \jgroups{}, scaling peers and global event throughput per second.
+    We evaluate \epto{} against \jgroups{} SEQUENCER, scaling peers and global event throughput per second.
 

### 2017-02-08 report/evaluation.tex
Fix figures and last typos

@@
 In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups.
-This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect.
+This is expected as \epto sends $c \cdot n \cdot \log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c  \cdot \log_2 n$ more messages sent in \epto if \jgroups is perfect.
 Here we are well within this ratio.
@@
 \end{table}
-\autoref{table:total-bandwidth-real-churn} also confirms our earlier results. \epto is still within the $c*log_2(n)$ ratio fixed earlier.
+\autoref{table:total-bandwidth-real-churn} also confirms our earlier results. \epto is still within the $c \cdot log_2(n)$ ratio fixed earlier.
 \par
@@
 These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies.
-These problems appeared 3 times out of 13 runs for \jgroups $1\{kill,add\}/minute$ and 2 times out of 7 for \jgroups with real churn.
+These problems appeared 3 times out of 13 runs using the scenario and 2 times out of 7 for \jgroups following a real trace.
 We have not investigated these problems further.

### 2017-02-08 report/figures/architecture.tex
Correct typos in report
Fix figures and last typos

@@
 \node[draw, text width=2.1cm, text centered] (bench) at (2.1, 2.6) {Benchmarks runner};
-\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python script}};
-\draw[dotted,red,thick] (0.8,2) rectangle (3.4,3.4);
+
+\draw (0.8,2.05) node[below right,red,scale=0.8] {\textit{LSDSuite module}};
+\draw[red] (0.9,0.1) rectangle (3.3,2.05);
+\node[draw, text width=2cm, text centered] (churnc) at (2.1,1.4) {Churn Class};
+\node[draw, text width=2cm, text centered] (benchc) at (2.1, 0.6) {Benchmark Class};
+\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python scripts}};
+\draw[dotted,red,thick] (0.8,0) rectangle (3.4,3.4);
 %
@@
 \draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
-\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto-tracker:}};
+\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto-tracker}};
 \draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);

### 2017-02-08 report/framework.tex
Fix figures and last typos

@@
 Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}.
-The synthetic churn provided by it was improved to support adding and removing nodes at the same time\mm{point out that when this is simultaneous it applies to different sets of nodes.}.
+The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
+When removing and adding an equal amount of nodes, both instructions execute without canceling each other.
+\mm{point out that when this is simultaneous it applies to different sets of nodes.}.
 

### 2017-02-08 report/main.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues
Fix more stuff pointed by Hugues
General comments.
Rework framework and background sections
Correct typos in report
Fix report based on Miguel's feedback
Fix figures and last typos

@@
 \usepackage{csquotes}
+\usepackage{cleveref}
 \usepackage[nolist]{acronym}
@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{url}
 \usepackage{xspace}
@@
 \makeatother
+
 %% Environment for comments: Set the boolean to false to produce a comment-free version.
 \newboolean{showcomments}
-\setboolean{showcomments}{true}
+\setboolean{showcomments}{false}
 \ifthenelse{\boolean{showcomments}}
@@
 \newcommand{\hm}[1]{\mynote{Hugues}{#1}{orange}}
+\newcommand{\mm}[1]{\mynote{Miguel}{#1}{gray}}
 
@@
 \newcommand{\jgroups}{\textsc{JGroups}\xspace}
+\newcommand{\sys}{\textsc{LSDSuite}\xspace}
 
@@
 
-\def\ieeetitle{EpTO TO DEFINE}
+\def\ieeetitle{EpTOTester}
 \def\ieeeauthor{X,X,X}
 
-\def\thesistitle{EpTO TO DEFINE}
-\def\thesissubtitle{X}
+\def\thesistitle{\epto}
+\def\thesissubtitle{Implementation of a Large-Scale Epidemic Total Order Algorithm
+}
 \def\thesisauthor{Jocelyn Thode}
@@
 			\vfill
+
+			\vspace{0.3in}
 			{\Large \today \par}
@@
 		\end{center}
+
+	\pagebreak
+
+	\section*{Abstract}
+
+	\noindent 
+	\epto is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time.
+In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm provided by the \jgroups toolkit.
+We first compare them in a perfect environment assessing the scalability in terms of events throughput and number of peers.
+We then compare them using a synthetic churn trace and a real one using the same metrics.
+	Our results show \epto is performing as expected, even though in the system sizes we tested \jgroups is still more efficient in a stable environment.
+	By extrapolating our results for larger system sizes we can clearly see that \jgroups starts to degrade with the system size. 
+	Still, further testing on a bigger cluster is required to evaluate how \epto performs when \jgroups is no longer able to keep up with scale.
+
+	\bigskip
+
+	Dr Hugues Mercier, Institutes of Computer Science and Mathematics, University of Neuchâtel, Supervisor
+
+	Dr Miguel Matos, HASLab - INESC TEC \&, University of Minho, Portugal,  Co-supervisor
+
 	\end{otherlanguage}
@@
 \input{introduction.tex}
-\input{ordering.tex}
 \input{definitions.tex}
+\input{background.tex}
 \input{framework.tex}
@@
 \input{conclusion.tex}
+
 \printbibliography
+
 \end{document}

### 2017-02-08 report/main.tex
Change Miguel's University

@@
 
-	Dr Hugues Mercier, Institutes of Computer Science and Mathematics, University of Neuchâtel, Supervisor
+	Dr Hugues Mercier, Institutes of Computer Science and Mathematics, Université de Neuchâtel, Supervisor
 
-	Dr Miguel Matos, HASLab - INESC TEC \&, University of Minho, Portugal,  Co-supervisor
+	Dr Miguel Matos, INESC-ID/IST, Universidade de Lisboa, Portugal,  Co-supervisor
 
